### 2015-05-02 MSDgroupwriteup.tex
modified script to get sparsity upates to tex doc

% \documentclass[12pt]{emulateapj}
 \documentclass[12pt,preprint]{aastex}
\usepackage[margin= 1.0in]{geometry}    % See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper} % or letter or a5paper or ... etc
\usepackage{float}
\usepackage{amssymb,amsmath}
\usepackage[]{epsfig,graphicx}
\usepackage{color}
\usepackage{verbatim}
\DeclareGraphicsRule{.tif}{png}{.jpg}{`convert Num1 `dirname Num1`/`basename Num1 .tif`.jpg}
\newcommand{\units}[1]{\ensuremath{\, \mathrm{#1}}}
\newcommand{\myemail}{egrand@astro.umd.edu}
\usepackage{enumitem}
\usepackage{natbib,twoopt}
\newcommand{\degree}{\ensuremath{^\circ}}
\citestyle{apj}
\bibliographystyle{aa}
\usepackage[maxfloats=25]{morefloats}
\usepackage[normalem]{ulem}
\usepackage{hyperref}

\begin{document}
\title{Clustering of Last.fm Music Artists}

 \author{Jordan Rosenblum, Justin Law \& Erin Grand}
 \affil{Data Science Institute, Columbia University, New York, NY 10027}

\date{\today}

\begin{abstract}
This is the abstract. 
\end{abstract}

This is the body.

\href{http://www.dtic.upf.edu/~ocelma/MusicRecommendationDataset/}{Music Recommendation Dataset}

Sparsity of $8.08199453451\times 10^{-5}$

\end{document}

### 2015-05-02 lastfm-erin.py

@@ -63,3 +63,6 @@ count = len(count[0])
 print count
 print len(users)*len(artists)
 print float(count)/(len(users)*len(artists))
+
+
+

### 2015-05-03 lastfm_db.py
get track data

@@ -1,0 +1,72 @@
+import sqlite3
+
+conn = sqlite3.connect('lastfm_tags.db')
+
+
+
+# Get list of all tags for a particular track id
+tid = 'TRNIEVD128F147645F'
+print 'We get all tags (with value) for track: %s' % tid
+sql = "SELECT tags.tag, tid_tag.val FROM tid_tag, tids, tags WHERE tags.ROWID=tid_tag.tag AND tid_tag.tid=tids.ROWID and tids.tid='%s'" % tid
+res = conn.execute(sql)
+data_tags = res.fetchall()
+print data_tags
+
+
+conn.close()
+
+
+
+# Create dictionaries for tracks to artists and titles
+f = open('unique_tracks.txt', 'r')
+data = []
+for line in f:
+    data.append(line)
+f.close()
+
+track_to_artist = {}
+track_to_title = {}
+song_to_track = {}
+
+for i in range(len(data)):
+    split = data[i].split('<SEP>')
+    track_to_artist[split[0]] = split[2]
+    track_to_title[split[0]] = split[3]
+    song_to_track[split[1]] = split[0]
+
+
+
+# Map songs id to track id
+f = open('kaggle_visible_evaluation_triplets.txt', 'r')
+data = []
+for line in f:
+    data.append(line)
+f.close()
+    
+
+users = set()
+songs = set()
+
+for i in range(len(data)):
+    split = data[i].split('\t')
+    userhash = split[0]
+    songhash = split[1]
+    users.add(userhash)
+    songs.add(songhash)
+        
+users = list(users) #user hashes
+songs = list(songs) #artist hashes
+
+i=0
+for song in songs:
+    if song not in song_to_track.keys():
+        i = i + 1
+
+
+
+
+
+
+
+
+

### 2015-05-03 lastfm_db.py
song <- track dict

@@ -57,13 +57,23 @@ for i in range(len(data)):
 users = list(users) #user hashes
 songs = list(songs) #artist hashes
 
-i=0
+# make sure that all the songs in the triplet set are in the dictonary
+#i=0
+#for song in songs:
+#    if song not in song_to_track:
+#        i = i + 1
+
+# dictonary from trackID to songID for only the songs in triplets
+
+triplet_song_to_track = {}
+triplet_track_to_artist = {}
+triplet_track_to_title = {}
+
 for song in songs:
-    if song not in song_to_track.keys():
-        i = i + 1
-
-
-
+    track = song_to_track[song]
+    triplet_song_to_track[song] = track
+    triplet_track_to_artis = track_to_artist[track]
+    triplet_track_to_title = track_to_title[track]
 
 
 

### 2015-05-03 MSDgroupwriteup.tex
small edits
new thoughts

@@
 
-Sparsity of $8.08199453451\times 10^{-5}$
+Sparsity of $8.08199453451\times 10^{-5}$ need updating
+
+The matrix was too big for our computers to handle the matrix factorization using the whole set of users and songs, so we subset the data.
+
+The full matrix was originally $163206$ users x $110000$ songs.
+
+Subset songs by how many times the song was listened to. i.e the popularity of the song
+= new number
+
+Subset users by how many songs they've listened to (how many popular song?)
+= new number
+
+How to split into training and test? 
+Random selection by user or element?
+Need matrixes to be the same size to compare them.
+
+Interesting plots:
+- Cumulative Distribution of number of songs with a given play count
+-
 

### 2015-05-03 MSDgroupwriteup.tex
more discussion

@@
 
-The full matrix was originally $163206$ users x $110000$ songs.
+The full matrix was originally $163206$ songs x $110000$ users.
 
@@
 
+SONG PLAY COUNT  \\
+mean = 28 \\
+max = 35432 \\
+std = 215.826789\\
+
+USER PLAY COUNT\\
+mean = 42\\
+max = 1305\\
+min = 5\\
+std = 53.31547\\
+
 How to split into training and test? 
-Random selection by user or element?
+Random selection by user or element? ELEMENT. 
 Need matrixes to be the same size to compare them.
@@
 - Cumulative Distribution of number of songs with a given play count
--
 

### 2015-05-03 MF_Analysis.py
get track data

@@ -1,0 +1,197 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from operator import itemgetter
+
+# Create M and M_test matrices here
+
+M = np.zeros([0,0])
+M_test = np.zeros([0,0])
+
+N1 = np.shape(M)[0]
+N2 = np.shape(M)[1]
+
+    
+omega = []
+    
+for i in xrange(0,N1):
+    for j in xrange(0,N2):
+        if M[i][j] != 0:
+            omega.append((i,j))
+            
+omega_test = []
+    
+for i in xrange(0,N1):
+    for j in xrange(0,N2):
+        if M_test[i][j] != 0:
+            omega_test.append((i,j))
+
+
+omegau = {}
+omegav = {}
+
+for (i,j) in omega:
+    if i in omegau.keys():
+        omegau[i].append(j)
+    else:
+        omegau[i] = [j]
+    if j in omegav.keys():
+        omegav[j].append(i)
+    else:
+        omegav[j] = [i]
+            
+var = 0.25
+d = 20
+I = np.identity(d)
+lamb = 10
+t1 = lamb*var*I
+
+mean = np.zeros([d])
+v = np.empty([N2,d])
+u = np.empty([N1,d])
+
+for j in xrange(0, N2):
+    v[j] = np.random.multivariate_normal(mean,(1/float(lamb))*I)
+
+def predict(u, v):
+    product = np.dot(u, v)
+    if product < 0:
+        pred = 0.0
+    elif product > 1:
+        pred = 1.0
+    else:
+        pred = product
+    return pred
+
+RMSE = []
+loglik = [] 
+           
+for iter in xrange(0,100):
+    print iter
+    for i in xrange(0, N1):
+        sum1 = np.zeros([d,d])
+        sum2 = np.zeros(d)
+        if i in omegau.keys():
+            for j in omegau[i]:
+                sum1 = sum1 + np.outer(v[j],np.transpose(v[j]))
+                sum2 = sum2 + np.dot(M[i][j],v[j])
+            u[i] = np.dot(np.linalg.inv(t1 + sum1), sum2)
+               
+    for j in xrange(0, N2):
+        sum1 = np.zeros([d,d])
+        sum2 = np.zeros(d)
+        if j in omegav.keys():
+            for i in omegav[j]:
+                sum1 = sum1 + np.outer(u[i],np.transpose(u[i]))
[…]

### 2015-05-03 normalize_by_column.py
normalize by song

@@ -1,0 +1,7 @@
+import numpy as np
+
+def scale_linear_bycolumn(matrix, high=1.0, low=0.0):
+    mins = np.min(matrix, axis=0)
+    maxs = np.max(matrix, axis=0)
+    rng = maxs - mins
+    return high - (((high - low) * (maxs - matrix)) / rng)

### 2015-05-03 DataFrames.py
Updating dataframes.py
Updated with plots of user and song data

@@ -1,10 +1,13 @@
 import numpy as np
 import sqlite3 as sqlite
 import pandas as pd
-import pandas.io.sql as psql
+import pandas.io.sql as psql.
+import matplotlib.pyplot as plt
+
 
 # http://www.kaggle.com/c/msdchallenge/data
-eval = pd.read_csv("kaggle_visible_evaluation_triplets.txt",sep='\t',header = None, names = ['user_id','artist_id','plays'])
+eval = pd.read_csv("kaggle_visible_evaluation_triplets.txt",sep='\t',header = None, names = ['user_id','song_id','plays'])
+#eval.pivot(index='user_id',columns='song_id', values='plays')
 
 # http://labrosa.ee.columbia.edu/millionsong/sites/default/files/AdditionalFiles/unique_tracks.txt
 unique_tracks = pd.read_csv("unique_tracks.txt",sep='<SEP>', header = None, names = ['track_id', 'song_id', 'arist_name', 'song_title'])
@@ -38,3 +41,61 @@ with con:
     track_metadata = psql.read_sql(sql, con)
 con.close()
 
+import matplotlib
+matplotlib.style.use('ggplot')
+
+import matplotlib.pyplot as plt
+
+# histogram of number of plays for users
+userhist = eval.groupby('user_id').sum()
+userhist = pd.DataFrame(userhist).reset_index()
+uservalues = userhist['plays'].values
+plt.hist(uservalues,50)
+plt.show()
+
+# cumulative sum of plays for users
+usercsum = userhist.groupby('plays').count().cumsum()
+usercsum = pd.DataFrame(usercsum).reset_index()
+usercsum.plot('plays','user_id')
+plt.show()
+
+# cumulative sum for plays for users in log
+usercsumlog = usercsum
+usercsumlog['plays'] = np.log10(usercsumlog['plays'])
+usercsumlog['user_id'] = usercsumlog['user_id']/usercsum['user_id'].max()
+usercsumlog.plot('plays','user_id')
+plt.show()
+
+# cumulative sum for plays vs users
+userplaycsum = userhist.sort('plays')
+userplaycsum['plays'] = userplaycsum['plays'].cumsum()
+userplaycsum['plays'] = userplaycsum['plays']/userplaycsum['plays'].max()
+userplaycsum.plot('user_id', 'plays')
+plt.show()
+
+# histogram of number of plays for songs
+songhist = eval.groupby('song_id').sum()
[…]

### 2015-05-03 DataFrames.py
Adding code to DataFrames.py to create matrix
Adding code to create matrix

@@ -1,13 +1,37 @@
 import numpy as np
 import sqlite3 as sqlite
 import pandas as pd
-import pandas.io.sql as psql.
+import pandas.io.sql as psql
 import matplotlib.pyplot as plt
 
 
 # http://www.kaggle.com/c/msdchallenge/data
 eval = pd.read_csv("kaggle_visible_evaluation_triplets.txt",sep='\t',header = None, names = ['user_id','song_id','plays'])
 #eval.pivot(index='user_id',columns='song_id', values='plays')
+
+userhist = eval.groupby('user_id').sum()
+userhist = pd.DataFrame(userhist).reset_index()
+usersub = userhist[userhist['plays']>100]
+
+songhist = eval.groupby('song_id').sum()
+songhist = pd.DataFrame(songhist).reset_index()
+songsub = songhist[songhist['plays']>100]
+
+sub = eval[eval['song_id'].isin(songsub['song_id'])]
+sub = sub[sub['user_id'].isin(usersub['user_id'])]
+
+len(set(sub['user_id']))
+len(set(sub['song_id']))
+
+sub.shape[0]/float(eval.shape[0])
+
+pivot = sub.pivot(index='user_id',columns='song_id', values='plays')
+user_index = pivot.index
+song_index = pivot.columns
+M = pivot.as_matrix()
+M = np.nan_to_num(M)
+M = np.asmatrix(M)
+M = M/np.max(M,axis=0)
 
 # http://labrosa.ee.columbia.edu/millionsong/sites/default/files/AdditionalFiles/unique_tracks.txt
 unique_tracks = pd.read_csv("unique_tracks.txt",sep='<SEP>', header = None, names = ['track_id', 'song_id', 'arist_name', 'song_title'])

### 2015-05-03 DataFrames.py
Adding last_fm tags loading to DataFrames.py

@@ -7,15 +7,14 @@ import matplotlib.pyplot as plt
 
 # http://www.kaggle.com/c/msdchallenge/data
 eval = pd.read_csv("kaggle_visible_evaluation_triplets.txt",sep='\t',header = None, names = ['user_id','song_id','plays'])
-#eval.pivot(index='user_id',columns='song_id', values='plays')
 
 userhist = eval.groupby('user_id').sum()
 userhist = pd.DataFrame(userhist).reset_index()
-usersub = userhist[userhist['plays']>100]
+usersub = userhist[userhist['plays']>29]
 
 songhist = eval.groupby('song_id').sum()
 songhist = pd.DataFrame(songhist).reset_index()
-songsub = songhist[songhist['plays']>100]
+songsub = songhist[songhist['plays']>34]
 
 sub = eval[eval['song_id'].isin(songsub['song_id'])]
 sub = sub[sub['user_id'].isin(usersub['user_id'])]
@@ -63,6 +62,14 @@ con = sqlite.connect("track_metadata.db")
 with con:
     sql = "SELECT * FROM songs"
     track_metadata = psql.read_sql(sql, con)
+con.close()
+
+# http://labrosa.ee.columbia.edu/millionsong/sites/default/files/lastfm/lastfm_tags.db
+# tag values indicate popularity of the tags on last.fm as a whole
+con = sqlite.connect("lastfm_tags.db")
+with con:
+    sql = "SELECT tags.tag, tids.tid, tid_tag.val FROM tid_tag, tids, tags WHERE tags.ROWID=tid_tag.tag AND tid_tag.tid=tids.ROWID"
+    lastfm_tags = psql.read_sql(sql, con)
 con.close()
 
 import matplotlib

### 2015-05-03 analysis_MF.py
MF

@@ -2,27 +2,26 @@ import numpy as np
 import matplotlib.pyplot as plt
 from operator import itemgetter
 
-# Create M and M_test matrices here
+# Create Mtrain and Mtest matrices here
 
-M = np.zeros([0,0])
-M_test = np.zeros([0,0])
+Mtrain = np.zeros([0,0])
+Mtest = np.zeros([0,0])
 
-N1 = np.shape(M)[0]
-N2 = np.shape(M)[1]
-
-    
+N1 = np.shape(Mtrain)[0]
+N2 = np.shape(Mtrain)[1]
+  
 omega = []
     
 for i in xrange(0,N1):
     for j in xrange(0,N2):
-        if M[i][j] != 0:
+        if Mtrain[i][j] != 0:
             omega.append((i,j))
             
 omega_test = []
     
 for i in xrange(0,N1):
     for j in xrange(0,N2):
-        if M_test[i][j] != 0:
+        if Mtest[i][j] != 0:
             omega_test.append((i,j))
 
 
@@ -38,8 +37,11 @@ for (i,j) in omega:
         omegav[j].append(i)
     else:
         omegav[j] = [i]
-            
-var = 0.25
+
+# Calculate var
+
+           
+var = ???????
 d = 20
 I = np.identity(d)
 lamb = 10
@@ -73,7 +75,7 @@ for iter in xrange(0,100):
         if i in omegau.keys():
             for j in omegau[i]:
                 sum1 = sum1 + np.outer(v[j],np.transpose(v[j]))
-                sum2 = sum2 + np.dot(M[i][j],v[j])
+                sum2 = sum2 + np.dot(Mtrain[i][j],v[j])
             u[i] = np.dot(np.linalg.inv(t1 + sum1), sum2)
                
     for j in xrange(0, N2):
@@ -82,20 +84,20 @@ for iter in xrange(0,100):
         if j in omegav.keys():
             for i in omegav[j]:
                 sum1 = sum1 + np.outer(u[i],np.transpose(u[i]))
-                sum2 = sum2 + np.dot(M[i][j],u[i])
+                sum2 = sum2 + np.dot(Mtrain[i][j],u[i])
             v[j] = np.dot(np.linalg.inv(t1 + sum1), sum2)
     
     sum3 = 0        
     for (i,j) in omega_test:
         prediction = predict(u[i], v[j])
-        actual = M_test[i][j]
[…]

### 2015-05-03 lastfm-erin.py
new code

@@ -1,17 +1,24 @@
+
+# coding: utf-8
+
+
+
 import numpy as np
 from collections import defaultdict
 from scipy import sparse
 
-
 f = open('kaggle_visible_evaluation_triplets.txt', 'r')
-users = set()
-artists = set()
-numplays = {}
-artistdic = defaultdict(list)
-userdic = defaultdict(list)
-
 data = f.read().split('\n')
 
+
+# In[2]:
+
+users = set()
+songs = set()
+numplays = {}
+numsongplayed = defaultdict(list)
+songdic = defaultdict(list)
+userdic = defaultdict(list)
 
 for line in data:
     split = line.strip().split('\t')
@@ -19,50 +26,115 @@ for line in data:
         print len(split), split
     else:
         userhash = split[0]
-        artisthash = split[1]
-        plays = split[2]
+        songhash = split[1]
+        plays = float(split[2])
         users.add(userhash)
-        artists.add(artisthash)
-        numplays[artisthash, userhash] = plays
-        #if artisthash not in userdic[userhash]:
-        userdic[userhash].append(artisthash)
-        #if userhash not in artistdic[artisthash]:
-        artistdic[artisthash].append(userhash)
+        songs.add(songhash)
+        numplays[songhash, userhash] = plays
+        if songhash not in userdic[userhash]:
+            userdic[userhash].append(plays)
+        if userhash not in songdic[songhash]:
+            songdic[songhash].append(userhash)
+        #if songhash not in numsongplayed:
+        numsongplayed[songhash].append(plays)
+
+
+# In[7]:
+
+# subset by number of songs a user has rated
+p1 = { key: value for key, value in userdic.items() if sum(value) > 29 }
+# subset by number of plays
+p2 = { key: value for key, value in numsongplayed.items() if sum(value) > 34 }
+
+# list of songs
+subsongs = p2.keys()
+
+# list of users who have rated those songs
+subusers = set()
+for s in subsongs:
+    for u in songdic[s]:
+        if u in p1:
+            subusers.add(u)
         
+p3 = { key: value for key, value in numplays.items() if key[0] in p2 and key[1] in p1 }
 
-n = len(users)
-m = len(artists)
+print len(subusers), len(users)
+print len(subsongs), len(songs)
[…]

### 2015-05-03 DataFrames.py
Adding sampling to DataFrames.py

@@ -3,6 +3,7 @@ import sqlite3 as sqlite
 import pandas as pd
 import pandas.io.sql as psql
 import matplotlib.pyplot as plt
+import random
 
 
 # http://www.kaggle.com/c/msdchallenge/data
@@ -24,13 +25,25 @@ len(set(sub['song_id']))
 
 sub.shape[0]/float(eval.shape[0])
 
-pivot = sub.pivot(index='user_id',columns='song_id', values='plays')
+sub_max = pd.DataFrame(sub.groupby('song_id').max()).reset_index()
+merged = pd.merge(sub,sub_max,on="song_id")
+merged['plays_x'] = merged['plays_x']/merged['plays_y']
+sub_norm = merged[['user_id_x', 'song_id', 'plays_x']]
+sub_norm.columns = ['user_id', 'song_id', 'plays']
+
+sample = random.sample(sub_norm.index, int(sub_norm.shape[0]*0.2))
+trainsub = sub_norm.copy()
+trainsub.ix[trainsub.index.isin(sample),'plays'] = 0
+
+testsub = sub_norm.copy()
+testsub.ix[~testsub.index.isin(sample),'plays'] = 0
+
+trainpivot = trainsub.pivot(index='user_id',columns='song_id', values='plays')
 user_index = pivot.index
 song_index = pivot.columns
-M = pivot.as_matrix()
-M = np.nan_to_num(M)
-M = np.asmatrix(M)
-M = M/np.max(M,axis=0)
+M_train = pivot.as_matrix()
+M_train = np.nan_to_num(M_train)
+
 
 # http://labrosa.ee.columbia.edu/millionsong/sites/default/files/AdditionalFiles/unique_tracks.txt
 unique_tracks = pd.read_csv("unique_tracks.txt",sep='<SEP>', header = None, names = ['track_id', 'song_id', 'arist_name', 'song_title'])

### 2015-05-04 DataFrames.py
Adding similarity data loading from last.fm

@@ -5,7 +5,6 @@ import pandas.io.sql as psql
 import matplotlib.pyplot as plt
 import random
 
-
 # http://www.kaggle.com/c/msdchallenge/data
 eval = pd.read_csv("kaggle_visible_evaluation_triplets.txt",sep='\t',header = None, names = ['user_id','song_id','plays'])
 
@@ -43,6 +42,11 @@ user_index = pivot.index
 song_index = pivot.columns
 M_train = pivot.as_matrix()
 M_train = np.nan_to_num(M_train)
+
+testpivot = testsub.pivot(index='user_id',columns='song_id', values='plays')
+M_test = pivot.as_matrix()
+M_test = np.nan_to_num(M_test)
+
 
 
 # http://labrosa.ee.columbia.edu/millionsong/sites/default/files/AdditionalFiles/unique_tracks.txt
@@ -84,6 +88,17 @@ with con:
     sql = "SELECT tags.tag, tids.tid, tid_tag.val FROM tid_tag, tids, tags WHERE tags.ROWID=tid_tag.tag AND tid_tag.tid=tids.ROWID"
     lastfm_tags = psql.read_sql(sql, con)
 con.close()
+
+# http://labrosa.ee.columbia.edu/millionsong/sites/default/files/lastfm/lastfm_similars.db
+# Similarity data from lastfm, the dest contains songs that consider the tid as similar while the src contains songs where tid considers as similar
+con = sqlite.connect("lastfm_similars.db")
+with con:
+    sql = "SELECT * FROM similars_dest"
+    lastfm_dest = psql.read_sql(sql, con)
+    sql = "SELECT * FROM similars_src"
+    lastfm_src = psql.read_sql(sql, con)
+con.close()
+
 
 import matplotlib
 matplotlib.style.use('ggplot')

### 2015-05-04 lastfm-erin.py
more updates

@@ -1,11 +1,14 @@
 
 # coding: utf-8
 
-
+# In[1]:
 
 import numpy as np
 from collections import defaultdict
 from scipy import sparse
+import matplotlib.pyplot as plt
+from operator import itemgetter
+
 
 f = open('kaggle_visible_evaluation_triplets.txt', 'r')
 data = f.read().split('\n')
@@ -39,7 +42,7 @@ for line in data:
         numsongplayed[songhash].append(plays)
 
 
-# In[7]:
+# In[3]:
 
 # subset by number of songs a user has rated
 p1 = { key: value for key, value in userdic.items() if sum(value) > 29 }
@@ -64,7 +67,7 @@ print 100. - (float(len(p3))/len(data) * 100)
 #print (float(len(subsongs)) * len(subusers)) / ( len(songs) * len(users)) * 100
 
 
-# In[8]:
+# In[4]:
 
 for key, value in p3.items():
     song = key[0]
@@ -78,7 +81,7 @@ for key, value in p3.items():
     p3[song, user] = normplay
 
 
-
+# In[5]:
 
 allkeys = p3.keys()
 
@@ -90,6 +93,7 @@ trainingKEYS, testKEYS = allkeys[:trainingN], allkeys[trainingN:]
 print len(trainingKEYS), len(testKEYS)
 
 
+# In[16]:
 
 n = len(subusers)
 m = len(subsongs)
@@ -112,13 +116,31 @@ for song in subsongs:
 Mtrain = np.zeros((n,m))
 Mtest = np.zeros((n,m))
 
+omega = []
+omega_test = []
+varcalc = []
+omegau = {}
+omegav = {}
+
 for key in trainingKEYS:
     plays = p3[key]
     song = key[0]
     user = key[1]
     userindex = user_dict[user]
     songindex = song_dict[song]
+    omega.append((userindex,songindex))
+    varcalc.append(float(plays))
     Mtrain[userindex, songindex] = float(plays)
+    i = userindex
+    j = songindex
+    if i in omegau:
+        omegau[i].append(j)
+    else:
+        omegau[i] = [j]
+    if j in omegav:
+        omegav[j].append(i)
+    else:
+        omegav[j] = [i]
 
 for key in testKEYS:
     plays = p3[key]
[…]

### 2015-05-04 normalize_by_column.py
Delete normalize_by_column.py

@@ -1,7 +1,0 @@
-import numpy as np
-
-def scale_linear_bycolumn(matrix, high=1.0, low=0.0):
-    mins = np.min(matrix, axis=0)
-    maxs = np.max(matrix, axis=0)
-    rng = maxs - mins
-    return high - (((high - low) * (maxs - matrix)) / rng)

### 2015-05-04 analysis_MF.py
comment out plot

@@ -4,26 +4,9 @@ from operator import itemgetter
 
 # Create Mtrain and Mtest matrices here
 
-Mtrain = np.zeros([0,0])
-Mtest = np.zeros([0,0])
-
 N1 = np.shape(Mtrain)[0]
 N2 = np.shape(Mtrain)[1]
   
-omega = []
-    
-for i in xrange(0,N1):
-    for j in xrange(0,N2):
-        if Mtrain[i][j] != 0:
-            omega.append((i,j))
-            
-omega_test = []
-    
-for i in xrange(0,N1):
-    for j in xrange(0,N2):
-        if Mtest[i][j] != 0:
-            omega_test.append((i,j))
-
 
 omegau = {}
 omegav = {}
@@ -38,10 +21,8 @@ for (i,j) in omega:
     else:
         omegav[j] = [i]
 
-# Calculate var
-
-           
-var = ???????
+# Calculate var          
+#var = ???????
 d = 20
 I = np.identity(d)
 lamb = 10

### 2015-05-04 lastfm-erin.py

@@ -242,7 +242,7 @@ for iter in xrange(0,10):
 
 # In[24]:
 
-get_ipython().magic(u'matplotlib inline')
+#get_ipython().magic(u'matplotlib inline')
 
 # Part 1: RMSE and likelihood
 fig = plt.figure()         

### 2015-05-04 MSDgroupwriteup.tex
updates
corrections
added Jordan's stuff to texfile

@@
 \newcommand{\units}[1]{\ensuremath{\, \mathrm{#1}}}
-\newcommand{\myemail}{egrand@astro.umd.edu}
 \usepackage{enumitem}
@@
 \usepackage{hyperref}
+\usepackage{ amssymb }
 
@@
 
-This is the body.
+\tableofcontents
+
+\section{Introduction}
+
+\section{Data Munging}
+
+\subsection{Matrix Factorization}
+The goal of matrix factorization is to use collaborative methods to build a recommendation system for users based on user ratings of objects (in this case songs).
+This will allow us to recommend certain songs to users based on their listening history and without the need for using content based approaches.
+Since our data set contains number of plays for a given user and song (rather than rating), we normalized plays for every song on a scale between 0 and 1 and used this as a proxy for rating.
+
+% Erin's part here on normalization.
+
+We then constructed training and testing matrices, of which both have $N_1$ users (rows denoted by $u_i$) and $N_2$ songs (columns denoted by $v_j$).
+Of course the matrices will be very sparse, containing zeros in all entries except for those in which a user (rows of matrix) has listened to a song (columns of matrix).
+The goal is to factor the training matrix into the product of two matrixes, $U$ and $V$.
+The matrix $U$ will be $N_1 \times d$ and the matrix $V$ will be $d \times N_2$.
+We want to learn a low-rank factorization (i.e.
+we choose $d$) so as to restrict the patterns we see in the rows and columns of our original matrix (e.g.
+we think a priori that if a user likes 1 top 100 song, the user may also like other top 100 songs).
+There is subjectivity in picking $d$ but $20$ is a common place to start.
+In the factorized matrices, the predicted rating will be $\hat{M}_{ij} = u_i^T v_j$
+
+Using a coordinate ascent algorithm over 100 iterations, each row ($u_i$) and column ($v_j$) of the training matrix is then updated (see equations \ref{eq1} and \ref{eq2}) in order to maximize the log joint likelihood (see equation \ref{eq3}).
+
+\begin{equation}
+u_i = \left( \lambda\sigma^2 I + \sum_{j \in \Omega_{w_i}} v_j v_j^T \right)^{-1}\left(\sum_{j \in \Omega_{w_i}} M_{ij} v_{j} \right)
+\label{eq1}
+\end{equation}
+
+\begin{equation}
+v_j = \left( \lambda\sigma^2 I + \sum_{i \in \Omega_{v_j}} u_i u_i^T \right)^{-1}\left(\sum_{j \in \Omega_{v_j}} M_{ij} u_{i} \right)
+\label{eq2}
+\end{equation}
+
+\begin{equation}
+\mathcal{L} = \sum_{(i,j) \in \Omega} \frac{1}{2\sigma^2} {|| M_{ij} - u_i^T v_j||}^2 - \sum_{i=1}^{N_1} \frac{\lambda}{2} ||u_i^2 || - \sum_{i=1}^{N_2} \frac{\lambda}{2} ||v_j^2 || + \text{constant}
+\label{eq3} 
+\end{equation}
+
+Note: We use a rank 20 matrix for factorization purposes, a $\lambda = 10$, and calculate the variance of the observations for our $\sigma^2$.
+Also, $\Omega$ is the set of all indices in the matrix which have an observation.
+
+We keep track of the root mean square error (RMSE) versus the testing set (i.e.
+how close our prediction is as compared to the actual normalized play count in the testing set) and the log joint likelihood of the training set as a function of iteration (see Figures \ref{fig:rmseplot} and \ref{fig:likeplot}).
+
+\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
+   \centering
+   \includegraphics[width=2in]{h845BuzFO9CNgAAAABJRU5ErkJggg==.png} 
+   \caption{(how close our prediction is as compared to the actual normalized play count in the testing set }
+   \label{fig:rmseplot}
+\end{figure}
+
+\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
+   \centering
+   \includegraphics[width=2in]{C9LAAAAABJRU5ErkJggg==.png} 
+   \caption{log joint likelihood of the training set as a function of iteration }
+   \label{fig:likeplot}
+\end{figure}
+
+\subsection{Kmeans Clustering}
+Next, we cluster each of the learned $u_i$ using k-means (describe more here later) so that the cluster centroids can be interpreted as listener/personality types.
+
+\section{Notes}
 
@@
 
-Sparsity of $8.08199453451\times 10^{-5}$ need updating
+Sparsity of $8.08199453451\times 10^{-5}$ (needs updating)
 
@@
 max = 35432 \\
+min = 1 \\
 std = 215.826789\\
@@
 
+29 - users
+34 - songs
+
 How to split into training and test? 
@@
 
+\section{Appendix}
+\href{https://github.com/eringrand/musicanalysis}{Link to our github repository} 
+\href{http://www.dtic.upf.edu/~ocelma/MusicRecommendationDataset/}{Link to the data set}
+
 \end{document}

### 2015-05-04 MSDgroupwriteup.tex
larger images

@@
 
-Using a coordinate ascent algorithm over 100 iterations, each row ($u_i$) and column ($v_j$) of the training matrix is then updated (see equations \ref{eq1} and \ref{eq2}) in order to maximize the log joint likelihood (see equation \ref{eq3}).
+Using a coordinate ascent algorithm over 100 iterations, each row ($u_i$) and column ($v_j$) of the training matrix is then updated (equations \ref{eq1} and \ref{eq2}) in order to maximize the log joint likelihood (equation \ref{eq3}).
 
@@
    \centering
-   \includegraphics[width=2in]{h845BuzFO9CNgAAAABJRU5ErkJggg==.png} 
-   \caption{(how close our prediction is as compared to the actual normalized play count in the testing set }
+   \includegraphics[width=4in]{h845BuzFO9CNgAAAABJRU5ErkJggg==.png} 
+   \caption{how close our prediction is as compared to the actual normalized play count in the testing set }
    \label{fig:rmseplot}
@@
    \centering
-   \includegraphics[width=2in]{C9LAAAAABJRU5ErkJggg==.png} 
+   \includegraphics[width=4in]{C9LAAAAABJRU5ErkJggg==.png} 
    \caption{log joint likelihood of the training set as a function of iteration }

### 2015-05-04 Justin_MF.py
Uploading MF code from ML

@@ -1,0 +1,38 @@
+import numpy as np
+
+RMSE = {}
+L={}
+iter = 100
+users = {}
+movies = {}
+for i in range(usernum):
+    users[i] = np.where(matrix[i] != 0)[0]
+
+for j in range(movienum):
+    movies[j] = np.where(matrix[:,j] != 0)[0]
+
+# Performing matrix factorization
+for c in range(iter):
+    for i in range(usernum):
+        inner = (v[:,users[i]] * v[:,users[i]].T)
+        outer = np.asmatrix(matrix[i][users[i]]) * v[:,users[i]].T
+        u[i] = ((rr + inner).I * outer.T).T
+
+    for j in range(movienum):
+        inner = (u[movies[j]].T * u[movies[j]])
+        outer = np.asmatrix(matrix[movies[j],j]) * u[movies[j]]
+        v[:,j] = ((rr + inner).I * outer.T)
+
+    predict = (u * v).round()
+    predict[predict>5.0] = 5.0
+    predict[predict<1.0] = 1.0
+
+    RMSE[c] = (np.mean((np.asarray(testmatrix[testmatrix!=0] - predict[testmatrix!=0]))**2))**.5
+
+    predict = (u * v)
+    first = np.sum((np.asarray(matrix[matrix!=0] - predict[matrix!=0]))**2)/(2*variance)
+    second = (np.linalg.norm(u)**2) * lmb / 2
+    third = (np.linalg.norm(v)**2) * lmb / 2
+    L[c] = - first - second - third
+
+    print c, RMSE[c], L[c]

### 2015-05-04 analysis_MF_new.py
MF

@@ -1,0 +1,165 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from operator import itemgetter
+
+# Create Mtrain and Mtest matrices here
+
+N1 = np.shape(Mtrain)[0]
+N2 = np.shape(Mtrain)[1]
+  
+
+omegau = {}
+omegav = {}
+
+for (i,j) in omega:
+    if i in omegau.keys():
+        omegau[i].append(j)
+    else:
+        omegau[i] = [j]
+    if j in omegav.keys():
+        omegav[j].append(i)
+    else:
+        omegav[j] = [i]
+
+# Calculate var          
+#var = ???????
+d = 20
+I = np.identity(d)
+lamb = 10
+t1 = lamb*var*I
+
+mean = np.zeros([d])
+v = np.empty([N2,d])
+u = np.empty([N1,d])
+
+for j in xrange(0, N2):
+    v[j] = np.random.multivariate_normal(mean,(1/float(lamb))*I)
+v = np.asmatrix(v)
+u = np.asmatrix(u)
+
+
+RMSE = {}
+L={}
+iter = 100
+users = {}
+movies = {}
+for i in range(N1):
+    users[i] = np.where(Mtrain[i] != 0)[0]
+
+for j in range(movienum):
+    movies[j] = np.where(Mtrain[:,j] != 0)[0]
+
+# Performing matrix factorization
+for c in range(iter):
+    for i in range(usernum):
+        inner = (v[:,users[i]] * v[:,users[i]].T)
+        outer = np.asmatrix(Mtrain[i][users[i]]) * v[:,users[i]].T
+        u[i] = ((t1 + inner).I * outer.T).T
+
+    for j in range(movienum):
+        inner = (u[movies[j]].T * u[movies[j]])
+        outer = np.asmatrix(Mtrain[movies[j],j]) * u[movies[j]]
+        v[:,j] = ((t1 + inner).I * outer.T)
+
+    predict = (u * v).round()
+    predict[predict>5.0] = 5.0
+    predict[predict<1.0] = 1.0
+
+    RMSE[c] = (np.mean((np.asarray(Mtest[Mtest!=0] - predict[Mtest!=0]))**2))**.5
+
+    predict = (u * v)
+    first = np.sum((np.asarray(Mtrain[Mtrain!=0] - predict[Mtrain!=0]))**2)/(2*var)
+    second = (np.linalg.norm(u)**2) * lamb / 2
[…]

### 2015-05-04 MF-current.py
needs to be optimized

@@ -1,0 +1,76 @@
+N1 = np.shape(Mtrain)[0]
+N2 = np.shape(Mtrain)[1]
+
+d = 20
+I = np.identity(d)
+lamb = 10
+t1 = lamb*var*I
+
+mean = np.zeros([d])
+v = np.empty([d,N2])
+u = np.empty([N1,d])
+
+for j in xrange(0, N2):
+    v[:,j] = np.random.multivariate_normal(mean,(1/float(lamb))*I)
+v = np.asmatrix(v)
+u = np.asmatrix(u)
+
+
+RMSE = []
+loglik=[]
+iter = 1
+users1 = {}
+songs1 = {}
+for i in range(N1):
+    users1[i] = np.where(Mtrain[i] != 0)[0]
+
+for j in range(N2):
+    songs1[j] = np.where(Mtrain[:,j] != 0)[0]
+
+
+def predict(u, v):
+    product = np.dot(u, v)
+    if product < 0:
+        pred = 0.0
+    elif product > 1:
+        pred = 1.0
+    else:
+        pred = product
+    return pred
+
+
+    # Performing matrix factorization
+for c in range(iter):
+    for i in range(N1):
+        inner = (v[:,users1[i]] * v[:,users1[i]].T)
+        outer = np.asmatrix(Mtrain[i][users1[i]]) * v[:,users1[i]].T
+        u[i] = ((t1 + inner).I * outer.T).T
+
+    for j in range(N2):
+        inner = (u[songs1[j]].T * u[songs1[j]])
+        outer = np.asmatrix(Mtrain[songs1[j],j]) * u[songs1[j]]
+        v[:,j] = ((t1 + inner).I * outer.T)
+
+    u = np.asarray(u)
+    v = np.asarray(v)
+    sum3 = 0        
+    for (i,j) in omega_test:
+        prediction = predict(u[i], v[:,j])
+        actual = Mtest[i][j]
+        sum3 = sum3 + (prediction - actual)**2
+    temp = (sum3/float(len(omega_test)))**0.5
+    RMSE.append(temp)
+    
+    sum4 = 0
+    for (i, j) in omega:
+        sum4 = sum4 + 0.5/var*np.power(Mtrain[i][j] - np.dot(u[i], v[:,j]),2)
+    sum4 = -sum4
+    sum5 = 0
+    for i in xrange(0, N1):
[…]

### 2015-05-04 MF-current.py
MF

@@ -12,8 +12,6 @@ u = np.empty([N1,d])
 
 for j in xrange(0, N2):
     v[:,j] = np.random.multivariate_normal(mean,(1/float(lamb))*I)
-v = np.asmatrix(v)
-u = np.asmatrix(u)
 
 
 RMSE = []
@@ -42,17 +40,15 @@ def predict(u, v):
     # Performing matrix factorization
 for c in range(iter):
     for i in range(N1):
-        inner = (v[:,users1[i]] * v[:,users1[i]].T)
-        outer = np.asmatrix(Mtrain[i][users1[i]]) * v[:,users1[i]].T
-        u[i] = ((t1 + inner).I * outer.T).T
+        inner = np.dot(v[:,users1[i]], v[:,users1[i]].T)
+        outer = np.dot(Mtrain[i][users1[i]], v[:,users1[i]].T)
+        u[i] = np.dot((t1 + inner).I * outer.T).T
 
     for j in range(N2):
-        inner = (u[songs1[j]].T * u[songs1[j]])
-        outer = np.asmatrix(Mtrain[songs1[j],j]) * u[songs1[j]]
-        v[:,j] = ((t1 + inner).I * outer.T)
+        inner = np.dot(u[songs1[j]].T * u[songs1[j]])
+        outer = np.dot(Mtrain[songs1[j],j] * u[songs1[j]])
+        v[:,j] = np.dot((t1 + inner).I, outer.T)
 
-    u = np.asarray(u)
-    v = np.asarray(v)
     sum3 = 0        
     for (i,j) in omega_test:
         prediction = predict(u[i], v[:,j])

### 2015-05-04 DataFrames.py
Adding stuff

@@ -40,17 +40,17 @@ testsub.ix[~testsub.index.isin(sample),'plays'] = 0
 trainpivot = trainsub.pivot(index='user_id',columns='song_id', values='plays')
 user_index = pivot.index
 song_index = pivot.columns
-M_train = pivot.as_matrix()
+M_train = trainpivot.as_matrix()
 M_train = np.nan_to_num(M_train)
 
 testpivot = testsub.pivot(index='user_id',columns='song_id', values='plays')
-M_test = pivot.as_matrix()
+M_test = testpivot.as_matrix()
 M_test = np.nan_to_num(M_test)
 
 
 
 # http://labrosa.ee.columbia.edu/millionsong/sites/default/files/AdditionalFiles/unique_tracks.txt
-unique_tracks = pd.read_csv("unique_tracks.txt",sep='<SEP>', header = None, names = ['track_id', 'song_id', 'arist_name', 'song_title'])
+unique_tracks = pd.read_csv("unique_tracks.txt",sep='<SEP>', header = None, names = ['tid', 'sid', 'arist_name', 'song_title'])
 
 # http://labrosa.ee.columbia.edu/millionsong/sites/default/files/AdditionalFiles/tracks_per_year.txt
 tracks_per_year = pd.read_csv("tracks_per_year.txt",sep='<SEP>', header = None, names =['year','track_id', 'song_title'])
@@ -129,7 +129,8 @@ plt.show()
 userplaycsum = userhist.sort('plays')
 userplaycsum['plays'] = userplaycsum['plays'].cumsum()
 userplaycsum['plays'] = userplaycsum['plays']/userplaycsum['plays'].max()
-userplaycsum.plot('user_id', 'plays')
+userplaycsum = userplaycsum.reset_index(drop=True).reset_index()
+userplaycsum.plot('index', 'plays')
 plt.show()
 
 # histogram of number of plays for songs
@@ -158,3 +159,20 @@ songplaycsum['plays'] = songplaycsum['plays'].cumsum()
 songplaycsum['plays'] = songplaycsum['plays']/songplaycsum['plays'].max()
 songplaycsum.plot('song_id', 'plays')
 plt.show()
+
+
+tags_100 = lastfm_tags[lastfm_tags['val']==100]
+
+centroid_songs = pd.read_csv("clustercenters.txt", sep=' ', header=None, names=['centroid','sid'])
+
[…]

### 2015-05-04 DataFrames.py
data frames

@@ -165,8 +165,9 @@ tags_100 = lastfm_tags[lastfm_tags['val']==100]
 
 centroid_songs = pd.read_csv("clustercenters.txt", sep=' ', header=None, names=['centroid','sid'])
 
-centroid_unique = pd.merge(unique_tracks, centroid_songs, on='sid')
-tag_centroid_unique = pd.merge(centroid_unique, tags_100, on='tid') 
+centroid_unique = pd.merge(centroid_songs, unique_tracks, on='sid')
+tags_100_list = tags_100.groupby('tid')['tag'].apply(', '.join).reset_index()
+tag_centroid_unique = pd.merge(centroid_unique, tags_100_list, on='tid', how='left') 
 
 #f = open("unique_tracks.txt")
 #tid = []

### 2015-05-04 DataFrames.py
Use artist terms

@@ -52,6 +52,9 @@ M_test = np.nan_to_num(M_test)
 # http://labrosa.ee.columbia.edu/millionsong/sites/default/files/AdditionalFiles/unique_tracks.txt
 unique_tracks = pd.read_csv("unique_tracks.txt",sep='<SEP>', header = None, names = ['tid', 'sid', 'arist_name', 'song_title'])
 
+# http://labrosa.ee.columbia.edu/millionsong/sites/default/files/AdditionalFiles/unique_artists.txt
+unique_artists = pd.read_csv("unique_artists.txt",sep='<SEP>',header = None, names = ['artist_id', 'artist_mbid', 'tid', 'artist_name'])
+
 # http://labrosa.ee.columbia.edu/millionsong/sites/default/files/AdditionalFiles/tracks_per_year.txt
 tracks_per_year = pd.read_csv("tracks_per_year.txt",sep='<SEP>', header = None, names =['year','track_id', 'song_title'])
 
@@ -168,6 +171,12 @@ centroid_songs = pd.read_csv("clustercenters.txt", sep=' ', header=None, names=[
 centroid_unique = pd.merge(centroid_songs, unique_tracks, on='sid')
 tags_100_list = tags_100.groupby('tid')['tag'].apply(', '.join).reset_index()
 tag_centroid_unique = pd.merge(centroid_unique, tags_100_list, on='tid', how='left') 
+
+artist_data = pd.merge(unique_artists, artist_term.groupby('artist_id')['term'].apply(', '.join).reset_index(), on='artist_id')
+
+centroid_artist = pd.merge(centroid_unique, artist_data, on='tid')
+term_centroid_artist = pd.merge(centroid_unique, centroid_artist, on='tid', how='left')
+
 
 #f = open("unique_tracks.txt")
 #tid = []

### 2015-05-04 MF-current.py
Update MF-current.py

@@ -70,3 +70,11 @@ for c in range(iter):
         sum6 = sum6 + 0.5*lamb*np.sum(v[:,j]**2)
     sum6 = -sum6
     loglik.append(sum4+sum5+sum6)
+    
+    
+    
+    
+    
+    
+    from sklearn.cluster import KMeans
+    kmeans_model = KMeans(n_clusters=10).fit_predict(u)

### 2015-05-04 MSDgroupwriteup.tex
new files

@@
 \usepackage{ amssymb }
+\newcommand{\TRANSPOSE}{\ensuremath{T}}
 
 \begin{document}
-\title{Clustering of Last.fm Music Artists}
+\title{Clustering of Last.fm Music}
 
@@
 There is subjectivity in picking $d$ but $20$ is a common place to start.
-In the factorized matrices, the predicted rating will be $\hat{M}_{ij} = u_i^T v_j$
+In the factorized matrices, the predicted rating will be $\hat{M}_{ij} = u_i^\TRANSPOSE  v_j$
 
@@
 \begin{equation}
-u_i = \left( \lambda\sigma^2 I + \sum_{j \in \Omega_{w_i}} v_j v_j^T \right)^{-1}\left(\sum_{j \in \Omega_{w_i}} M_{ij} v_{j} \right)
+u_i = \left( \lambda\sigma^2 I + \sum_{j \in \Omega_{u_i}} v_j v_j^\TRANSPOSE \right)^{-1}\left(\sum_{j \in \Omega_{u_i}} M_{ij} v_{j} \right)
 \label{eq1}
@@
 \begin{equation}
-v_j = \left( \lambda\sigma^2 I + \sum_{i \in \Omega_{v_j}} u_i u_i^T \right)^{-1}\left(\sum_{j \in \Omega_{v_j}} M_{ij} u_{i} \right)
+v_j = \left( \lambda\sigma^2 I + \sum_{i \in \Omega_{v_j}} u_i u_i^\TRANSPOSE  \right)^{-1}\left(\sum_{j \in \Omega_{v_j}} M_{ij} u_{i} \right)
 \label{eq2}
@@
 \begin{equation}
-\mathcal{L} = \sum_{(i,j) \in \Omega} \frac{1}{2\sigma^2} {|| M_{ij} - u_i^T v_j||}^2 - \sum_{i=1}^{N_1} \frac{\lambda}{2} ||u_i^2 || - \sum_{i=1}^{N_2} \frac{\lambda}{2} ||v_j^2 || + \text{constant}
+\mathcal{L} = \sum_{(i,j) \in \Omega} \frac{1}{2\sigma^2} {|| M_{ij} - u_i^\TRANSPOSE  v_j||}^2 - \sum_{i=1}^{N_1} \frac{\lambda}{2} ||u_i^2 || - \sum_{i=1}^{N_2} \frac{\lambda}{2} ||v_j^2 || + \text{constant}
 \label{eq3} 
@@
 
-Note: We use a rank 20 matrix for factorization purposes, a $\lambda = 10$, and calculate the variance of the observations for our $\sigma^2$.
+\emph{Note}: We use a rank 20 matrix for factorization purposes, a $\lambda = 10$, and calculate the variance of the observations for our $\sigma^2$.
 Also, $\Omega$ is the set of all indices in the matrix which have an observation.
@@
    \label{fig:likeplot}
+\end{figure}
+
+\begin{figure}[ht]
+\begin{minipage}[b]{0.45\linewidth}
+\centering
+\includegraphics[width=\textwidth]{h845BuzFO9CNgAAAABJRU5ErkJggg==.png}
+\caption{default}
+\label{fig:figure1}
+\end{minipage}
+\hspace{0.5cm}
+\begin{minipage}[b]{0.45\linewidth}
+\centering
+\includegraphics[width=\textwidth]{C9LAAAAABJRU5ErkJggg==.png}
+\caption{default}
+\label{fig:figure2}
+\end{minipage}
 \end{figure}

### 2015-05-04 whateveryoulike.py
to look at

@@ -1,0 +1,405 @@
+
+# coding: utf-8
+
+# In[1]:
+
+import numpy as np
+from collections import defaultdict
+from scipy import sparse
+import matplotlib.pyplot as plt
+from operator import itemgetter
+
+f = open('kaggle_visible_evaluation_triplets.txt', 'r')
+data = f.read().split('\n')
+
+
+# In[2]:
+
+users = set()
+songs = set()
+numplays = {}
+numsongplayed = defaultdict(list)
+songdic = defaultdict(list)
+userdic = defaultdict(list)
+
+for line in data:
+    split = line.strip().split('\t')
+    if len(split) != 3:
+        print len(split), split
+    else:
+        userhash = split[0]
+        songhash = split[1]
+        plays = float(split[2])
+        users.add(userhash)
+        songs.add(songhash)
+        numplays[songhash, userhash] = plays
+        if songhash not in userdic[userhash]:
+            userdic[userhash].append(plays)
+        if userhash not in songdic[songhash]:
+            songdic[songhash].append(userhash)
+        #if songhash not in numsongplayed:
+        numsongplayed[songhash].append(plays)
+
+
+# In[3]:
+
+# subset by number of songs a user has rated
+p1 = { key: value for key, value in userdic.items() if sum(value) > 29 }
+# subset by number of plays
+p2 = { key: value for key, value in numsongplayed.items() if sum(value) > 34 }
+
+# list of songs
+subsongs = p2.keys()
+
+# list of users who have rated those songs
+subusers = set()
+for s in subsongs:
+    for u in songdic[s]:
+        if u in p1:
+            subusers.add(u)
+        
+p3 = { key: value for key, value in numplays.items() if key[0] in p2 and key[1] in p1 }
+
+print len(subusers), len(users)
+print len(subsongs), len(songs)
+print 100. - (float(len(p3))/len(data) * 100)
+#print (float(len(subsongs)) * len(subusers)) / ( len(songs) * len(users)) * 100
+
+
+# In[13]:
+
+for key, value in p3.items():
+    song = key[0]
+    user = key[1]
+    playcounts = p1[user]
[…]

### 2015-05-04 whateveryoulike.py
update of Erin's code

@@ -8,6 +8,7 @@ from collections import defaultdict
 from scipy import sparse
 import matplotlib.pyplot as plt
 from operator import itemgetter
+from sklearn.cluster import KMeans
 
 f = open('kaggle_visible_evaluation_triplets.txt', 'r')
 data = f.read().split('\n')
@@ -66,7 +67,7 @@ print 100. - (float(len(p3))/len(data) * 100)
 #print (float(len(subsongs)) * len(subusers)) / ( len(songs) * len(users)) * 100
 
 
-# In[13]:
+# In[4]:
 
 for key, value in p3.items():
     song = key[0]
@@ -80,7 +81,7 @@ for key, value in p3.items():
     p3[song, user] = normplay
 
 
-# In[4]:
+# In[5]:
 
 allkeys = p3.keys()
 
@@ -92,7 +93,7 @@ trainingKEYS, testKEYS = allkeys[:trainingN], allkeys[trainingN:]
 print len(trainingKEYS), len(testKEYS)
 
 
-# In[5]:
+# In[6]:
 
 n = len(subusers)
 m = len(subsongs)
@@ -151,7 +152,7 @@ for key in testKEYS:
     Mtest[userindex, songindex] = float(plays)
 
 
-# In[6]:
+# In[7]:
 
 var = np.var(varcalc)
 print var
@@ -190,7 +191,7 @@ RMSE = []
 loglik=[]
 
 
-# In[10]:
+# In[15]:
 
 def predict(u, v):
     product = np.dot(u, v)
@@ -242,11 +243,6 @@ for c in range(iter):
     loglik.append(sum4+sum5+sum6)
 
 
-# In[134]:
-
-#print RMSE, loglik
-
-
 # In[13]:
 
 get_ipython().magic(u'matplotlib inline')
@@ -264,11 +260,6 @@ plt.subplot(1,2,2)
 plt.plot(loglik)
 plt.xlabel('Iteration')
 plt.ylabel('Log Joint Likelihood')
-
-
-# In[14]:
-
-RMSE
 
 
 # In[18]:
@@ -291,18 +282,12 @@ RMSE
 
 def kmeans(data, obs, clusters):    
     lik = []
-    
     c_old = np.zeros([obs]) - 1
-    
     mu = np.empty([clusters, np.shape(data)[1]])
-    
     temp = np.random.random_integers(0, np.shape(data)[0]-1, clusters)
-    
     for k in xrange(0, clusters):
[…]

### 2015-05-05 NMF.py
NMF

@@ -1,0 +1,45 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from sklearn.preprocessing import normalize
+
+
+# X should be songs by users (i.e. N2 x N1)
+X = M.T
+
+rank = 25
+
+W = np.random.uniform(size=[N2,rank])
+H = np.random.uniform(size=[rank,N1])
+
+
+divobj = []
+for t in xrange(0,200):
+    print t
+    temp1 = normalize(np.transpose(W), norm='l1', axis=1)
+    purple = X / (np.dot(W,H) + 10**(-16))
+    H = H * np.dot(temp1,purple)
+    temp2 = normalize(np.transpose(H), norm='l1', axis=0) 
+    purple = X / (np.dot(W,H) + 10**(-16))
+    W = W * np.dot(purple,temp2)
+    temp = np.sum((X * np.log(1/(np.dot(W,H) + 10**(-16)))) + (np.dot(W,H)))
+    divobj.append(temp)
+  
+
+# Plot objective function
+plt.plot(divobj)
+plt.xlabel('Iteration')
+plt.ylabel('Objective function')
+plt.title('NMF with Divergence Penalty')
+
+
+# Top 10 words
+W_norm = normalize(W, norm='l1', axis=0)
+
+
+for col in range(rank):
+    temp = np.argsort(W_norm[:,col])[::-1][0:10]
+    print 'Column of W: ' + str(col)
+    for j in temp:
+        #index of song is j
+        print index_to_song_dict[j] + ' ' + str(W_norm[:,col][j])
+    print '\n'

### 2015-05-05 whateveryoulike.py
Adding mAP calculation

@@ -150,6 +150,12 @@ for key in testKEYS:
     songindex = song_dict[song]
     omega_test.append((userindex,songindex))
     Mtest[userindex, songindex] = float(plays)
+    i = userindex
+    j = songindex
+    if i in omegau_test:
+        omegau_test[i].append(j)
+    else:
+        omegau_test[i] = [j]
 
 
 # In[7]:
@@ -242,6 +248,52 @@ for c in range(iter):
     sum6 = -sum6
     loglik.append(sum4+sum5+sum6)
 
+
+def apk(actual, predicted, k=500):
+    """
+    Computes the average precision at k.
+
+    This function computes the average prescision at k between two lists of
+    items.
+
+    Parameters
+    ----------
+    actual : list
+             A list of elements that are to be predicted (order doesn't matter)
+    predicted : list
+                A list of predicted elements (order does matter)
+    k : int, optional
+        The maximum number of predicted elements
+
+    Returns
+    -------
+    score : double
+            The average precision at k over the input lists
+
+    """
+    if len(predicted)>k:
+        predicted = predicted[:k]
+
+    score = 0.0
+    num_hits = 0.0
+
+    for i,p in enumerate(predicted):
+        if p in actual and p not in predicted[:i]:
+            num_hits += 1.0
+            score += num_hits / (i+1.0)
+
+    if not actual:
+        return 1.0
+
+    return score / min(len(actual), k)
+
+predict_m = np.dot(u,v)
+apk_sum = 0
+
+for i in range(len(u)):
+    apk_sum += apk(omegau_test[i],np.argsort(M_train[i])[::1][:500])
+
+map = apk_sum/len(u)
 
 # In[13]:
 

### 2015-05-05 whateveryoulike.py
Editing incorrect matrix name

@@ -291,7 +291,7 @@ predict_m = np.dot(u,v)
 apk_sum = 0
 
 for i in range(len(u)):
-    apk_sum += apk(omegau_test[i],np.argsort(M_train[i])[::1][:500])
+    apk_sum += apk(omegau_test[i],np.argsort(predict_m[i])[::1][:500])
 
 map = apk_sum/len(u)
 

### 2015-05-05 UserItemBasedSimilarity.py
item based rec

@@ -1,0 +1,198 @@
+from operator import itemgetter
+import numpy as np
+from collections import defaultdict
+
+
+
+f = open('kaggle_visible_evaluation_triplets.txt', 'r')
+data = f.read().split('\n')
+
+
+# In[2]:
+
+users = set()
+songs = set()
+numplays = {}
+numsongplayed = defaultdict(list)
+songdic = defaultdict(list)
+userdic = defaultdict(list)
+
+for line in data:
+    split = line.strip().split('\t')
+    if len(split) != 3:
+        print len(split), split
+    else:
+        userhash = split[0]
+        songhash = split[1]
+        plays = float(split[2])
+        users.add(userhash)
+        songs.add(songhash)
+        numplays[songhash, userhash] = plays
+        if songhash not in userdic[userhash]:
+            userdic[userhash].append(plays)
+        if userhash not in songdic[songhash]:
+            songdic[songhash].append(userhash)
+        #if songhash not in numsongplayed:
+        numsongplayed[songhash].append(plays)
+
+
+# In[3]:
+
+# subset by number of songs a user has rated
+p1 = { key: value for key, value in userdic.items() if sum(value) > 98 }
+# subset by number of plays
+p2 = { key: value for key, value in numsongplayed.items() if sum(value) > 52 }
+
+# list of songs
+subsongs = p2.keys()
+
+# list of users who have rated those songs
+subusers = set()
+for s in subsongs:
+    for u in songdic[s]:
+        if u in p1:
+            subusers.add(u)
+        
+p3 = { key: value for key, value in numplays.items() if key[0] in p2 and key[1] in p1 }
+
+print len(subusers), len(users)
+print len(subsongs), len(songs)
+print 100. - (float(len(p3))/len(data) * 100)
+#print (float(len(subsongs)) * len(subusers)) / ( len(songs) * len(users)) * 100
+
+
+# In[4]:
+
+for key, value in p3.items():
+    song = key[0]
+    user = key[1]
+    playcounts = p1[user]
+    maxs = np.max(playcounts)
+    rng = maxs - 0.0
+    high = 1.0
+    low = 0.0 
+    normplay = high - (((high - low) * (maxs - value)) / rng)
+    p3[song, user] = normplay
+
+
+# In[5]:
+
+allkeys = p3.keys()
+
+np.random.shuffle(allkeys)
+
[…]

### 2015-05-05 cleanedcodeall.py
cleaned code

@@ -1,0 +1,436 @@
+
+# coding: utf-8
+
+# In[18]:
+
+def norm(data_dict, norm_dict, typeofnorm, high=1.0, low=0.0):
+# Normilization of User Playcount
+    if typeofnorm == "user":
+        for key, value in data_dict.items():
+            song = key[0]
+            user = key[1]
+            playcounts = norm_dict[user]
+            maxs = np.max(playcounts)
+            rng = maxs - low
+            normplay = high - (((high - low) * (maxs - value)) / rng)
+            data_dict[song, user] = normplay
+    if typeofnorm == "song":
+        for key, value in data_dict.items():
+            song = key[0]
+            user = key[1]
+            playcounts = norm_dict[song]
+            maxs = np.max(playcounts)
+            rng = maxs - low
+            normplay = high - (((high - low) * (maxs - value)) / rng)
+            data_dict[song, user] = normplay
+    if typeofnorm == "none":
+        return data_dict
+    if typeofnorm == "binary":
+        for key, value in data_dict.items():
+            song = key[0]
+            user = key[1]
+            data_dict[song, user] = 1.0
+    return data_dict
+
+
+# In[11]:
+
+def apk(actual, predicted, k=500):
+    """
+    Computes the average precision at k.
+
+    This function computes the average prescision at k between two lists of
+    items.
+
+    Parameters
+    ----------
+    actual : list
+             A list of elements that are to be predicted (order doesn't matter)
+    predicted : list
+                A list of predicted elements (order does matter)
+    k : int, optional
+        The maximum number of predicted elements
+
+    Returns
+    -------
+    score : double
+            The average precision at k over the input lists
+
+    """
+    if len(predicted)>k:
+        predicted = predicted[:k]
+
+    score = 0.0
+    num_hits = 0.0
+
+    for i,p in enumerate(predicted):
+        if p in actual and p not in predicted[:i]:
+            num_hits += 1.0
+            score += num_hits / (i+1.0)
+
+    if not actual:
+        return 1.0
+
+    return score / min(len(actual), k)
+
+
+# In[12]:
+
[…]

### 2015-05-05 cleanedcodeall.py
small changes

@@ -1,7 +1,7 @@
 
 # coding: utf-8
 
-# In[18]:
+# In[32]:
 
 def norm(data_dict, norm_dict, typeofnorm, high=1.0, low=0.0):
 # Normilization of User Playcount
@@ -33,7 +33,7 @@ def norm(data_dict, norm_dict, typeofnorm, high=1.0, low=0.0):
     return data_dict
 
 
-# In[11]:
+# In[33]:
 
 def apk(actual, predicted, k=500):
     """
@@ -74,7 +74,7 @@ def apk(actual, predicted, k=500):
     return score / min(len(actual), k)
 
 
-# In[12]:
+# In[34]:
 
 def MF(Mtrain, Mtest, var, omega, omega_test, omegau, omegau_test, d=20, iter=15):
     
@@ -171,7 +171,7 @@ def MF(Mtrain, Mtest, var, omega, omega_test, omegau, omegau_test, d=20, iter=15
     return map
 
 
-# In[13]:
+# In[35]:
 
 def rawkmeans(Mtrain, omegau, ncl=10):
     kmeans_setup = KMeans(n_clusters=ncl)
@@ -249,7 +249,7 @@ def rawkmeans(Mtrain, omegau, ncl=10):
     return map
 
 
-# In[23]:
+# In[36]:
 
 def fromthetop(data, typeofnorm, alg, d=20, iter=15, ncl=10):
     users = set()
@@ -378,15 +378,14 @@ def fromthetop(data, typeofnorm, alg, d=20, iter=15, ncl=10):
         else: 
             var = np.var(varcalc)
 
-        
     if alg == "MF":
         print "Running MF Algorithm with d ="+str(d)+" with "+str(iter)+" iteration."
-        map = MF(Mtrain, Mtest, var, omega, omega_test, omegau, omegau_test)
+        mapval = MF(Mtrain, Mtest, var, omega, omega_test, omegau, omegau_test) 
             
     if alg == "Kmeans":
-        map = rawkmeans(Mtrain, omegau, ncl)
+        mapval = rawkmeans(Mtrain, omegau, ncl)
         
-    return map
+    return mapval
 
 
 # In[6]:
@@ -403,22 +402,22 @@ f = open('kaggle_visible_evaluation_triplets.txt', 'r')
 data = f.read().split('\n')
 
 
-# In[ ]:
+# In[37]:
 
[…]

### 2015-05-06 NMF.py
Update NMF.py

@@ -43,3 +43,14 @@ for col in range(rank):
         #index of song is j
         print index_to_song_dict[j] + ' ' + str(W_norm[:,col][j])
     print '\n'
+    
+    
+    
+    
+# Loop over all users the below code 
+rec_user = {}
+for user in users:
+    H_column = H[:,user]
+    col = np.argmax(H_column)
+    W_column = W[:,col]
+    rec_user[user] = np.argsort(W_column)[::-1][0:500]

### 2015-05-06 tfidf_MF.py
Uploading tfidf_MF

@@ -1,0 +1,242 @@
+import numpy as np
+#import sqlite3 as sqlite
+import pandas as pd
+#import pandas.io.sql as psql
+import matplotlib.pyplot as plt
+import random
+
+import matplotlib
+matplotlib.style.use('ggplot')
+
+def apk(actual, predicted, k=500):
+    """
+    Computes the average precision at k.
+
+    This function computes the average prescision at k between two lists of
+    items.
+
+    Parameters
+    ----------
+    actual : list
+             A list of elements that are to be predicted (order doesn't matter)
+    predicted : list
+                A list of predicted elements (order does matter)
+    k : int, optional
+        The maximum number of predicted elements
+
+    Returns
+    -------
+    score : double
+            The average precision at k over the input lists
+
+    """
+    if len(predicted)>k:
+        predicted = predicted[:k]
+
+    score = 0.0
+    num_hits = 0.0
+
+    for i,p in enumerate(predicted):
+        if p in actual and p not in predicted[:i]:
+            num_hits += 1.0
+            score += num_hits / (i+1.0)
+
+    if not actual:
+        return 1.0
+
+    return score / min(len(actual), k)
+
+f = open("kaggle_visible_evaluation_triplets.txt", 'rb')
+eval = pd.read_csv(f,sep='\t',header = None, names = ['user_id','sid','plays'])
+
+userhist = eval.groupby('user_id').count()
+userhist = pd.DataFrame(userhist).reset_index()
+usersub = userhist[userhist['plays']>27]
+
+songhist = eval.groupby('sid').count()
+songhist = pd.DataFrame(songhist).reset_index()
+songsub = songhist[songhist['plays']>22]
+
+sub = eval[eval['sid'].isin(songsub['sid'])]
+sub = sub[sub['user_id'].isin(usersub['user_id'])]
+
+len(set(sub['user_id']))
+len(set(sub['sid']))
+
+sub.shape[0]/float(eval.shape[0])
+
+sub.shape[0]/float(len(set(sub['user_id']))*len(set(sub['sid'])))
+
+#sub_max = pd.DataFrame(sub.groupby('sid').max()).reset_index()
+#merged = pd.merge(sub,sub_max,on="sid")
[…]

### 2015-05-06 whateveryoulike.py
Update whateveryoulike.py

@@ -443,3 +443,4 @@ np.savetxt("UintoKmeans", u)
 
 
 
+[x for x in recommend[user1] if x[0] not in user_to_song[user1]][0:500]

### 2015-05-06 artist_based.py
Adding artist_based algorithm recommender

@@ -1,0 +1,154 @@
+import numpy as np
+import sqlite3 as sqlite
+import pandas as pd
+import pandas.io.sql as psql
+import matplotlib.pyplot as plt
+import random
+
+from collections import OrderedDict
+
+import matplotlib
+matplotlib.style.use('ggplot')
+
+def apk(actual, predicted, k=500):
+    """
+    Computes the average precision at k.
+
+    This function computes the average prescision at k between two lists of
+    items.
+
+    Parameters
+    ----------
+    actual : list
+             A list of elements that are to be predicted (order doesn't matter)
+    predicted : list
+                A list of predicted elements (order does matter)
+    k : int, optional
+        The maximum number of predicted elements
+
+    Returns
+    -------
+    score : double
+            The average precision at k over the input lists
+
+    """
+    if len(predicted)>k:
+        predicted = predicted[:k]
+
+    score = 0.0
+    num_hits = 0.0
+
+    for i,p in enumerate(predicted):
+        if p in actual and p not in predicted[:i]:
+            num_hits += 1.0
+            score += num_hits / (i+1.0)
+
+    if not actual:
+        return 1.0
+
+    return score / min(len(actual), k)
+
+# load observation data
+f = open("kaggle_visible_evaluation_triplets.txt", 'rb')
+eval = pd.read_csv(f,sep='\t',header = None, names = ['user_id','sid','plays'])
+
+# count number of songs per user and subset
+userhist = eval.groupby('user_id').count()
+userhist = pd.DataFrame(userhist).reset_index()
+usersub = userhist[userhist['plays']>27]
+
+# count number of users per song and subset
+songhist = eval.groupby('sid').count()
+songhist = pd.DataFrame(songhist).reset_index()
+songsub = songhist[songhist['plays']>22]
+
+# subset the whole dataset
+sub = eval[eval['sid'].isin(songsub['sid'])]
+sub = sub[sub['user_id'].isin(usersub['user_id'])]
+
+# sampling/splitting the dataset
+sample = random.sample(sub.index, int(sub.shape[0]*0.2))
+trainsub = sub.copy()
+trainsub.ix[trainsub.index.isin(sample),'plays'] = 0
+testsub = sub.copy()
+testsub.ix[~trainsub.index.isin(sample),'plays'] = 0
+
+# generating a matrix out of dataframe
+trainpivot = trainsub.pivot(index='user_id',columns='sid', values='plays')
+# creating the mapping of user index to user id
[…]

### 2015-05-06 writeup/MSDgroupwriteup.tex

--- 2015-05-04 MSDgroupwriteup.tex
@@
 \section{Data Munging}
+
+\section{Algorithms}
 
@@
 
-\begin{figure}[ht]
-\begin{minipage}[b]{0.45\linewidth}
-\centering
-\includegraphics[width=\textwidth]{h845BuzFO9CNgAAAABJRU5ErkJggg==.png}
-\caption{default}
-\label{fig:figure1}
-\end{minipage}
-\hspace{0.5cm}
-\begin{minipage}[b]{0.45\linewidth}
-\centering
-\includegraphics[width=\textwidth]{C9LAAAAABJRU5ErkJggg==.png}
-\caption{default}
-\label{fig:figure2}
-\end{minipage}
-\end{figure}
+%
+%\begin{figure}[ht]
+%\begin{minipage}[b]{0.45\linewidth}
+%\centering
+%\includegraphics[width=\textwidth]{h845BuzFO9CNgAAAABJRU5ErkJggg==.png}
+%\caption{default}
+%\label{fig:figure1}
+%\end{minipage}
+%\hspace{0.5cm}
+%\begin{minipage}[b]{0.45\linewidth}
+%\centering
+%\includegraphics[width=\textwidth]{C9LAAAAABJRU5ErkJggg==.png}
+%\caption{default}
+%\label{fig:figure2}
+%\end{minipage}
+%\end{figure}
 
-\subsection{Kmeans Clustering}
+\subsection{Naked Kmeans }
 Next, we cluster each of the learned $u_i$ using k-means (describe more here later) so that the cluster centroids can be interpreted as listener/personality types.
+
+\subsection{Non-Negative Matrix Factorization}
+
+\subsection{User based and Item based}
+
+\subsection{Using Tags}
+
+\section{Results}
 

### 2015-05-06 writeup/MSDgroupwriteup.tex
correct MAP values

@@
 \section{Data Munging}
+\subsection{Data}
+\subsubsection{Data Statistics}
+\subsubsection{Problems and Solutions}
+\subsection{Normalization}
 
@@
 
-\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
-   \centering
-   \includegraphics[width=4in]{h845BuzFO9CNgAAAABJRU5ErkJggg==.png} 
-   \caption{how close our prediction is as compared to the actual normalized play count in the testing set }
-   \label{fig:rmseplot}
-\end{figure}
-
-\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
-   \centering
-   \includegraphics[width=4in]{C9LAAAAABJRU5ErkJggg==.png} 
-   \caption{log joint likelihood of the training set as a function of iteration }
-   \label{fig:likeplot}
-\end{figure}
+%\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
+%   \centering
+%   \includegraphics[width=4in]{h845BuzFO9CNgAAAABJRU5ErkJggg==.png} 
+%   \caption{how close our prediction is as compared to the actual normalized play count in the testing set }
+%   \label{fig:rmseplot}
+%\end{figure}
+%
+%\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
+%   \centering
+%   \includegraphics[width=4in]{C9LAAAAABJRU5ErkJggg==.png} 
+%   \caption{log joint likelihood of the training set as a function of iteration }
+%   \label{fig:likeplot}
+%\end{figure}
 
@@
 
-\subsection{User based and Item based}
+\subsection{User Based and Item Based Recommendation}
 
-\subsection{Using Tags}
+\subsection{Artist Based Recommendation}
 
 \section{Results}
+\subsection{Matrix Factorization}
+
+\subsection{Naked Kmeans }
+Next, we cluster each of the learned $u_i$ using k-means (describe more here later) so that the cluster centroids can be interpreted as listener/personality types.
+
+\subsection{Non-Negative Matrix Factorization}
+
+\subsection{User Based and Item Based Recommendation}
+
+\subsection{Artist Based Recommendation}
 

### 2015-05-06 popularity_baseline.py
Adding algorithm to calculate popularity baseline

@@ -1,0 +1,118 @@
+import numpy as np
+import pandas as pd
+import random
+
+def apk(actual, predicted, k=500):
+    """
+    Computes the average precision at k.
+
+    This function computes the average prescision at k between two lists of
+    items.
+
+    Parameters
+    ----------
+    actual : list
+             A list of elements that are to be predicted (order doesn't matter)
+    predicted : list
+                A list of predicted elements (order does matter)
+    k : int, optional
+        The maximum number of predicted elements
+
+    Returns
+    -------
+    score : double
+            The average precision at k over the input lists
+
+    """
+    if len(predicted)>k:
+        predicted = predicted[:k]
+
+    score = 0.0
+    num_hits = 0.0
+
+    for i,p in enumerate(predicted):
+        if p in actual and p not in predicted[:i]:
+            num_hits += 1.0
+            score += num_hits / (i+1.0)
+
+    if not actual:
+        return 1.0
+
+    return score / min(len(actual), k)
+
+# load observation data
+f = open("kaggle_visible_evaluation_triplets.txt", 'rb')
+eval = pd.read_csv(f,sep='\t',header = None, names = ['user_id','sid','plays'])
+
+# count number of songs per user and subset
+userhist = eval.groupby('user_id').count()
+userhist = pd.DataFrame(userhist).reset_index()
+usersub = userhist[userhist['plays']>27]
+
+# count number of users per song and subset
+songhist = eval.groupby('sid').count()
+songhist = pd.DataFrame(songhist).reset_index()
+songsub = songhist[songhist['plays']>22]
+
+# subset the whole dataset
+sub = eval[eval['sid'].isin(songsub['sid'])]
+sub = sub[sub['user_id'].isin(usersub['user_id'])]
+
+# sampling/splitting the dataset
+sample = random.sample(sub.index, int(sub.shape[0]*0.2))
+trainsub = sub.copy()
+trainsub.ix[trainsub.index.isin(sample),'plays'] = 0
+testsub = sub.copy()
+testsub.ix[~trainsub.index.isin(sample),'plays'] = 0
+
+# generating a matrix out of dataframe
+trainpivot = trainsub.pivot(index='user_id',columns='sid', values='plays')
+# creating the mapping of user index to user id
+user_index = pd.DataFrame(trainpivot.index).reset_index()
+user_index.columns = [['user_index','user_id']]
[…]

### 2015-05-06 Justin_MF.py
Updating list of references

@@ -1,38 +1,0 @@
-import numpy as np
-
-RMSE = {}
-L={}
-iter = 100
-users = {}
-movies = {}
-for i in range(usernum):
-    users[i] = np.where(matrix[i] != 0)[0]
-
-for j in range(movienum):
-    movies[j] = np.where(matrix[:,j] != 0)[0]
-
-# Performing matrix factorization
-for c in range(iter):
-    for i in range(usernum):
-        inner = (v[:,users[i]] * v[:,users[i]].T)
-        outer = np.asmatrix(matrix[i][users[i]]) * v[:,users[i]].T
-        u[i] = ((rr + inner).I * outer.T).T
-
-    for j in range(movienum):
-        inner = (u[movies[j]].T * u[movies[j]])
-        outer = np.asmatrix(matrix[movies[j],j]) * u[movies[j]]
-        v[:,j] = ((rr + inner).I * outer.T)
-
-    predict = (u * v).round()
-    predict[predict>5.0] = 5.0
-    predict[predict<1.0] = 1.0
-
-    RMSE[c] = (np.mean((np.asarray(testmatrix[testmatrix!=0] - predict[testmatrix!=0]))**2))**.5
-
-    predict = (u * v)
-    first = np.sum((np.asarray(matrix[matrix!=0] - predict[matrix!=0]))**2)/(2*variance)
-    second = (np.linalg.norm(u)**2) * lmb / 2
-    third = (np.linalg.norm(v)**2) * lmb / 2
-    L[c] = - first - second - third
-
-    print c, RMSE[c], L[c]

### 2015-05-06 MFfunction.py
MF function

@@ -1,0 +1,101 @@
+def MF(Mtrain, Mtest, var, omega, omega_test, omegau, omegau_test, d=20, iter=15):
+    
+    N1 = np.shape(Mtrain)[0]
+    N2 = np.shape(Mtrain)[1]
+
+    users1 = {}
+    songs1 = {}
+    for i in range(N1):
+        users1[i] = np.where(Mtrain[i] != 0)[0]
+
+    for j in range(N2):
+        songs1[j] = np.where(Mtrain[:,j] != 0)[0]
+
+    I = np.identity(d)
+    lamb = 10
+    t1 = lamb*var*I
+
+    mean = np.zeros([d])
+    v = np.empty([d,N2])
+    u = np.empty([N1,d])
+
+    for j in xrange(0, N2):
+        v[:,j] = np.random.multivariate_normal(mean,(1/float(lamb))*I)
+
+    RMSE = []
+    loglik=[]
+
+    def predict(u, v):
+        product = np.dot(u, v)
+        if product < 0.:
+            pred = 0.0
+        elif product > 1.:
+            pred = 1.0
+        else:
+            pred = product
+        return pred
+
+    # Performing matrix factorization
+    for c in xrange(0,iter):
+        for i in range(N1):
+            inner = np.dot(v[:,users1[i]], v[:,users1[i]].T)
+            outer = np.dot(Mtrain[i][users1[i]], v[:,users1[i]].T)
+            u[i] = np.dot((np.linalg.inv(t1 + inner)), outer.T).T
+
+        for j in range(N2):
+            inner = np.dot(u[songs1[j]].T, u[songs1[j]])
+            outer = np.dot(Mtrain[songs1[j],j], u[songs1[j]])
+            v[:,j] = np.dot(np.linalg.inv(t1 + inner), outer.T)
+
+        sum3 = 0        
+        for (i,j) in omega_test:
+            prediction = predict(u[i], v[:,j])
+            actual = Mtest[i][j]
+            sum3 = sum3 + (prediction - actual)**2
+        temp = (sum3/float(len(omega_test)))**0.5
+        RMSE.append(temp)
+
+        sum4 = 0
+        for (i, j) in omega:
+            sum4 = sum4 + 0.5/var*np.power(Mtrain[i][j] - np.dot(u[i], v[:,j]),2)
+        sum4 = -sum4
+        sum5 = 0
+        for i in xrange(0, N1):
[…]

### 2015-05-06 MFfunction.py
Editing map calculation

@@ -89,12 +89,11 @@ def MF(Mtrain, Mtest, var, omega, omega_test, omegau, omegau_test, d=20, iter=15
 
     recommend = {}
     rec = {}
-    for i in omegau_test:
-        if i in omegau:
-            recommend[i] = np.argsort(predict_m[i])[::-1]
-            rec[i] = [x for x in recommend[i] if x not in omegau[i]][0:500]
-            apk_sum += apk(omegau_test[i], rec[i])
-            counter += 1
+    for i in list(set(omegau_test).intersection(omegau)):
+        recommend[i] = np.argsort(predict_m[i])[::-1]
+        rec[i] = [x for x in recommend[i] if x not in omegau[i]][0:500]
+        apk_sum += apk(omegau_test[i], rec[i])
+        counter += 1
 
     mapval = apk_sum/counter
 

### 2015-05-06 writeup/MSDgroupwriteup.tex
MF function
added ref notes

@@
 
+To Reference a paper: 
+The Million Song Dataset Challenge: \cite{McFee:2012:MSD:2187980.2188222} \\
+Million Song Dataset Recommendation Project Report: \cite{li2012million} \\
+codebook-based scalable music tagging with poisson matrix factorization: \cite{liangcodebook} \\
+Matrix Factorization Techniques for Recommender Systems: \cite{koren2009matrix} \\
+A Preliminary Study on a Recommender System for the Million Songs Dataset Challenge: \cite{aiolli2013preliminary} \\
+
+\bibliography{msdref}
+
 \section{Appendix}
 \href{https://github.com/eringrand/musicanalysis}{Link to our github repository}
-\href{http://www.dtic.upf.edu/~ocelma/MusicRecommendationDataset/}{Link to the data set}
 

### 2015-05-06 algorithms/matrix_factorization.py
Moving "finished" algorithms to folder

@@ -1,0 +1,241 @@
+import numpy as np
+#import sqlite3 as sqlite
+import pandas as pd
+#import pandas.io.sql as psql
+import matplotlib.pyplot as plt
+import random
+
+import matplotlib
+matplotlib.style.use('ggplot')
+
+def apk(actual, predicted, k=500):
+    """
+    Computes the average precision at k.
+
+    This function computes the average prescision at k between two lists of
+    items.
+
+    Parameters
+    ----------
+    actual : list
+             A list of elements that are to be predicted (order doesn't matter)
+    predicted : list
+                A list of predicted elements (order does matter)
+    k : int, optional
+        The maximum number of predicted elements
+
+    Returns
+    -------
+    score : double
+            The average precision at k over the input lists
+
+    """
+    if len(predicted)>k:
+        predicted = predicted[:k]
+
+    score = 0.0
+    num_hits = 0.0
+
+    for i,p in enumerate(predicted):
+        if p in actual and p not in predicted[:i]:
+            num_hits += 1.0
+            score += num_hits / (i+1.0)
+
+    if not actual:
+        return 1.0
+
+    return score / min(len(actual), k)
+
+def MF(Mtrain, Mtest, var, omega, omega_test, omegau, omegau_test, d=20, iter=15):
+    
+    N1 = np.shape(Mtrain)[0]
+    N2 = np.shape(Mtrain)[1]
+
+    users1 = {}
+    songs1 = {}
+    for i in range(N1):
+        users1[i] = np.where(Mtrain[i] != 0)[0]
+
+    for j in range(N2):
+        songs1[j] = np.where(Mtrain[:,j] != 0)[0]
+
+    I = np.identity(d)
+    lamb = 10
+    t1 = lamb*var*I
+
+    mean = np.zeros([d])
+    v = np.empty([d,N2])
+    u = np.empty([N1,d])
+
+    for j in xrange(0, N2):
+        v[:,j] = np.random.multivariate_normal(mean,(1/float(lamb))*I)
+
+    RMSE = []
+    loglik=[]
+
+    def predict(u, v):
+        product = np.dot(u, v)
+        if product < 0.:
+            pred = 0.0
+        elif product > 1.:
+            pred = 1.0
+        else:
+            pred = product
+        return pred
+
[…]

### 2015-05-06 DataFrames.py
Cleaning DataFrames.py

@@ -5,30 +5,64 @@ import pandas.io.sql as psql
 import matplotlib.pyplot as plt
 import random
 
+import matplotlib
+matplotlib.style.use('ggplot')
+
+
+# 34 user plays, 29 song plays
+# 98 user plays, 52 song plays
+# 27 user songcounts, 22 song usercounts
+
+
 # http://www.kaggle.com/c/msdchallenge/data
-eval = pd.read_csv("kaggle_visible_evaluation_triplets.txt",sep='\t',header = None, names = ['user_id','song_id','plays'])
+f = open("kaggle_visible_evaluation_triplets.txt", 'rb')
+eval = pd.read_csv(f,sep='\t',header = None, names = ['user_id','sid','plays'])
 
-userhist = eval.groupby('user_id').sum()
+#userhist = eval.groupby('user_id').sum()
+#userhist = pd.DataFrame(userhist).reset_index()
+#usersub = userhist[userhist['plays']>98]
+
+#songhist = eval.groupby('sid').sum()
+#songhist = pd.DataFrame(songhist).reset_index()
+#songsub = songhist[songhist['plays']>52]
+
+#sub = eval[eval['sid'].isin(songsub['sid'])]
+#sub = sub[sub['user_id'].isin(usersub['user_id'])]
+
+#len(set(sub['user_id']))
+#len(set(sub['sid']))
+
+#sub.shape[0]/float(eval.shape[0])
+
+#sub.shape[0]/float(len(set(sub['user_id']))*len(set(sub['sid'])))
+
+# for number of songs instead of plays
+
+userhist = eval.groupby('user_id').count()
 userhist = pd.DataFrame(userhist).reset_index()
-usersub = userhist[userhist['plays']>29]
+usersub = userhist[userhist['plays']>27]
 
-songhist = eval.groupby('song_id').sum()
+songhist = eval.groupby('sid').count()
 songhist = pd.DataFrame(songhist).reset_index()
-songsub = songhist[songhist['plays']>34]
+songsub = songhist[songhist['plays']>22]
 
-sub = eval[eval['song_id'].isin(songsub['song_id'])]
+sub = eval[eval['sid'].isin(songsub['sid'])]
 sub = sub[sub['user_id'].isin(usersub['user_id'])]
 
 len(set(sub['user_id']))
-len(set(sub['song_id']))
+len(set(sub['sid']))
 
 sub.shape[0]/float(eval.shape[0])
 
-sub_max = pd.DataFrame(sub.groupby('song_id').max()).reset_index()
[…]

### 2015-05-06 writeup/MSDgroupwriteup.tex
updates

--- 2015-05-06 writeup/MSDgroupwriteup.tex
@@
 \newcommand{\degree}{\ensuremath{^\circ}}
-\citestyle{apj}
-\bibliographystyle{aa}
+\bibliographystyle{apalike}
 \usepackage[maxfloats=25]{morefloats}
@@
 
+To Reference a paper:
+
+The Million Song Dataset Challenge: \citep{McFee:2012:MSD:2187980.2188222} \\
+Million Song Dataset Recommendation Project Report: \citep{li2012million} \\
+codebook-based scalable music tagging with poisson matrix factorization: \citep{liangcodebook} \\
+Matrix Factorization Techniques for Recommender Systems: \citep{koren2009matrix} \\
+A Preliminary Study on a Recommender System for the Million Songs Dataset Challenge: \citep{aiolli2013preliminary} \\
+
+\bibliography{msdref}
+
 \section{Appendix}
 \href{https://github.com/eringrand/musicanalysis}{Link to our github repository}
-\href{http://www.dtic.upf.edu/~ocelma/MusicRecommendationDataset/}{Link to the data set}
 

### 2015-05-06 algorithms/UserAndSongBasedRec.py
user/item based rec

@@ -1,0 +1,248 @@
+from operator import itemgetter
+import numpy as np
+from collections import defaultdict
+import pandas as pd
+import random
+
+
+
+
+
+
+f = open("kaggle_visible_evaluation_triplets.txt", 'rb')
+eval = pd.read_csv(f,sep='\t',header = None, names = ['user_id','sid','plays'])
+
+userhist = eval.groupby('user_id').count()
+userhist = pd.DataFrame(userhist).reset_index()
+usersub = userhist[userhist['plays']>27]
+
+songhist = eval.groupby('sid').count()
+songhist = pd.DataFrame(songhist).reset_index()
+songsub = songhist[songhist['plays']>22]
+
+sub = eval[eval['sid'].isin(songsub['sid'])]
+sub = sub[sub['user_id'].isin(usersub['user_id'])]
+
+len(set(sub['user_id']))
+len(set(sub['sid']))
+
+sub.shape[0]/float(eval.shape[0])
+
+sub.shape[0]/float(len(set(sub['user_id']))*len(set(sub['sid'])))
+
+#sub_max = pd.DataFrame(sub.groupby('sid').max()).reset_index()
+#merged = pd.merge(sub,sub_max,on="sid")
+#merged['plays_x'] = merged['plays_x']/merged['plays_y']
+#sub_norm = merged[['user_id_x', 'sid', 'plays_x']]
+#sub_norm.columns = ['user_id', 'sid', 'plays']
+
+sample = random.sample(sub.index, int(sub.shape[0]*0.2))
+
+trainsub = sub.copy()
+trainsub.ix[trainsub.index.isin(sample),'plays'] = 0
+
+testsub = sub.copy()
+testsub.ix[~trainsub.index.isin(sample),'plays'] = 0
+
+trainpivot = trainsub.pivot(index='user_id',columns='sid', values='plays')
+
+user_index = pd.DataFrame(trainpivot.index).reset_index()
+user_index.columns = [['user_index','user_id']]
+trainsub = pd.merge(trainsub, user_index, on='user_id')
+testsub = pd.merge(testsub, user_index, on='user_id')
+song_index = pd.DataFrame(trainpivot.columns).reset_index()
+song_index.columns = [['song_index','sid']]
+trainsub = pd.merge(trainsub, song_index, on='sid')
+testsub = pd.merge(testsub, song_index, on='sid')
+
+M_train = trainpivot.as_matrix()
[…]

### 2015-05-07 writeup/MSDgroupwriteup.tex
edits to intro

@@
-% \documentclass[12pt]{emulateapj}
 \documentclass[12pt,preprint]{aastex}
+ %\documentclass[12pt]{emulateapj}
 \usepackage[margin= 1.0in]{geometry}    % See geometry.pdf to learn the layout options. There are lots.
@@
 \usepackage{enumitem}
-\usepackage{natbib,twoopt}
+\usepackage{natbib}
 \newcommand{\degree}{\ensuremath{^\circ}}
-\bibliographystyle{apalike}
 \usepackage[maxfloats=25]{morefloats}
@@
 \newcommand{\TRANSPOSE}{\ensuremath{T}}
+
+\bibliographystyle{apalike}
 
@@
 \section{Introduction}
+In the online world, recommendation systems are used in shopping, entertainment and even dating. 
+Because of their popularity and usefulness, there are lots of incentives to implement a good recommendation system, but a lack of large datasets to learn the best algorithms from. 
+In the music industry, companies use recommendation systems to provide a better music streaming radio service to their listeners. 
+With the music industry shifting away from a more traditional distribution model of physical CDs or records, 
+having an efficient and well performing recommendation system is very important.
 
-\section{Data Munging}
+In building a recommendation system, there are many problems to tackle. This is way some methods that take more information about the user and the song into account.
+There is no set similarity measure between two songs, as evident by the fact that songs by the same artist are often very different. As such, there are many different types of algorithms to explore.
+
+For our project, we chose to explore various song recommendation systems algorithms such as matrix factorization, non-negative matrix factorization as well as user based and item based similarity measures.
+% memory based collaborative filtering
+
 \subsection{Data}
+The main source of data for this project was the Million Song Data Set (MSD)  \citep{Bertin-Mahieux2011}. We used the sample data set from the Kaggle competition as well as metadata from MSD.
+
+The full data set had  $163206$ songs  $110000$ users, and the number of play counts 
+The scale of the data was too big for our computers to explore, so we subset the data down to a size easily handled by an average laptop.
+
+We choose to subset the data to popular songs and the most active users. We chose a subset of 10\% the size of the original matrix.   % Because?
+
+len(value) > 27 users
+len(value) > 22 songs
+
 \subsubsection{Data Statistics}
-\subsubsection{Problems and Solutions}
-\subsection{Normalization}
+
+- Sparsity 
+-
+
+SONG PLAY COUNT  \\
+mean = 28 \\
+max = 35432 \\
+min = 1 \\
+std = 215.826789\\
+
+USER PLAY COUNT\\
+mean = 42\\
+max = 1305\\
+min = 5\\
+std = 53.31547\\
+
+\subsection{Data Analysis}
+We did a random selection by data element to split between test and train.
+
+\subsection{Testing}
+We tested our algorithms against each other using a mean average precision 
+score for each method.  The mean average precision (MAP) algorithm came directly from the Kaggle MSD challenge \citep{kaggleMSD}. The score was used to rank the competition entries.
+
+The MAP scores were computed by calculating the precision between a list of recommendations for a user and that user's test set.
+The MAP score is then the average precision value over all the users in the test set.
+
+The bench mark we used to recommend to the top 500 songs to every user.
 
@@
 The goal of matrix factorization is to use collaborative methods to build a recommendation system for users based on user ratings of objects (in this case songs).
-This will allow us to recommend certain songs to users based on their listening history and without the need for using content based approaches.
+This allowed us to recommend certain songs to users based on their listening history and without the need for using content based approaches.
 Since our data set contains number of plays for a given user and song (rather than rating), we normalized plays for every song on a scale between 0 and 1 and used this as a proxy for rating.
 
-% Erin's part here on normalization.
-
 We then constructed training and testing matrices, of which both have $N_1$ users (rows denoted by $u_i$) and $N_2$ songs (columns denoted by $v_j$).
-Of course the matrices will be very sparse, containing zeros in all entries except for those in which a user (rows of matrix) has listened to a song (columns of matrix).
+Of course the matrices were very sparse, containing zeros in all entries except for those in which a user (rows of matrix) has listened to a song (columns of matrix).
 The goal is to factor the training matrix into the product of two matrixes, $U$ and $V$.
@@
 
-\emph{Note}: We use a rank 20 matrix for factorization purposes, a $\lambda = 10$, and calculate the variance of the observations for our $\sigma^2$.
-Also, $\Omega$ is the set of all indices in the matrix which have an observation.
+\emph{Note: We use a rank 20 matrix for factorization purposes, a $\lambda = 10$, and calculate the variance of the observations for our $\sigma^2$.
+Also, $\Omega$ is the set of all indices in the matrix which have an observation}.
 
@@
 
-\subsection{Naked Kmeans }
-Next, we cluster each of the learned $u_i$ using k-means (describe more here later) so that the cluster centroids can be interpreted as listener/personality types.
-
-\subsection{Non-Negative Matrix Factorization}
+%\subsection{Naked Kmeans }
+%\subsection{Non-Negative Matrix Factorization}
 
 \subsection{User Based and Item Based Recommendation}
+Intro to Memory-Based Collaborative Filtering
 
-\subsection{Artist Based Recommendation}
+A simpler set of methods for determining which songs to recommend to individual users is by using user- or item/song- based similarity (both also considered collaborative filtering).
+These only consider whether or not a user listened to a song and does not take into account the play count.
+Similarity scores between users (i.e.
+user-based) and songs (i.e.
+item-based) were calculated using cosine similarity.
+Intuitively, they work by recommending users songs that other similar users like (user-based) or by recommending songs which are similar to the songs the user has already listened to (item-based).
 
-\section{Results}
-\subsection{Matrix Factorization}
+User-Based Collaborative Filtering
 
-\subsection{Naked Kmeans }
-Next, we cluster each of the learned $u_i$ using k-means (describe more here later) so that the cluster centroids can be interpreted as listener/personality types.
+For user-based recommendation, we first calculated the similarity score between every pair of users, u and v, using the equation below:
 
-\subsection{Non-Negative Matrix Factorization}
+Inline image 1
+\begin{equation}
+sim(u,v) = \frac{\# common items}{{\# items(u)}^{1/2} {\# items (v)}^{1/2}}
+\end{equation}
 
-\subsection{User Based and Item Based Recommendation}
+Then, for each user, we looked at each song in the dataset and found all other users who have listened to that song.
+Next, we added up the similarity score of each of those users, v, with the original user, u, and got the weight for a particular song, i:
 
-\subsection{Artist Based Recommendation}
+$w_i = \sum_{v \epsilon V} sim(u, v)$
+
+That sum is a proxy for how likely the user is to like that particular song.
+We went through that process for all songs and then recommended the user the top 500 songs which had the highest scores.
+This method led to a MAP value of 0.0383.
+
+Item-Based Collaborative Filtering
+
+For item-based recommendation, which turned out to be our best algorithm in terms of a MAP value, we first calculated the similarity score between every pair of songs, i and j, using the equation below and saved the most similar songs to each song in the process:
+
+Inline image 2
+
+Then, for each user, we found all the songs listened to.
+Next, we got the most similar songs to each of those songs listened to by the user.
+If the same similar songs came up multiple times, we added the weights together.
+In other words, for each song, b, that was found to be similar to one of the songs a that the user listened to, we calculated the score for that song, b:
+
+$w_b = \sum_{a \epsilon A} sim(a, b)$
+
+Lastly, we sorted the songs by scores and recommended the user the top 500 songs which had the highest scores. This method led to a MAP value of 0.0503.
+
+%\subsection{Artist Based Recommendation}
+
+\section{Conclusion}
 
 \section{Notes}
-
-\href{http://www.dtic.upf.edu/~ocelma/MusicRecommendationDataset/}{Music Recommendation Dataset}
-
 Sparsity of $8.08199453451\times 10^{-5}$ (needs updating)
 
-The matrix was too big for our computers to handle the matrix factorization using the whole set of users and songs, so we subset the data.
-
-The full matrix was originally $163206$ songs x $110000$ users.
-
-Subset songs by how many times the song was listened to. i.e the popularity of the song
-= new number
-
-Subset users by how many songs they've listened to (how many popular song?)
-= new number
-
-SONG PLAY COUNT  \\
-mean = 28 \\
-max = 35432 \\
-min = 1 \\
-std = 215.826789\\
-
-USER PLAY COUNT\\
-mean = 42\\
-max = 1305\\
-min = 5\\
-std = 53.31547\\
+%The matrix was too big for our computers to handle the matrix factorization using the whole set of users and songs, so we subset the data. 
+%Subset songs by how many times the song was listened to. i.e the popularity of the song
+%= new number 
+%Subset users by how many songs they've listened to (how many popular song?)
+%= new number
 
@@
 34 - songs 
-
-How to split into training and test? 
-Random selection by user or element? ELEMENT. 
-Need matrixes to be the same size to compare them.
+60 \%
 
@@
 
+\small{
 \section{Appendix}
-\href{https://github.com/eringrand/musicanalysis}{Link to our github repository}
+
+\begin{itemize}
+
+  \item Link to our github repository: \url{https://github.com/eringrand/musicanalysis}
+  \item The MSD Kaggle challenge: \url{https://www.kaggle.com/c/msdchallenge}
+
+\end{itemize}
+}
 

### 2015-05-07 writeup/MSDgroupwriteup.tex
put in user

@@
 
-\subsection{User Based and Item Based Recommendation}
-Intro to Memory-Based Collaborative Filtering
+\subsection{User Based and Item Based Collaborative Filtering}
+%\subsubsection{Intro to Memory-Based Collaborative Filtering}
 
@@
 
-User-Based Collaborative Filtering
+\subsubsection{User-Based Collaborative Filtering}
 
-For user-based recommendation, we first calculated the similarity score between every pair of users, u and v, using the equation below:
+For user-based recommendation, we first calculated the similarity score between every pair of users, $u$ and $v$, using the equation below:
 
-Inline image 1
 \begin{equation}
-sim(u,v) = \frac{\# common items}{{\# items(u)}^{1/2} {\# items (v)}^{1/2}}
+sim(u,v) = \frac{\text{\# common items}(u, v)}{{\text{\# items}(u)}^{1/2} \times {\text{\# items}(v)}^{1/2}}
 \end{equation}
@@
 Then, for each user, we looked at each song in the dataset and found all other users who have listened to that song.
-Next, we added up the similarity score of each of those users, v, with the original user, u, and got the weight for a particular song, i:
+Next, we added up the similarity score of each of those users, v, with the original user, u, and got the weight for a particular song, $i$:
 
-$w_i = \sum_{v \epsilon V} sim(u, v)$
+$$w_i = \sum_{v \epsilon V} sim(u, v)$$
 
-That sum is a proxy for how likely the user is to like that particular song.
-We went through that process for all songs and then recommended the user the top 500 songs which had the highest scores.
+The sum is a proxy for how likely the user is to like that particular song.
+We went through that process for all songs and recommended the user the top 500 songs which had the highest scores.
 This method led to a MAP value of 0.0383.
 
-Item-Based Collaborative Filtering
+\subsubsection{Item-Based Collaborative Filtering}
 
-For item-based recommendation, which turned out to be our best algorithm in terms of a MAP value, we first calculated the similarity score between every pair of songs, i and j, using the equation below and saved the most similar songs to each song in the process:
+For item-based recommendation, which turned out to be our best algorithm in terms of a MAP value, we first calculated the similarity score between every pair of songs, $i$ and $j$, using the equation below and saved the most similar songs to each song in the process:
 
-Inline image 2
+\begin{equation}
+sim(i,j) = \frac{\text{\# common users}(i, j)}{{\text{\# items}(i)}^{1/2} \times {\text{\# items}(j)}^{1/2}}
+\end{equation}
 
@@
 If the same similar songs came up multiple times, we added the weights together.
-In other words, for each song, b, that was found to be similar to one of the songs a that the user listened to, we calculated the score for that song, b:
+In other words, for each song, b, that was found to be similar to one of the songs a that the user listened to, we calculated the score for that song, $b$:
 
-$w_b = \sum_{a \epsilon A} sim(a, b)$
+$$w_b = \sum_{a \epsilon A} sim(a, b)$$
 

### 2015-05-07 writeup/MSDgroupwriteup.tex
Update MSDgroupwriteup.tex

@@
 In the online world, recommendation systems are used in shopping, entertainment and even dating. 
-Because of their popularity and usefulness, there are lots of incentives to implement a good recommendation system, but a lack of large datasets to learn the best algorithms from. 
+Because of their popularity and usefulness, there are numerous incentives to implement a good recommendation system. 
 In the music industry, companies use recommendation systems to provide a better music streaming radio service to their listeners. 
@@
 
-In building a recommendation system, there are many problems to tackle. This is way some methods that take more information about the user and the song into account.
+In building a recommendation system, there are many problems to tackle. This is why some methods take more information about the user and the song into account.
 There is no set similarity measure between two songs, as evident by the fact that songs by the same artist are often very different. As such, there are many different types of algorithms to explore.
@@
 \subsection{Data}
-The main source of data for this project was the Million Song Data Set (MSD)  \citep{Bertin-Mahieux2011}. We used the sample data set from the Kaggle competition as well as metadata from MSD.
+The main source of data for this project was the Million Song Data Set (MSD)  \citep{Bertin-Mahieux2011}.
+We used the sample data set from the Kaggle competition which contains tuples of user id, song id, and play count.
 
-The full data set had  $163206$ songs  $110000$ users, and the number of play counts 
-The scale of the data was too big for our computers to explore, so we subset the data down to a size easily handled by an average laptop.
+The sample data set had  $163206$ songs  $110000$ users, and the number of play counts 
+The scale of the data was too large to explore on our computers, so we subset the data down to a size easily handled by an average laptop.
 
-We choose to subset the data to popular songs and the most active users. We chose a subset of 10\% the size of the original matrix.   % Because?
-
-len(value) > 27 users
-len(value) > 22 songs
+We choose to subset the data to users who have listened to more than 27 songs and songs which were listened to by more than 22 users.
+This resulted in a datset which was 10\% the size of the original.
+Despite subsetting the data, we are still dealking with a matrix of $8130$ users by $11861$, so our algorithms will take a fair amount of time to run.
 
@@
 
-- Sparsity 
+- Sparsity .0016086
 -
@@
 \subsection{Data Analysis}
-We did a random selection by data element to split between test and train.
+In order to perform the analysis, we had to fill in a matrix which contained play counts for user and song pairs.
+We then did a random selection by data element to split between the training (80\%) and testing (20\%) sets.
 

### 2015-05-07 writeup/MSDgroupwriteup.tex
Update MSDgroupwriteup.tex

@@
 This resulted in a datset which was 10\% the size of the original.
-Despite subsetting the data, we are still dealking with a matrix of $8130$ users by $11861$, so our algorithms will take a fair amount of time to run.
+Despite subsetting the data, we are still dealing with a matrix of $8130$ users by $11861$ songs, so our algorithms will take a fair amount of time to run.
 

### 2015-05-07 writeup/MSDgroupwriteup.tex

@@
 
-The sample data set had  $163206$ songs  $110000$ users, and the number of play counts 
+The sample data set had $110000$ users, $163206$ songs, and the number of play counts for each user and song pair.
 The scale of the data was too large to explore on our computers, so we subset the data down to a size easily handled by an average laptop.
@@
 \subsection{Data Analysis}
-In order to perform the analysis, we had to fill in a matrix which contained play counts for user and song pairs.
-We then did a random selection by data element to split between the training (80\%) and testing (20\%) sets.
+In order to perform the analysis, we had to fill in training and testing matrices which contained play counts for user and song pairs.
+We did a random selection of data elements to split between the training (80\%) and testing (20\%) sets.
 
@@
 
-The bench mark we used to recommend to the top 500 songs to every user.
+The benchmark we used to compare our MAP scores against was recommending the top 500 most popular songs to every user.
+This naive method resulted in a MAP score of 0.013 (top 500 songs based on counts) or 0.012 (top 500 songs based on plays).
 

### 2015-05-07 algorithms/artist_based.py
Artist-based results and fixed algorithm

@@ -142,13 +142,15 @@ popularsonglist = list(eval_songmeta.song_index.values)
 # apk calculation
 apk_sum = 0
 
-for key_user_index in list(set(omegau_test).intersection(user_songs.user_index.values)):
+for key_user_index in list(set(omegau_test).intersection(omegau_train)):
     songlist = []
     for key_artist_id in list(user_songs[user_songs.user_index==key_user_index].artist_id.values):
         songlist += artist_songdict[key_artist_id]
     # appending the popularsonglist to the generated songlist, assume 600 is more than enough to account for any overlaps
     songlist = list(OrderedDict.fromkeys(songlist + popularsonglist))[:600]
     # starting from np.delete, this function is just to remove the songs in songlist that already exist in omegau_train
-    apk_sum = apk(omegau_test[key_user_index],np.delete(songlist,np.nonzero(np.in1d(songlist,omegau_train[key_user_index])))[:500])
+    apk_sum += apk(omegau_test[key_user_index],np.delete(songlist,np.nonzero(np.in1d(songlist,omegau_train[key_user_index])))[:500])
 
-map = apk_sum/len(set(omegau_test).intersection(user_songs.user_index.values))
+map = apk_sum/len(list(set(omegau_test).intersection(omegau_train)))
+
+

### 2015-05-07 writeup/MSDgroupwriteup.tex
Update MSDgroupwriteup.tex

@@
 
-The benchmark we used to compare our MAP scores against was recommending the top 500 most popular songs to every user.
-This naive method resulted in a MAP score of 0.013 (top 500 songs based on counts) or 0.012 (top 500 songs based on plays).
+We used two different baseline benchmarks to compare our MAP scores against.
+THe first was to simply recommend the top 500 most popular songs to every user.
+This naive method resulted in a MAP score of 0.013 (top 500 songs based on number of users who listened to the song) or 0.012 (top 500 songs based on number of plays the song had).
+The second method was to first recommend a user the most popular songs of artists the user already had played (but different songs from that artist).
+If this did not result in enough recommendations, we would then recommend the overall most popular songs as well at the end(similar to baseline method 1).
+Although this method is conservative in that it does not go beyond the tastes the user has already listened to, it resulted in a significant improvement in the MAP score to 0.045.
 
@@
 We went through that process for all songs and recommended the user the top 500 songs which had the highest scores.
-This method led to a MAP value of 0.0383.
+This method led to a MAP value of 0.0377.
 
@@
 
-Lastly, we sorted the songs by scores and recommended the user the top 500 songs which had the highest scores. This method led to a MAP value of 0.0503.
+Lastly, we sorted the songs by scores and recommended the user the top 500 songs which had the highest scores. This method led to a MAP value of 0.0479.
 

### 2015-05-07 writeup/MSDgroupwriteup.tex
Update MSDgroupwriteup.tex

@@
 THe first was to simply recommend the top 500 most popular songs to every user.
-This naive method resulted in a MAP score of 0.013 (top 500 songs based on number of users who listened to the song) or 0.012 (top 500 songs based on number of plays the song had).
+This naive method resulted in a MAP score of 0.0138 (top 500 songs based on number of users who listened to the song) or 0.0126 (top 500 songs based on number of plays the song had).
 The second method was to first recommend a user the most popular songs of artists the user already had played (but different songs from that artist).

### 2015-05-07 writeup/MSDgroupwriteup.tex
Update MSDgroupwriteup.tex

@@
 We tested our algorithms against each other using a mean average precision 
-score for each method.  The mean average precision (MAP) algorithm came directly from the Kaggle MSD challenge \citep{kaggleMSD}. The score was used to rank the competition entries.
+score for each method.  The mean average precision (MAP) calculation came directly from the Kaggle MSD challenge \citep{kaggleMSD}. The score was used to rank the competition entries.
 
-The MAP scores were computed by calculating the precision between a list of recommendations for a user and that user's test set.
-The MAP score is then the average precision value over all the users in the test set.
+The MAP@500 scores were computed by calculating the precision between a list of recommendations for a user and that user's test set.
+Specifically, we calculate the average precision at each element in the list of recommended items (i.e.
+percentage of correct items among first k recommendations) and sum these up for the first 500 recommendations.
+We then averaged the score over all users.
+Intuitively, the score looks at the percentage of recommendations that were correct (i.e.
+in the test set) but also takes order into account.
+So it is preferable to recommend the songs in the user's test set earlier in the list of 500.
 
 We used two different baseline benchmarks to compare our MAP scores against.
-THe first was to simply recommend the top 500 most popular songs to every user.
+The first was to simply recommend the top 500 most popular songs to every user.
 This naive method resulted in a MAP score of 0.0138 (top 500 songs based on number of users who listened to the song) or 0.0126 (top 500 songs based on number of plays the song had).

### 2015-05-07 writeup/MSDgroupwriteup.tex
Update MSDgroupwriteup.tex

@@
 Specifically, we calculate the average precision at each element in the list of recommended items (i.e.
-percentage of correct items among first k recommendations) and sum these up for the first 500 recommendations.
+percentage of correct items among first k recommendations) and sum these up for the first 500 recommendations (i.e.
+try for each $k \epsilon {1, ..., 500}$).
 We then averaged the score over all users.
-Intuitively, the score looks at the percentage of recommendations that were correct (i.e.
-in the test set) but also takes order into account.
+Intuitively, the score looks at the percentage of recommendations that were in-line with the test set of songs actually listened to but also takes order into account.
 So it is preferable to recommend the songs in the user's test set earlier in the list of 500.
+Basically, the MAP score is similar to AUC of the ROC curve discussed in class except it takes ordering of the recommendations into account.
 
@@
 This naive method resulted in a MAP score of 0.0138 (top 500 songs based on number of users who listened to the song) or 0.0126 (top 500 songs based on number of plays the song had).
-The second method was to first recommend a user the most popular songs of artists the user already had played (but different songs from that artist).
-If this did not result in enough recommendations, we would then recommend the overall most popular songs as well at the end(similar to baseline method 1).
-Although this method is conservative in that it does not go beyond the tastes the user has already listened to, it resulted in a significant improvement in the MAP score to 0.045.
+The second method was to first recommend a user the most popular songs of artists the user already had already played (but different songs from that artist).
+If this did not result in enough recommendations, we would then recommend the overall most popular songs as well at the end (similar to baseline method 1).
+Although this method is conservative in that it does not go beyond the tastes the user already has in the training set, it resulted in a significant improvement in the MAP score to 0.0448.
 

### 2015-05-07 writeup/MSDgroupwriteup.tex
Update MSDgroupwriteup.tex

@@
 We did a random selection of data elements to split between the training (80\%) and testing (20\%) sets.
+Next, we implemented various algorithms on the dataset in order to recommend songs to users, including matrix factorization, user-based collaborative filtering, and item-based collaborative filtering.
 
@@
 
-The MAP@500 scores were computed by calculating the precision between a list of recommendations for a user and that user's test set.
-Specifically, we calculate the average precision at each element in the list of recommended items (i.e.
-percentage of correct items among first k recommendations) and sum these up for the first 500 recommendations (i.e.
-try for each $k \epsilon {1, ..., 500}$).
+Specifically, we used MAP@500 scores which take into account the first 500 recommendations given to each user.
+Generally, the score looks at the precision between a list of recommendations for a user and that user's test set.
+This was done by calculating the average precision at each element in the list of recommended items (i.e.
+percentage of correct items among first k recommendations) and averaging these for the first 500 recommendations (i.e.
+for each $k \epsilon {1, ..., 500}$, calculate the precision and average them together).
 We then averaged the score over all users.
@@
 So it is preferable to recommend the songs in the user's test set earlier in the list of 500.
-Basically, the MAP score is similar to AUC of the ROC curve discussed in class except it takes ordering of the recommendations into account.
+The MAP score is similar to the AUC of the ROC curve discussed in class except it takes ordering of the recommendations into account, which is an important factor to consider in a recommendation system.
 

### 2015-05-07 writeup/MSDgroupwriteup.tex
Table
Update MSDgroupwriteup.tex

@@
 \begin{abstract}
-This is the abstract. 
+We attempt to build a recommendation system for a subset of the Million Song Dataset.
+We explore various algorithms including matrix factorization, user-based collaborative filtering, and item-based collaborative filtering.
+In each case we recommend the top 500 songs that our algorithm returns to each user and compare against the testing set of what the user actually listened to.
+Matrix factorization gives the worst results, only slightly above the popularity baseline, while artist-based popularity and user-based and item-based collaborative filtering methods yield better results.
 \end{abstract}
@@
 \section{Introduction}
-In the online world, recommendation systems are used in shopping, entertainment and even dating. 
+In the online world, recommendation systems are becoming increasingly important in various industries including in retail, entertainment, and even dating. 
 Because of their popularity and usefulness, there are numerous incentives to implement a good recommendation system. 
-In the music industry, companies use recommendation systems to provide a better music streaming radio service to their listeners. 
+In the music industry, companies use recommendation systems to provide a better service to their listeners. 
 With the music industry shifting away from a more traditional distribution model of physical CDs or records, 
@@
 
-In building a recommendation system, there are many problems to tackle. This is why some methods take more information about the user and the song into account.
+In building a recommendation system, we had to take into account various trade-offs in the type of models we chose to fit.
+For example, some methods take more information about the user and the song into account (e.g.
+metadata) while others use a collaborative filtering approach without taking into account user or song characteristic such as age, genre, location, etc.
 There is no set similarity measure between two songs, as evident by the fact that songs by the same artist are often very different. As such, there are many different types of algorithms to explore.
 
-For our project, we chose to explore various song recommendation systems algorithms such as matrix factorization, non-negative matrix factorization as well as user based and item based similarity measures.
+For our project, we explored various song recommendation systems algorithms such as popularity recommendations, artist based recommensations, matrix factorization, and user and item based similarity measures.
 % memory based collaborative filtering
@@
 %\subsubsection{Intro to Memory-Based Collaborative Filtering}
-
 A simpler set of methods for determining which songs to recommend to individual users is by using user- or item/song- based similarity (both also considered collaborative filtering).
@@
 \subsubsection{User-Based Collaborative Filtering}
-
 For user-based recommendation, we first calculated the similarity score between every pair of users, $u$ and $v$, using the equation below:
@@
 \subsubsection{Item-Based Collaborative Filtering}
-
 For item-based recommendation, which turned out to be our best algorithm in terms of a MAP value, we first calculated the similarity score between every pair of songs, $i$ and $j$, using the equation below and saved the most similar songs to each song in the process:
@@
 \section{Conclusion}
+After comparing the various algorithms, the simpler models gave us the best results.
+Our most complicated model, matrix factorization, gave us results even below the popularity baseline.
+On the other hand simply recommending users songs by artists they have already listened to (the artist based baseline) resulted in a significantly better recommendation system.
+Lastly, our song-based collaborative filtering technique produced the best results despite being a relatively simple algorithm as compared to matrix factorization.
+Our final results for each technique employed are summarized in the table below.
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{lc}
+
+\hline
+\bf{Algorithm} &  \bf{MAP score}\\ \hline
+Artist Based  Baseline  & 0.0448    \\ 
+%\hline
+Top 500 Songs by Count  Baseline &  0.0138  \\ 
+%\hline
+Top 500 Songs by Plays  Baseline &  0.0126  \\ 
+%\hline
+User Based  CF  &  0.0377 \\ 
+%\hline
+Item Based  CF  &  0.0479 \\ 
+%\hline
+Matrix Factorization  &   0.0143  \\ 
+%\hline
+\end{tabular}
+\end{center}
+\end{table}
 

### 2015-05-07 writeup/MSDgroupwriteup.tex
Update MSDgroupwriteup.tex

@@
 This resulted in a datset which was 10\% the size of the original.
-Despite subsetting the data, we are still dealing with a matrix of $8130$ users by $11861$ songs, so our algorithms will take a fair amount of time to run.
+Despite subsetting the data, we are still dealing with a matrix of $8130$ users by $11861$ songs, so some of our algorithms will take a fair amount of time to run.
 
@@
 After comparing the various algorithms, the simpler models gave us the best results.
-Our most complicated model, matrix factorization, gave us results even below the popularity baseline.
-On the other hand simply recommending users songs by artists they have already listened to (the artist based baseline) resulted in a significantly better recommendation system.
-Lastly, our song-based collaborative filtering technique produced the best results despite being a relatively simple algorithm as compared to matrix factorization.
+Our most complicated model, matrix factorization, gave us results barely above the popularity baseline.
+On the other hand, simply recommending users songs by artists they have already listened to (the artist-based baseline) resulted in a significantly better recommendation system.
+Our song-based collaborative filtering technique produced the best results despite being a relatively simple algorithm as compared to matrix factorization.
 Our final results for each technique employed are summarized in the table below.

### 2015-05-07 writeup/MSDgroupwriteup.tex

@@
 In building a recommendation system, we had to take into account various trade-offs in the type of models we chose to fit.
-For example, some methods take more information about the user and the song into account (e.g.
-metadata) while others use a collaborative filtering approach without taking into account user or song characteristic such as age, genre, location, etc.
+For example, some methods take more information about the user and the song into account (i.e.
+content-based) while others use a collaborative filtering approach without taking into account user or song characteristics such as age, genre, location, etc.
 There is no set similarity measure between two songs, as evident by the fact that songs by the same artist are often very different. As such, there are many different types of algorithms to explore.

### 2015-05-07 writeup/MSDgroupwriteup.tex
Update MSDgroupwriteup.tex

@@
 Intuitively, they work by recommending users songs that other similar users like (user-based) or by recommending songs which are similar to the songs the user has already listened to (item-based).
+The downside to this method is the computational complexity.
+Given the quadratic runtime of these algorithms, they are more difficult to scale to a larger dataset.
+As for storage requirements, we will have to store all pairs of similarity scores for users in the case of user-based and songs in the case of item-based.
+The details are described below.
 

### 2015-05-07 writeup/MSDgroupwriteup.tex

@@
 \subsection{Matrix Factorization}
-The goal of matrix factorization is to use collaborative methods to build a recommendation system for users based on user ratings of objects (in this case songs).
-This allowed us to recommend certain songs to users based on their listening history and without the need for using content based approaches.
-Since our data set contains number of plays for a given user and song (rather than rating), we normalized plays for every song on a scale between 0 and 1 and used this as a proxy for rating.
-
-We then constructed training and testing matrices, of which both have $N_1$ users (rows denoted by $u_i$) and $N_2$ songs (columns denoted by $v_j$).
-Of course the matrices were very sparse, containing zeros in all entries except for those in which a user (rows of matrix) has listened to a song (columns of matrix).
-The goal is to factor the training matrix into the product of two matrixes, $U$ and $V$.
-The matrix $U$ will be $N_1 \times d$ and the matrix $V$ will be $d \times N_2$.
-We want to learn a low-rank factorization (i.e.
-we choose $d$) so as to restrict the patterns we see in the rows and columns of our original matrix (e.g.
-we think a priori that if a user likes 1 top 100 song, the user may also like other top 100 songs).
-There is subjectivity in picking $d$ but $20$ is a common place to start.
+Matrix factorization is a type of collaborative filtering algorithm that can be used to build a recommendation system for users based on user feedback of objects (in this case songs).
+This allows us to recommend songs to users based on their listening history without the need for content based approaches.
+Training and testing matrices were constructed, of which both have $N_1$ users (rows denoted by $u_i$) and $N_2$ songs (columns denoted by $v_j$).
+As users have only listened to a fraction of all the songs available, the matrices are sparse, containing zeros in all entries except for those in which a user (a user representing a single row) has listened to.
+The goal of matrix factorization is to factor the training matrix into the product of two matrices $U$ and $V$.
+The matrix $U$ is $N_1 \times d$ and the matrix $V$ is $d \times N_2$.
+Another goal of matrix factorization is typically to learn a low-rank factorization of the original matrix while restricting the patterns that are generated in the product of the factorized matrix (e.g.
+if a user likes a song in a specific group of songs, the user may also like other songs in that group of songs).
+The choice is $d$ can be subjective but $20$ is a common value to start with.
 In the factorized matrices, the predicted rating will be $\hat{M}_{ij} = u_i^\TRANSPOSE  v_j$
-
 Using a coordinate ascent algorithm over 100 iterations, each row ($u_i$) and column ($v_j$) of the training matrix is then updated (equations \ref{eq1} and \ref{eq2}) in order to maximize the log joint likelihood (equation \ref{eq3}).
-
 \begin{equation}
@@
 \end{equation}
-
 \begin{equation}
@@
 
-\emph{Note: We use a rank 20 matrix for factorization purposes, a $\lambda = 10$, and calculate the variance of the observations for our $\sigma^2$.
-Also, $\Omega$ is the set of all indices in the matrix which have an observation}.
-
+\emph{Note}: $\lambda = 10$ was used and $\Omega$ is the set of all indices in the matrix which have an observation.
 We keep track of the root mean square error (RMSE) versus the testing set (i.e.
 how close our prediction is as compared to the actual normalized play count in the testing set) and the log joint likelihood of the training set as a function of iteration (see Figures \ref{fig:rmseplot} and \ref{fig:likeplot}).
+There are several hyper parameters that can be tweaked in order to achieve a better mAP result using matrix factorization.
+The main ones that  was varied were the variance, the rank d of the factorization, and the iterations.
+As the project was being done using Python and the main computational package (\emph{scikit-learn}) did not have probabilistic matrix factorization functions, we decided to apply the algorithm that we had wrote ourselves for a different class.
+
+After testing around with a few parameters, it was soon discovered that matrix factorization results were significantly below the popularity baseline and changing the hyper-parameters around did not help much.
+The best MAP value of 0.0032 were achieved was using a high rank d of 80 over 100 iterations with a variance of 0.0001.
+The log joint likelihood chart can be used to verify that the probabilistic matrix factorization algorithm is running properly in that the values are monotonically decreasing which corresponds to a decreasing training error.
+The plot shows that there is more room for improving the training error, but the MAP plot that was taken from the same run shows that there is marginal difference in terms of improving the MAP score after the initial 5 iterations.
+Due to this result, further exploration of the probabilistic matrix factorization algorithm was limited to 30 iterations to reduce computational time.
+
+One explanation for the very low MAP scores using PMF was that the number of plays that were used are different from what PMF is typically used for which is bounded ratings.
+Using implicit user feedback (number of plays) instead of explicit user feedback (ratings) have been found in literature to create various issues.
+One of the main issue is that implicit feedback is inherently noisy does not measure negative feedback.
+Users may have played a song to simply find out that the dislike the song.
+The other main issue is that the exact number of plays has could have significantly different meaning measured between different users.
+For example, an avid user who has played a song five times could have different meaning to a user that has at most only played songs five times.
+
+In order to tackle the issue of using implicit user feedback, we tried four different scaling schemas independently.
+These include normalizing user play counts by the total number of plays for the user, normalizing song play counts by the total number of plays for the song, making the data binary (changing all non-zero play counts to 1), and also applying tfidf.
+While testing out these schemas, we continued to vary the hyperparameters d and variance in order to determine if they changed the MAP results significantly.
+
+From the experiments, it was found that the rank d did not contribute significantly to the MAP values especially in settings where the MAP values were close to the baseline.
+Higher variance values were found to actually significantly increase the MAP values.
+In terms of scaling schemas, making the data binary produced the best result, with the second best being normalizing the data against the total number of plays for each user.
+Perhaps it can be said that using number of plays introduces more noise than simply considering if the user has played the song or not.
+Interestingly enough, normalizing the data against total number of plays for each song performed the worst and this most likely relates to the fact that number of plays is more meaningful relative to each user rather than to each song.
+
+Regardless, the best result achieved for PMF in the end was 0.0143 which was only slightly above the popularity baseline.
+Based on literature review, we have found that this is better or in line with what other people have achieved with using PMF on play counts.
+Nevertheless, this indicates that PMF is not suitable for the MSD dataset and this can be attributed to the implicit feedback issue or simply that the dataset is just too sparse.
+However, it is still possible to apply matrix factorization techniques to this dataset through using metadata and this has been demonstrated to be viable by \citet{li2012million} which achieved MAP values exceeding 0.1.
 

### 2015-05-07 writeup/MSDgroupwriteup.tex
small edits

@@
 
-For our project, we explored various song recommendation systems algorithms such as popularity recommendations, artist based recommensations, matrix factorization, and user and item based similarity measures.
+For our project, we explored various song recommendation systems algorithms such as popularity recommendations, artist based recommendations, matrix factorization, and user and item based similarity measures.
 % memory based collaborative filtering
@@
 We choose to subset the data to users who have listened to more than 27 songs and songs which were listened to by more than 22 users.
-This resulted in a datset which was 10\% the size of the original.
+This resulted in a dataset which was 10\% the size of the original.
 Despite subsetting the data, we are still dealing with a matrix of $8130$ users by $11861$ songs, so some of our algorithms will take a fair amount of time to run.
@@
 This was done by calculating the average precision at each element in the list of recommended items (i.e.
-percentage of correct items among first k recommendations) and averaging these for the first 500 recommendations (i.e.
-for each $k \epsilon {1, ..., 500}$, calculate the precision and average them together).
+percentage of correct items among first $k$ recommendations) and averaging these for the first 500 recommendations (i.e.
+for each $k \in \{1, ..., 500\}$, calculate the precision and average them together).
 We then averaged the score over all users.
@@
 
-\emph{Note}: $\lambda = 10$ was used and $\Omega$ is the set of all indices in the matrix which have an observation.
-We keep track of the root mean square error (RMSE) versus the testing set (i.e.
+\emph{Note: $\lambda = 10$ was used and $\Omega$ is the set of all indices in the matrix which have an observation.}
+%We keep track of the root mean square error (RMSE) versus the testing set (i.e.
 how close our prediction is as compared to the actual normalized play count in the testing set) and the log joint likelihood of the training set as a function of iteration (see Figures \ref{fig:rmseplot} and \ref{fig:likeplot}).
-There are several hyper parameters that can be tweaked in order to achieve a better mAP result using matrix factorization.
+
+There are several hyper parameters that can be tweaked in order to achieve a better MAP result using matrix factorization.
 The main ones that  was varied were the variance, the rank d of the factorization, and the iterations.
@@
 These include normalizing user play counts by the total number of plays for the user, normalizing song play counts by the total number of plays for the song, making the data binary (changing all non-zero play counts to 1), and also applying tfidf.
-While testing out these schemas, we continued to vary the hyperparameters d and variance in order to determine if they changed the MAP results significantly.
+While testing out these schemas, we continued to vary the hyper-parameters: $d$ and $\sigma$, in order to determine if they changed the MAP results significantly.
 
-From the experiments, it was found that the rank d did not contribute significantly to the MAP values especially in settings where the MAP values were close to the baseline.
+From the experiments, it was found that the rank $d$ did not contribute significantly to the MAP values especially in settings where the MAP values were close to the baseline.
 Higher variance values were found to actually significantly increase the MAP values.

### 2015-05-07 writeup/MSDgroupwriteup.tex
Update MSDgroupwriteup.tex

@@
 The matrix $U$ is $N_1 \times d$ and the matrix $V$ is $d \times N_2$.
-Another goal of matrix factorization is typically to learn a low-rank factorization of the original matrix while restricting the patterns that are generated in the product of the factorized matrix (e.g.
+Another goal of matrix factorization is typically to learn a low-rank factorization of the original matrix while restricting the patterns that are generated in the product of the factorized matrices (e.g.
 if a user likes a song in a specific group of songs, the user may also like other songs in that group of songs).
-The choice is $d$ can be subjective but $20$ is a common value to start with.
+The choice of $d$ can be subjective but $20$ is a common value to start with.
 In the factorized matrices, the predicted rating will be $\hat{M}_{ij} = u_i^\TRANSPOSE  v_j$
@@
 \begin{equation}
-v_j = \left( \lambda\sigma^2 I + \sum_{i \in \Omega_{v_j}} u_i u_i^\TRANSPOSE  \right)^{-1}\left(\sum_{j \in \Omega_{v_j}} M_{ij} u_{i} \right)
+v_j = \left( \lambda\sigma^2 I + \sum_{i \in \Omega_{v_j}} u_i u_i^\TRANSPOSE  \right)^{-1}\left(\sum_{i \in \Omega_{v_j}} M_{ij} u_{i} \right)
 \label{eq2}
@@
 
-There are several hyper parameters that can be tweaked in order to achieve a better MAP result using matrix factorization.
-The main ones that  was varied were the variance, the rank d of the factorization, and the iterations.
+There are several hyper-parameters that can be tweaked in order to achieve a better MAP result using matrix factorization.
+The main ones that we varied were the variance, the rank $d$ of the factorization, and the number of iterations.
 As the project was being done using Python and the main computational package (\emph{scikit-learn}) did not have probabilistic matrix factorization functions, we decided to apply the algorithm that we had wrote ourselves for a different class.
 
-After testing around with a few parameters, it was soon discovered that matrix factorization results were significantly below the popularity baseline and changing the hyper-parameters around did not help much.
-The best MAP value of 0.0032 were achieved was using a high rank d of 80 over 100 iterations with a variance of 0.0001.
+After testing various values of the parameters, it was soon discovered that matrix factorization results were significantly below the popularity baseline and changing the hyper-parameters around did not help much.
+The best MAP value of 0.0032 was achieved using a high rank $d$ of 80 over 100 iterations with a variance of 0.0001.
 The log joint likelihood chart can be used to verify that the probabilistic matrix factorization algorithm is running properly in that the values are monotonically decreasing which corresponds to a decreasing training error.
-The plot shows that there is more room for improving the training error, but the MAP plot that was taken from the same run shows that there is marginal difference in terms of improving the MAP score after the initial 5 iterations.
+The plot shows that there is more room for improvement of the training error, but the MAP plot that was taken from the same run shows that there is only a marginal difference in terms of improving the MAP score after the initial 5 iterations.
 Due to this result, further exploration of the probabilistic matrix factorization algorithm was limited to 30 iterations to reduce computational time.
@@
 Using implicit user feedback (number of plays) instead of explicit user feedback (ratings) have been found in literature to create various issues.
-One of the main issue is that implicit feedback is inherently noisy does not measure negative feedback.
-Users may have played a song to simply find out that the dislike the song.
-The other main issue is that the exact number of plays has could have significantly different meaning measured between different users.
-For example, an avid user who has played a song five times could have different meaning to a user that has at most only played songs five times.
+One of the main issues is that implicit feedback is inherently noisy and does not measure negative feedback.
+Users may have played a song to simply find out that they dislike the song.
+The other main issue is that the exact number of plays could have a significantly different meaning across different users.
+For example, an avid user who has played a song five times could have a different meaning to a user that has at most only played songs five times.
 
 In order to tackle the issue of using implicit user feedback, we tried four different scaling schemas independently.
-These include normalizing user play counts by the total number of plays for the user, normalizing song play counts by the total number of plays for the song, making the data binary (changing all non-zero play counts to 1), and also applying tfidf.
-While testing out these schemas, we continued to vary the hyper-parameters: $d$ and $\sigma$, in order to determine if they changed the MAP results significantly.
+These included normalizing user play counts by the total number of plays for the user, normalizing song play counts by the total number of plays for the song, making the data binary (changing all non-zero play counts to 1), and  applying tf-idf.
+While testing out these schemas, we continued to vary the hyper-parameters, $d$ and variance, in order to determine if they changed the MAP results significantly.
 
@@
 Based on literature review, we have found that this is better or in line with what other people have achieved with using PMF on play counts.
-Nevertheless, this indicates that PMF is not suitable for the MSD dataset and this can be attributed to the implicit feedback issue or simply that the dataset is just too sparse.
+Nevertheless, this indicates that PMF is not suitable for the MSD dataset and this can be attributed to the implicit feedback issue or simply to the fact that the dataset is too sparse.
 However, it is still possible to apply matrix factorization techniques to this dataset through using metadata and this has been demonstrated to be viable by \citet{li2012million} which achieved MAP values exceeding 0.1.

### 2015-05-07 writeup/MSDgroupwriteup.tex
Update MSDgroupwriteup.tex

@@
 
-We used two different baseline benchmarks to compare our MAP scores against.
+We used two different baseline benchmarks \citep{McFee:2012:MSD:2187980.2188222} to compare our MAP scores against.
 The first was to simply recommend the top 500 most popular songs to every user.
@@
 user-based) and songs (i.e.
-item-based) were calculated using cosine similarity.
+item-based) were calculated using cosine similarity \citep{aiolli2013preliminary} and \citep{li2012million}.
 Intuitively, they work by recommending users songs that other similar users like (user-based) or by recommending songs which are similar to the songs the user has already listened to (item-based).

### 2015-05-07 writeup/MSDgroupwriteup.tex
Update MSDgroupwriteup.tex

@@
 \begin{document}
-\title{Clustering of Last.fm Music}
+\title{Music Recommendation System Using the Million Song Dataset}
 
@@
 
-\subsection{Data}
+\subsection{Data and Statistics}
 The main source of data for this project was the Million Song Data Set (MSD)  \citep{Bertin-Mahieux2011}.
@@
 This resulted in a dataset which was 10\% the size of the original.
-Despite subsetting the data, we are still dealing with a matrix of $8130$ users by $11861$ songs, so some of our algorithms will take a fair amount of time to run.
-
-\subsubsection{Data Statistics}
+The sparsity of the data was .0016, meaning that we only had play counts for 0.16\% of the user, song pairs (i.e.
+not all users listen to all songs).
+Despite subsetting the data, though, we are still dealing with a matrix of $8130$ users by $11861$ songs, so some of our algorithms will take a fair amount of time to run.
 
@@
 
-$$w_i = \sum_{v \epsilon V} sim(u, v)$$
+$$w_i = \sum_{v \in V} sim(u, v)$$
 
@@
 
-$$w_b = \sum_{a \epsilon A} sim(a, b)$$
+$$w_b = \sum_{a \in A} sim(a, b)$$
 

### 2015-05-07 writeup/MSDgroupwriteup.tex
Update MSDgroupwriteup.tex

@@
 
-\subsection{Matrix Factorization}
+\subsection{Probabilistic Matrix Factorization}
 Matrix factorization is a type of collaborative filtering algorithm that can be used to build a recommendation system for users based on user feedback of objects (in this case songs).
@@
 The main ones that we varied were the variance, the rank $d$ of the factorization, and the number of iterations.
-As the project was being done using Python and the main computational package (\emph{scikit-learn}) did not have probabilistic matrix factorization functions, we decided to apply the algorithm that we had wrote ourselves for a different class.
+As the project was being done using Python and the main computational package (\emph{scikit-learn}) did not have probabilistic matrix factorization functions, we decided to apply the algorithm that we wrote ourselves for a different class \citep{koren2009matrix}.
 
@@
   \item The MSD Kaggle challenge: \url{https://www.kaggle.com/c/msdchallenge}
+  \item The user, song, play count dataset: \url{https://www.kaggle.com/c/msdchallenge/download/kaggle_visible_evaluation_triplets.zip}
 

### 2015-05-07 writeup/MSDgroupwriteup.tex
Update MSDgroupwriteup.tex

@@
 Nevertheless, this indicates that PMF is not suitable for the MSD dataset and this can be attributed to the implicit feedback issue or simply to the fact that the dataset is too sparse.
-However, it is still possible to apply matrix factorization techniques to this dataset through using metadata and this has been demonstrated to be viable by \citet{li2012million} which achieved MAP values exceeding 0.1.
+However, it is still possible to apply matrix factorization techniques to this dataset through using metadata and this has been demonstrated to be viable by \citep{liangcodebook} which achieved MAP values exceeding 0.1.
 

### 2015-05-07 writeup/MSDgroupwriteup.tex
new pdf

@@
 
-\section{Notes}
-Sparsity of $8.08199453451\times 10^{-5}$ (needs updating)
+%To Reference a paper:
 
-%The matrix was too big for our computers to handle the matrix factorization using the whole set of users and songs, so we subset the data. 
-%Subset songs by how many times the song was listened to. i.e the popularity of the song
-%= new number 
-%Subset users by how many songs they've listened to (how many popular song?)
-%= new number
-
-29 - users
-34 - songs 
-60 \%
-
-Interesting plots:
-- Cumulative Distribution of number of songs with a given play count
-
-To Reference a paper:
-
-The Million Song Dataset Challenge: \citep{McFee:2012:MSD:2187980.2188222} \\
-Million Song Dataset Recommendation Project Report: \citep{li2012million} \\
-codebook-based scalable music tagging with poisson matrix factorization: \citep{liangcodebook} \\
-Matrix Factorization Techniques for Recommender Systems: \citep{koren2009matrix} \\
-A Preliminary Study on a Recommender System for the Million Songs Dataset Challenge: \citep{aiolli2013preliminary} \\
+%The Million Song Dataset Challenge: \citep{McFee:2012:MSD:2187980.2188222} \\
+%Million Song Dataset Recommendation Project Report: \citep{li2012million} \\
+%codebook-based scalable music tagging with poisson matrix factorization: \citep{liangcodebook} \\
+%Matrix Factorization Techniques for Recommender Systems: \citep{koren2009matrix} \\
+%A Preliminary Study on a Recommender System for the Million Songs Dataset Challenge: \citep{aiolli2013preliminary} \\
 

### 2015-05-07 DataFrames.py

@@ -2,17 +2,28 @@ import numpy as np
 import sqlite3 as sqlite
 import pandas as pd
 import pandas.io.sql as psql
-import matplotlib.pyplot as plt
 import random
-
 import matplotlib
+import matplotlib.pyplot as plt
+from matplotlib.ticker import FuncFormatter
+
 matplotlib.style.use('ggplot')
+
+def to_percent(y, position):
+    # Ignore the passed in position. This has the effect of scaling the default
+    # tick locations.
+    s = str(100 * y)
+
+    # The percent symbol needs escaping in latex
+    if matplotlib.rcParams['text.usetex'] == True:
+        return s + r'$\%$'
+    else:
+        return s + '%'
 
 
 # 34 user plays, 29 song plays
 # 98 user plays, 52 song plays
 # 27 user songcounts, 22 song usercounts
-
 
 # http://www.kaggle.com/c/msdchallenge/data
 f = open("kaggle_visible_evaluation_triplets.txt", 'rb')
@@ -33,7 +44,6 @@ eval = pd.read_csv(f,sep='\t',header = None, names = ['user_id','sid','plays'])
 #len(set(sub['sid']))
 
 #sub.shape[0]/float(eval.shape[0])
-
 #sub.shape[0]/float(len(set(sub['user_id']))*len(set(sub['sid'])))
 
 # for number of songs instead of plays
@@ -57,31 +67,28 @@ sub.shape[0]/float(eval.shape[0])
 sub.shape[0]/float(len(set(sub['user_id']))*len(set(sub['sid'])))
 
 ##
-
 sub_max = pd.DataFrame(sub.groupby('sid').max()).reset_index()
 merged = pd.merge(sub,sub_max,on="sid")
 merged['plays_x'] = merged['plays_x']/merged['plays_y']
 sub_norm = merged[['user_id_x', 'sid', 'plays_x']]
 sub_norm.columns = ['user_id', 'sid', 'plays']
 
-sample = random.sample(sub_norm.index, int(sub_norm.shape[0]*0.2))
-trainsub = sub_norm.copy()
-trainsub.ix[trainsub.index.isin(sample),'plays'] = 0
+#sample = random.sample(sub_norm.index, int(sub_norm.shape[0]*0.2))
+#trainsub = sub_norm.copy()
+#trainsub.ix[trainsub.index.isin(sample),'plays'] = 0
 
-testsub = sub_norm.copy()
[…]

### 2015-05-07 writeup/MSDgroupwriteup.tex
change in table
Update MSDgroupwriteup.tex

@@
 
+\citep{hu2008collaborative}
+
 %\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
@@
 
-\section{Conclusion}
+\section{Results}
 After comparing the various algorithms, the simpler models gave us the best results.
 Our most complicated model, matrix factorization, gave us results barely above the popularity baseline.
-On the other hand, simply recommending users songs by artists they have already listened to (the artist-based baseline) resulted in a significantly better recommendation system.
+On the other hand, simply recommending users songs by artists they have already listened to (the artist-based popularity baseline) resulted in a significantly better recommendation system.
 Our song-based collaborative filtering technique produced the best results despite being a relatively simple algorithm as compared to matrix factorization.
@@
 \bf{Algorithm} &  \bf{MAP score}\\ \hline
-Artist Based  Baseline  & 0.0448    \\ 
+Artist Based Popularity Baseline  & 0.0448    \\ 
 %\hline
@@
 \end{table}
+
+\section{Conclusion}
+Apart from the algorithms discussed above, we also explored building a recommendation system based on K-Means Clustering (by both user and song) and Non-Negative Matrix Factorization (NMF).
+In the end, both of these methods yielded worse results as compared to Probabalistic Matrix Factorization so we decided not to explore them further.
+Next steps include incorporating tags and metadata present in the dataset (e.g.
+year, genre, audio metadata).
+It would also be useful to expand the size of the subset and distribute the workload to multiple machines so that more data could be analyzed.
 

### 2015-05-07 writeup/MSDgroupwriteup.tex
updates
plots

@@
 
-- Sparsity .0016086
--
-
 SONG PLAY COUNT  \\
@@
 We tested our algorithms against each other using a mean average precision 
-score for each method.  The mean average precision (MAP) calculation came directly from the Kaggle MSD challenge \citep{kaggleMSD}. The score was used to rank the competition entries.
+score for each method.
+The mean average precision (MAP) calculation came directly from the Kaggle MSD challenge \citep{McFee:2012:MSD:2187980.2188222}.
+The score was used to rank the competition entries.
 
@@
 \bf{Algorithm} &  \bf{MAP score}\\ \hline
+Item Based  CF  &  0.0479 \\ 
 Artist Based Popularity Baseline  & 0.0448    \\ 
-%\hline
+User Based  CF  &  0.0377 \\ 
+Matrix Factorization  &   0.0143  \\ 
 Top 500 Songs by Count  Baseline &  0.0138  \\ 
-%\hline
 Top 500 Songs by Plays  Baseline &  0.0126  \\ 
-%\hline
-User Based  CF  &  0.0377 \\ 
-%\hline
-Item Based  CF  &  0.0479 \\ 
-%\hline
-Matrix Factorization  &   0.0143  \\ 
-%\hline
 \end{tabular}

### 2015-05-07 writeup/MSDgroupwriteup.tex
Update MSDgroupwriteup.tex

@@
 \section{Introduction}
-In the online world, recommendation systems are becoming increasingly important in various industries including in retail, entertainment, and even dating. 
+In the online world, recommendation systems are becoming increasingly important in various industries including retail, entertainment, and even dating. 
 Because of their popularity and usefulness, there are numerous incentives to implement a good recommendation system. 
@@
 content-based) while others use a collaborative filtering approach without taking into account user or song characteristics such as age, genre, location, etc.
+In this paper, we focus on collaborative filtering methods.
 There is no set similarity measure between two songs, as evident by the fact that songs by the same artist are often very different. As such, there are many different types of algorithms to explore.

### 2015-05-07 writeup/MSDgroupwriteup.tex
Update MSDgroupwriteup.tex

@@
 One explanation for the very low MAP scores using PMF was that the number of plays that were used are different from what PMF is typically used for which is bounded ratings.
-Using implicit user feedback (number of plays) instead of explicit user feedback (ratings) have been found in literature to create various issues.
+Using implicit user feedback (number of plays) instead of explicit user feedback (ratings) has been found in literature to create various issues.
 One of the main issues is that implicit feedback is inherently noisy and does not measure negative feedback.
@@
 The other main issue is that the exact number of plays could have a significantly different meaning across different users.
-For example, an avid user who has played a song five times could have a different meaning to a user that has at most only played songs five times.
+For example, an avid user who has played a song five times could have a different meaning as compared to a user that has at most only played songs five times.
 

### 2015-05-07 DataFrames.py
plots

@@ -158,23 +158,6 @@ plt.xlabel("Number of Songs Played")
 plt.ylabel("Count")
 plt.show()
 
-#
-usercsum = usercount.groupby('plays').count().cumsum()
-usercsum = pd.DataFrame(usercsum).reset_index()
-#formatter = FuncFormatter(to_percent)
-#plt.gca().yaxis.set_major_formatter(formatter)
-#plt.show()
-
-usercsumlog = usercsum
-usercsumlog['plays'] = np.log10(usercsumlog['plays'])
-usercsumlog['user_id'] = usercsumlog['user_id']/usercsum['user_id'].max()
-usercsumlog.plot('plays','user_id')
-plt.title("CDF")
-plt.xlabel("Number of Songs Played (log scale)")
-plt.ylabel("Percentage of Users")
-formatter = FuncFormatter(to_percent)
-plt.gca().yaxis.set_major_formatter(formatter)
-plt.show()
 
 # histogram of number of user counts for songs
 songcount = eval.groupby('sid').count()
@@ -182,24 +165,41 @@ songcount = pd.DataFrame(songcount).reset_index()
 songvalues = songcount['plays'].values
 songvalues = np.log10(songvalues)
 plt.hist(songvalues,50)
-plt.title("histogram of number of plays for each songs")
-plt.xlabel("Number of user counts")
+plt.title("Number of Songs with a Given Number of Users")
+plt.xlabel("Number of Users")
 plt.ylabel("Count")
 plt.show()
 
-# cumulative sum for song counts vs users
-userplaycsum = userhist.sort('plays')
-userplaycsum['plays'] = userplaycsum['plays'].cumsum()
-userplaycsum['plays'] = userplaycsum['plays']/userplaycsum['plays'].max()
-userplaycsum = userplaycsum.reset_index(drop=True).reset_index()
-userplaycsum.plot('index', 'plays')
-plt.legend().set_visible(False)
-plt.title("Cumulative sum of plays for users")
-plt.ylabel("Perecentage of plays")
-plt.xlabel("Amount of users sorted with increasing amount of plays")
+
+#
+usercsum = usercount.groupby('plays').count().cumsum()
+usercsum = pd.DataFrame(usercsum).reset_index()
+usercsumlog = usercsum
+usercsumlog['plays'] = np.log10(usercsumlog['plays'])
[…]

### 2015-05-07 writeup/MSDgroupwriteup.tex

@@
 These included normalizing user play counts by the total number of plays for the user, normalizing song play counts by the total number of plays for the song, making the data binary (changing all non-zero play counts to 1), and  applying tf-idf.
-While testing out these schemas, we continued to vary the hyper-parameters, $d$ and variance, in order to determine if they changed the MAP results significantly.
+While testing out these schemas, we continued to vary the hyper-parameters (i.e.
+$d$ and variance) in order to determine if they changed the MAP results significantly.
 
@@
 
-Regardless, the best result achieved for PMF in the end was 0.0143 which was only slightly above the popularity baseline.
-Based on literature review, we have found that this is better or in line with what other people have achieved with using PMF on play counts.
+Regardless, the best result achieved for PMF in the end was 0.0143, which was only slightly above the popularity baseline.
+Based on literature review, we have found that this is better or in line with what others have achieved with using PMF on play counts.
 Nevertheless, this indicates that PMF is not suitable for the MSD dataset and this can be attributed to the implicit feedback issue or simply to the fact that the dataset is too sparse.
-However, it is still possible to apply matrix factorization techniques to this dataset through using metadata and this has been demonstrated to be viable by \citep{liangcodebook} which achieved MAP values exceeding 0.1.
+However, it is still possible to apply matrix factorization techniques to this dataset through using metadata and this has been demonstrated to be viable by \citet{liangcodebook} which achieved MAP values exceeding 0.1.
 

### 2015-05-07 DataFrames.py
plotting

@@ -170,7 +170,6 @@ plt.xlabel("Number of Users")
 plt.ylabel("Count")
 plt.show()
 
-
 #
 usercsum = usercount.groupby('plays').count().cumsum()
 usercsum = pd.DataFrame(usercsum).reset_index()
@@ -183,8 +182,8 @@ plt.xlabel("Number of Songs Played (log scale)")
 plt.ylabel("Percentage of Users")
 formatter = FuncFormatter(to_percent)
 plt.gca().yaxis.set_major_formatter(formatter)
+plt.legend().set_visible(False)
 plt.show()
-
 
 songcsum = songcount.groupby('plays').count().cumsum()
 songcsum = pd.DataFrame(songcsum).reset_index()
@@ -197,25 +196,10 @@ plt.xlabel("Number of Users (log scale)")
 plt.ylabel("Percentage of Songs")
 formatter = FuncFormatter(to_percent)
 plt.gca().yaxis.set_major_formatter(formatter)
+plt.legend().set_visible(False)
 plt.show()
 
-
-
-#cumulative sum of plays for artists
-#songcsum = songcount.groupby('plays').count().cumsum()
-#songcsum = pd.DataFrame(songcsum).reset_index()
-#songcsum.plot('plays','sid')
-#plt.show()
-
-#cumulative sum for plays for artists in log
-#songcsumlog = songcsum
-#songcsumlog['plays'] = np.log10(songcsumlog['plays'])
-#songcsumlog['sid'] = songcsumlog['sid']/songcsum['sid'].max()
-#songcsumlog.plot('plays','sid')
-#plt.show()
-
-
-# histogram of number of plays for users
+#histogram of number of plays for users
 #userhist = eval.groupby('user_id').sum()
 #userhist = pd.DataFrame(userhist).reset_index()
 #uservalues = userhist['plays'].values
@@ -226,7 +210,7 @@ plt.show()
 #plt.ylabel("Count")
 #plt.show()
 
-# cumulative sum for plays vs users
+#cumulative sum for plays vs users
 #userplaycsum = userhist.sort('plays')
 #userplaycsum['plays'] = userplaycsum['plays'].cumsum()
 #userplaycsum['plays'] = userplaycsum['plays']/userplaycsum['plays'].max()

### 2015-05-07 writeup/MSDgroupwriteup.tex
Update MSDgroupwriteup.tex

@@
 user-based) and songs (i.e.
-item-based) were calculated using cosine similarity \citep{aiolli2013preliminary} and \citep{li2012million}.
+item-based) were calculated using cosine similarity \citep{aiolli2013preliminary, li2012million}.
 Intuitively, they work by recommending users songs that other similar users like (user-based) or by recommending songs which are similar to the songs the user has already listened to (item-based).
@@
 Then, for each user, we looked at each song in the dataset and found all other users who have listened to that song.
-Next, we added up the similarity score of each of those users, v, with the original user, u, and got the weight for a particular song, $i$:
+Next, we added up the similarity score of each of those users, $v$, with the original user, $u$, and got the weight for a particular song, $i$:
 
@@
 If the same similar songs came up multiple times, we added the weights together.
-In other words, for each song, b, that was found to be similar to one of the songs a that the user listened to, we calculated the score for that song, $b$:
+In other words, for each song, $b$, that was found to be similar to one of the songs, $a$, that the user listened to, we calculated the score for that song, $b$:
 
@@
 Next steps include incorporating tags and metadata present in the dataset (e.g.
-year, genre, audio metadata).
+year, genre, and audio metadata).
 It would also be useful to expand the size of the subset and distribute the workload to multiple machines so that more data could be analyzed.

### 2015-05-08 writeup/MSDgroupwriteup.tex
Update MSDgroupwriteup.tex
correct plots

@@
 
-%\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
-%   \centering
-%   \includegraphics[width=4in]{h845BuzFO9CNgAAAABJRU5ErkJggg==.png} 
-%   \caption{how close our prediction is as compared to the actual normalized play count in the testing set }
-%   \label{fig:rmseplot}
-%\end{figure}
-%
-%\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
-%   \centering
-%   \includegraphics[width=4in]{C9LAAAAABJRU5ErkJggg==.png} 
-%   \caption{log joint likelihood of the training set as a function of iteration }
-%   \label{fig:likeplot}
-%\end{figure}
+\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
+   \centering
+   \includegraphics[width=4in]{h845BuzFO9CNgAAAABJRU5ErkJggg==.png} 
+   \caption{how close our prediction is as compared to the actual normalized play count in the testing set }
+   \label{fig:rmseplot}
+\end{figure}
+
+\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
+   \centering
+   \includegraphics[width=4in]{C9LAAAAABJRU5ErkJggg==.png} 
+   \caption{log joint likelihood of the training set as a function of iteration }
+   \label{fig:likeplot}
+\end{figure}
 

### 2015-05-08 DataFrames.py

@@ -170,35 +170,6 @@ plt.xlabel("Number of Users")
 plt.ylabel("Count")
 plt.show()
 
-#
-usercsum = usercount.groupby('plays').count().cumsum()
-usercsum = pd.DataFrame(usercsum).reset_index()
-usercsumlog = usercsum
-usercsumlog['plays'] = np.log10(usercsumlog['plays'])
-usercsumlog['user_id'] = usercsumlog['user_id']/usercsum['user_id'].max()
-usercsumlog.plot('plays','user_id')
-plt.title("Cumulative Sum of Songs Played by Users")
-plt.xlabel("Number of Songs Played (log scale)")
-plt.ylabel("Percentage of Users")
-formatter = FuncFormatter(to_percent)
-plt.gca().yaxis.set_major_formatter(formatter)
-plt.legend().set_visible(False)
-plt.show()
-
-songcsum = songcount.groupby('plays').count().cumsum()
-songcsum = pd.DataFrame(songcsum).reset_index()
-songcsumlog = songcsum
-songcsumlog['plays'] = np.log10(songcsumlog['plays'])
-songcsumlog['sid'] = songcsumlog['sid']/songcsum['sid'].max()
-songcsumlog.plot('plays','sid')
-plt.title("Cumulative Sum of Users")
-plt.xlabel("Number of Users (log scale)")
-plt.ylabel("Percentage of Songs")
-formatter = FuncFormatter(to_percent)
-plt.gca().yaxis.set_major_formatter(formatter)
-plt.legend().set_visible(False)
-plt.show()
-
 #histogram of number of plays for users
 #userhist = eval.groupby('user_id').sum()
 #userhist = pd.DataFrame(userhist).reset_index()
@@ -211,17 +182,37 @@ plt.show()
 #plt.show()
 
 #cumulative sum for plays vs users
-#userplaycsum = userhist.sort('plays')
-#userplaycsum['plays'] = userplaycsum['plays'].cumsum()
-#userplaycsum['plays'] = userplaycsum['plays']/userplaycsum['plays'].max()
-#userplaycsum = userplaycsum.reset_index(drop=True).reset_index()
-#userplaycsum.plot('index', 'plays')
-#plt.legend().set_visible(False)
-#plt.title("Cumulative sum of plays for users")
-#plt.ylabel("Perecentage of plays")
-#plt.xlabel("Amount of users sorted with increasing amount of plays")
+userplaycsum = userhist.sort('plays')
[…]

### 2015-05-08 algorithms/final/main.py
Starting final set of alogrithms

@@ -1,0 +1,109 @@
+import numpy as np
+import pandas as pd
+import random
+
+# load observation data
+f = open("kaggle_visible_evaluation_triplets.txt", 'rb')
+eval = pd.read_csv(f,sep='\t',header = None, names = ['user_id','sid','plays'])
+
+# count number of songs per user and subset
+userhist = eval.groupby('user_id').count()
+userhist = pd.DataFrame(userhist).reset_index()
+usersub = userhist[userhist['plays']>27]
+
+# count number of users per song and subset
+songhist = eval.groupby('sid').count()
+songhist = pd.DataFrame(songhist).reset_index()
+songsub = songhist[songhist['plays']>22]
+
+# subset the whole dataset
+sub = eval[eval['sid'].isin(songsub['sid'])]
+sub = sub[sub['user_id'].isin(usersub['user_id'])]
+
+# sampling/splitting the dataset
+sample = random.sample(sub.index, int(sub.shape[0]*0.2))
+trainsub = sub.copy()
+trainsub.ix[trainsub.index.isin(sample),'plays'] = 0
+testsub = sub.copy()
+testsub.ix[~trainsub.index.isin(sample),'plays'] = 0
+
+# generating a matrix out of dataframe
+trainpivot = trainsub.pivot(index='user_id',columns='sid', values='plays')
+# creating the mapping of user index to user id
+user_index = pd.DataFrame(trainpivot.index).reset_index()
+user_index.columns = [['user_index','user_id']]
+trainsub = pd.merge(trainsub, user_index, on='user_id')
+testsub = pd.merge(testsub, user_index, on='user_id')
+# creating the mapping of song index to song id
+song_index = pd.DataFrame(trainpivot.columns).reset_index()
+song_index.columns = [['song_index','sid']]
+trainsub = pd.merge(trainsub, song_index, on='sid')
+testsub = pd.merge(testsub, song_index, on='sid')
+
+# creating omegau_test
+testplays = testsub[testsub.plays>0]
+test_usergroup = testsub[testsub.plays>0].groupby('user_index')
+omegau_test = {}
+for i in list(set(testplays.user_index.values)):
+    omegau_test[i] = list(test_usergroup.get_group(i)['song_index'])
+# creating omegau_train
+trainplays = trainsub[trainsub.plays>0]
[…]

### 2015-05-08 algorithms/final/main.py
Updating main.py

@@ -76,14 +76,30 @@ M_train_binary[M_train_binary != 0] = 1
 plotdata = []
 
 # Bunch of plotting stuff
+# First and Second PMF plot
+MAP, L =  MF(M_train,M_test,1.0, omega, omega_test, omegau_train, omegau_test,80,100)
+
+plt.plot([0] + range(4,100,5), MAP)
+plt.xlabel('Iteration')
+plt.ylabel('MAP')
+plt.title('MAP across iterations for default PMF')
+plt.show()
+
+plt.plot(L)
+plt.xlabel('Iteration')
+plt.ylabel('Log Joint Likelihood')
+plt.title('Log Joint Likelihood across iterations for default PMF')
+plt.show()
+
+# Third PMF plot
 iter = 30
 for var in [100, 10, 1, 0.01, 0.001]:
     for d in [10, 20, 40, 80]:
-        plotdata.append([0,'Default',var,d,MF(M_train,M_test,var, omega, omega_test, omegau_train, omegau_test,d,iter)])
-        plotdata.append([1,'Row-norm',var,d,MF(M_train,M_test,var, omega, omega_test, omegau_train, omegau_test,d,iter)])
-        plotdata.append([2,'Col-norm',var,d,MF(M_train,M_test,var, omega, omega_test, omegau_train, omegau_test,d,iter)])
-        plotdata.append([3,'Binary',var,d,MF(M_train,M_test,var, omega, omega_test, omegau_train, omegau_test,d,iter)])
-        plotdata.append([4,'TF-IDF',var,d,MF(M_train,M_test,var, omega, omega_test, omegau_train, omegau_test,d,iter)])
+        plotdata.append([0,'Default',var,d,MF2(M_train,M_test,var, omegau_train, omegau_test,d,iter)])
+        plotdata.append([1,'Row-norm',var,d,MF2(M_train,M_test,var, omegau_train, omegau_test,d,iter)])
+        plotdata.append([2,'Col-norm',var,d,MF2(M_train,M_test,var, omegau_train, omegau_test,d,iter)])
+        plotdata.append([3,'Binary',var,d,MF2(M_train,M_test,var, omegau_train, omegau_test,d,iter)])
+        plotdata.append([4,'TF-IDF',var,d,MF2(M_train,M_test,var, omegau_train, omegau_test,d,iter)])
             
 plotdata = pd.DataFrame(plotdata,columns=['method','methodname','variance','d','MAP'])
 

### 2015-05-08 algorithms/final/main.py
Updating main.py again

@@ -73,8 +73,6 @@ M_train_tfidf = M_train_rownorm*idf
 M_train_binary = M_train.copy()
 M_train_binary[M_train_binary != 0] = 1
 
-plotdata = []
-
 # Bunch of plotting stuff
 # First and Second PMF plot
 MAP, L =  MF(M_train,M_test,1.0, omega, omega_test, omegau_train, omegau_test,80,100)
@@ -92,6 +90,8 @@ plt.title('Log Joint Likelihood across iterations for default PMF')
 plt.show()
 
 # Third PMF plot
+plotdata = []
+
 iter = 30
 for var in [100, 10, 1, 0.01, 0.001]:
     for d in [10, 20, 40, 80]:

### 2015-05-08 algorithms/final/main.py
Updating main.py again and again

@@ -2,6 +2,7 @@ import numpy as np
 import pandas as pd
 import random
 
+### Data Loading Section
 # load observation data
 f = open("kaggle_visible_evaluation_triplets.txt", 'rb')
 eval = pd.read_csv(f,sep='\t',header = None, names = ['user_id','sid','plays'])
@@ -52,7 +53,23 @@ train_usergroup = trainplays.groupby('user_index')
 omegau_train = {}
 for i in list(set(trainplays.user_index.values)):
     omegau_train[i] = list(train_usergroup.get_group(i)['song_index'])
+    
+# creating tuple lists
+omega = [tuple(x) for x in trainsub[trainsub.plays>0][['user_index','song_index']].values]
+omega_test = [tuple(x) for x in testsub[trainsub.plays>0][['user_index','song_index']].values]
 
+### Data Exploratory and Plotting Section
+
+
+
+### Popularity Baseline Section
+
+
+
+### Artist-based Popularity Baseline Section
+
+
+### PMF Section
 # Processing M_train, row normalization, column normalization, and tfidf
 rowsum = np.sum(M_train,axis=1).reshape((-1,1))
 rowsum[rowsum == 0] = 1E-16
@@ -123,3 +140,14 @@ for name, group in groups:
 plt.legend(bbox_to_anchor=[-0.1, 2.3], loc='center', ncol=5)
 plt.suptitle("PMF over different hyper-parameters and scaling methods", fontsize = 20)
 plt.show()
+
+
+### Item and user-based section
+
+
+
+
+### K-Means section
+
+
+### NMF section

### 2015-05-08 algorithms/final/main.py
Updating main.py....

@@ -112,11 +112,11 @@ plotdata = []
 iter = 30
 for var in [100, 10, 1, 0.01, 0.001]:
     for d in [10, 20, 40, 80]:
-        plotdata.append([0,'Default',var,d,MF2(M_train,M_test,var, omegau_train, omegau_test,d,iter)])
-        plotdata.append([1,'Row-norm',var,d,MF2(M_train,M_test,var, omegau_train, omegau_test,d,iter)])
-        plotdata.append([2,'Col-norm',var,d,MF2(M_train,M_test,var, omegau_train, omegau_test,d,iter)])
-        plotdata.append([3,'Binary',var,d,MF2(M_train,M_test,var, omegau_train, omegau_test,d,iter)])
-        plotdata.append([4,'TF-IDF',var,d,MF2(M_train,M_test,var, omegau_train, omegau_test,d,iter)])
+        plotdata.append([0,'Default',var,d,MF2(M_train,M_test,var, omegau_train, omegau_test, d, iter)])
+        plotdata.append([1,'Row-norm',var,d,MF2(M_train,M_test,var, omegau_train, omegau_test, d, iter)])
+        plotdata.append([2,'Col-norm',var,d,MF2(M_train,M_test,var, omegau_train, omegau_test, d, iter)])
+        plotdata.append([3,'Binary',var,d,MF2(M_train,M_test,var, omegau_train, omegau_test, d, iter)])
+        plotdata.append([4,'TF-IDF',var,d,MF2(M_train,M_test,var, omegau_train, omegau_test, d, iter)])
             
 plotdata = pd.DataFrame(plotdata,columns=['method','methodname','variance','d','MAP'])
 

### 2015-05-08 algorithms/final/main.py
Updating main.py?

@@ -1,6 +1,10 @@
 import numpy as np
 import pandas as pd
+import matplotlib as plt
+import matplotlib
 import random
+
+matplotlib.style.use('ggplot')
 
 ### Data Loading Section
 # load observation data
@@ -40,6 +44,14 @@ song_index = pd.DataFrame(trainpivot.columns).reset_index()
 song_index.columns = [['song_index','sid']]
 trainsub = pd.merge(trainsub, song_index, on='sid')
 testsub = pd.merge(testsub, song_index, on='sid')
+
+# Generating the default M_train and M_test matrices
+M_train = trainpivot.as_matrix()
+M_train = np.nan_to_num(M_train)
+
+testpivot = testsub.pivot(index='user_id',columns='sid', values='plays')
+M_test = testpivot.as_matrix()
+M_test = np.nan_to_num(M_test)
 
 # creating omegau_test
 testplays = testsub[testsub.plays>0]

### 2015-05-08 writeup/MSDgroupwriteup.tex
?
smaller text size + images

@@
-\documentclass[12pt,preprint]{aastex}
+\documentclass[11pt,preprint]{aastex}
  %\documentclass[12pt]{emulateapj}
@@
 
-SONG PLAY COUNT  \\
-mean = 28 \\
-max = 35432 \\
-min = 1 \\
-std = 215.826789\\
+\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
+   \centering
+   \includegraphics[width=4in]{../plots/final/cumsum-songs.png} 
+   \caption{ }
+   \label{fig:t}
+\end{figure}
 
-USER PLAY COUNT\\
-mean = 42\\
-max = 1305\\
-min = 5\\
-std = 53.31547\\
+\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
+   \centering
+   \includegraphics[width=4in]{../plots/final/cumsum-users.png} 
+   \caption{ }
+   \label{fig:t}
+\end{figure}
+
+\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
+   \centering
+   \includegraphics[width=4in]{../plots/final/hist-song.png} 
+   \caption{ }
+   \label{fig:t}
+\end{figure}
+
+\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
+   \centering
+   \includegraphics[width=4in]{../plots/final/hist_user_song_count.png} 
+   \caption{ }
+   \label{fig:t}
+\end{figure}
 
@@
 Using a coordinate ascent algorithm over 100 iterations, each row ($u_i$) and column ($v_j$) of the training matrix is then updated (equations \ref{eq1} and \ref{eq2}) in order to maximize the log joint likelihood (equation \ref{eq3}).
+
 \begin{equation}
@@
 \end{equation}
+
 \begin{equation}
@@
    \centering
-   \includegraphics[width=4in]{h845BuzFO9CNgAAAABJRU5ErkJggg==.png} 
-   \caption{how close our prediction is as compared to the actual normalized play count in the testing set }
-   \label{fig:rmseplot}
-\end{figure}
-
-\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
-   \centering
-   \includegraphics[width=4in]{C9LAAAAABJRU5ErkJggg==.png} 
-   \caption{log joint likelihood of the training set as a function of iteration }
-   \label{fig:likeplot}
+   \includegraphics[width=6in]{../plots/final/PMF.png} 
+   \caption{ }
+   \label{fig:t}
 \end{figure}

### 2015-05-08 algorithms/final/main.py
update final code

@@ -4,6 +4,9 @@ import matplotlib as plt
 import matplotlib
 import random
 
+from UserAndSongBasedRec import user_based_rec
+from UserAndSongBasedRec import item_based_rec
+
 matplotlib.style.use('ggplot')
 
 ### Data Loading Section
@@ -65,6 +68,14 @@ train_usergroup = trainplays.groupby('user_index')
 omegau_train = {}
 for i in list(set(trainplays.user_index.values)):
     omegau_train[i] = list(train_usergroup.get_group(i)['song_index'])
+    
+# creating omegav_train
+trainplay = trainsub[trainsub.plays>0] 
+train_songgroup = trainplay.groupby('song_index')
+omegav_train = {}
+
+for i in list(set(trainplay.song_index.values)):
+    omegav_train[i] = list(train_songgroup.get_group(i)['user_index'])   
     
 # creating tuple lists
 omega = [tuple(x) for x in trainsub[trainsub.plays>0][['user_index','song_index']].values]
@@ -132,7 +143,6 @@ for var in [100, 10, 1, 0.01, 0.001]:
             
 plotdata = pd.DataFrame(plotdata,columns=['method','methodname','variance','d','MAP'])
 
-plotdata = pd.merge(plotdata,method_index,on='method')
 plotdata['logvar'] = np.log10(plotdata['variance'])
 
 groups = plotdata.sort(['d','method']).groupby('d')
@@ -156,7 +166,13 @@ plt.show()
 
 ### Item and user-based section
 
+#Call item-based recommendation function
+item_based_map = item_based_rec(omegau_train, omegav_train, np.shape(M_train)[1], omegau_test)
+print item_based_map
 
+#Call user-based recommendation fucntion
+user_based_map = user_based_rec(omegau_train, omegav_train, np.shape(M_train)[0], omegau_test)
+print user_based_map
 
 
 ### K-Means section

### 2015-05-08 algorithms/final/artist_based.py
Updating algorithms

@@ -1,0 +1,21 @@
+import numpy as np
+import pandas as pd
+import random
+
+from apk import apk
+
+def artist_based(user_songs, artist_songdict, popularsonglist, omegau_test, omegau_train):
+    apk_sum = 0
+
+    for key_user_index in list(set(omegau_test).intersection(omegau_train)):
+        songlist = []
+        for key_artist_id in list(user_songs[user_songs.user_index==key_user_index].artist_id.values):
+            songlist += artist_songdict[key_artist_id]
+        # appending the popularsonglist to the generated songlist, assume 600 is more than enough to account for any overlaps
+        songlist = list(OrderedDict.fromkeys(songlist + popularsonglist))[:600]
+        # starting from np.delete, this function is just to remove the songs in songlist that already exist in omegau_train
+        apk_sum += apk(omegau_test[key_user_index],np.delete(songlist,np.nonzero(np.in1d(songlist,omegau_train[key_user_index])))[:500])
+
+    return apk_sum/len(list(set(omegau_test).intersection(omegau_train)))
+
+

### 2015-05-08 algorithms/final/main.py

@@ -4,6 +4,13 @@ import matplotlib as plt
 import matplotlib
 import random
 
+from matrix_factorization import MF
+from matrix_factorization import MF2
+
+from popularity_baseline import popularity_baseline
+
+from artist_based import artist_based
+
 from UserAndSongBasedRec import user_based_rec
 from UserAndSongBasedRec import item_based_rec
 
@@ -68,13 +75,10 @@ train_usergroup = trainplays.groupby('user_index')
 omegau_train = {}
 for i in list(set(trainplays.user_index.values)):
     omegau_train[i] = list(train_usergroup.get_group(i)['song_index'])
-    
 # creating omegav_train
-trainplay = trainsub[trainsub.plays>0] 
-train_songgroup = trainplay.groupby('song_index')
+train_songgroup = trainplays.groupby('song_index')
 omegav_train = {}
-
-for i in list(set(trainplay.song_index.values)):
+for i in list(set(trainplays.song_index.values)):
     omegav_train[i] = list(train_songgroup.get_group(i)['user_index'])   
     
 # creating tuple lists
@@ -86,8 +90,8 @@ omega_test = [tuple(x) for x in testsub[trainsub.plays>0][['user_index','song_in
 
 
 ### Popularity Baseline Section
-
-
+popbase = popularity_baseline(trainsub,omegau_train,omegau_test)
+print "The popularity baseline based on counts is " + str(popbase[0]) + ' and based on plays is ' + str(popbase[1])
 
 ### Artist-based Popularity Baseline Section
 

### 2015-05-08 algorithms/final/main.py
Updating main and adding loaddataframes
Loaddataframes.py isn't used in main.py, just additional code written to
load data

@@ -3,6 +3,9 @@ import pandas as pd
 import matplotlib as plt
 import matplotlib
 import random
+import pandas.io.sql as psql
+import sqlite3 as sqlite
+from collections import OrderedDict
 
 from matrix_factorization import MF
 from matrix_factorization import MF2
@@ -18,6 +21,7 @@ matplotlib.style.use('ggplot')
 
 ### Data Loading Section
 # load observation data
+# kaggle_visible_evaluation_triplets.txt is available at https://www.kaggle.com/c/msdchallenge/data
 f = open("kaggle_visible_evaluation_triplets.txt", 'rb')
 eval = pd.read_csv(f,sep='\t',header = None, names = ['user_id','sid','plays'])
 
@@ -85,15 +89,60 @@ for i in list(set(trainplays.song_index.values)):
 omega = [tuple(x) for x in trainsub[trainsub.plays>0][['user_index','song_index']].values]
 omega_test = [tuple(x) for x in testsub[trainsub.plays>0][['user_index','song_index']].values]
 
+
+
 ### Data Exploratory and Plotting Section
+
+
 
 
 
 ### Popularity Baseline Section
 popbase = popularity_baseline(trainsub,omegau_train,omegau_test)
-print "The popularity baseline based on counts is " + str(popbase[0]) + ' and based on plays is ' + str(popbase[1])
+print "The popularity baseline MAP based on counts is " + str(popbase[0]) + ' and based on plays is ' + str(popbase[1])
 
 ### Artist-based Popularity Baseline Section
+
+# Load track metadata to get artist_id
+# http://labrosa.ee.columbia.edu/millionsong/sites/default/files/AdditionalFiles/track_metadata.db
+con = sqlite.connect("track_metadata.db")
+with con:
+    sql = "SELECT * FROM songs"
+    track_metadata = psql.read_sql(sql, con)
+con.close()
+
+# Renaming columns to match other dataframes
+track_metadata=track_metadata.rename(columns = {'track_id':'tid'})
+track_metadata=track_metadata.rename(columns = {'song_id':'sid'})
+# Subsetting track_metadata to only have songs on the song_index on the subset data
+track_metadata=pd.merge(track_metadata,song_index, on='sid')
+
+# Group by songs and sum the amount of plays, then sort in descending amount of plays
+eval_songs = eval.groupby('sid').sum().reset_index().sort('plays',ascending=False)[['sid','plays']]
+# Subset songs to subset while including track metadata
[…]

### 2015-05-08 algorithms/final/main.py
Adding a plot and updating main.py to create plot

@@ -170,16 +170,16 @@ M_train_binary[M_train_binary != 0] = 1
 # First and Second PMF plot
 MAP, L =  MF(M_train,M_test,1.0, omega, omega_test, omegau_train, omegau_test,80,100)
 
+plt.subplot(1,2,1)
 plt.plot([0] + range(4,100,5), MAP)
-plt.xlabel('Iteration')
+plt.xlabel('Iterations')
 plt.ylabel('MAP')
 plt.title('MAP across iterations for default PMF')
-plt.show()
-
+plt.subplot(1,2,2)
 plt.plot(L)
-plt.xlabel('Iteration')
-plt.ylabel('Log Joint Likelihood')
-plt.title('Log Joint Likelihood across iterations for default PMF')
+plt.xlabel('Iterations')
+plt.ylabel('Log joint likelihood')
+plt.title('Log joint likelihood across iterations for default PMF')
 plt.show()
 
 # Third PMF plot

### 2015-05-08 writeup/MSDgroupwriteup.tex
table
Update MSDgroupwriteup.tex

@@
 We used the sample data set from the Kaggle competition which contains tuples of user id, song id, and play count.
+We also used data which links songs to artists through track ids.
+The links to these datasets can be found in the Appendix.
 
-The sample data set had $110000$ users, $163206$ songs, and the number of play counts for each user and song pair.
+The sample data set had $110000$ users, $163206$ songs, and the number of play counts for each user and song pair. The statistics are summarized below:
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{lcc}
+\hline
+Measurement & Plays Across Songs & Plays Across Users \\
+\hline
+Count &  163206 & 110000 \\
+Mean  &     8.890194 & 13.190300  \\
+Std      &    46.780258 & 8.070827 \\
+Min      &    1.0  & 5.0  \\
+Max     &   5043 &  53 \\
+\end{tabular}
+\end{center}
+\end{table}
+
 The scale of the data was too large to explore on our computers, so we subset the data down to a size easily handled by an average laptop.
@@
    \includegraphics[width=4in]{../plots/final/cumsum-songs.png} 
-   \caption{ }
+   \caption{Percentage of users accounted for versus songs with increasing number of plays.}
    \label{fig:t}
@@
    \includegraphics[width=4in]{../plots/final/cumsum-users.png} 
-   \caption{ }
+   \caption{Percentage of songs accounted for versus users with increasing number of plays.}
    \label{fig:t}
@@
 
-%\subsection{Naked Kmeans }
-%\subsection{Non-Negative Matrix Factorization}
-
 \subsection{User Based and Item Based Collaborative Filtering}
@@
 \begin{tabular}{lc}
-
 \hline
@@
   \item The user, song, play count dataset: \url{https://www.kaggle.com/c/msdchallenge/download/kaggle_visible_evaluation_triplets.zip}
+  \item Data to link songs to artists through track ids: \url{http://labrosa.ee.columbia.edu/millionsong/sites/default/files/AdditionalFiles/track_metadata.db}
 

### 2015-05-08 DataFrames.py
captions and table updates

@@ -155,7 +155,7 @@ uservalues = usercount['plays'].values
 plt.hist(uservalues,50)
 plt.title("Number of Users Who've played a Given Number of Songs")
 plt.xlabel("Number of Songs Played")
-plt.ylabel("Count")
+plt.ylabel("Count of Users")
 plt.show()
 
 
@@ -165,9 +165,9 @@ songcount = pd.DataFrame(songcount).reset_index()
 songvalues = songcount['plays'].values
 songvalues = np.log10(songvalues)
 plt.hist(songvalues,50)
-plt.title("Number of Songs with a Given Number of Users")
-plt.xlabel("Number of Users")
-plt.ylabel("Count")
+plt.title("Number of Songs that have a Given Number of Listeners")
+plt.xlabel("Number of Listeners (log scale)")
+plt.ylabel("Count of Songs")
 plt.show()
 
 #histogram of number of plays for users
@@ -198,6 +198,7 @@ plt.legend().set_visible(False)
 plt.show()
 
 
+#cumulative sum for plays vs songs
 songplaycsum = songhist.sort('plays')
 songplaycsum['plays'] = songplaycsum['plays'].cumsum()
 songplaycsum['plays'] = songplaycsum['plays']/songplaycsum['plays'].max()
@@ -236,16 +237,4 @@ plt.show()
 #songcsumlog['plays'] = np.log10(songcsumlog['plays'])
 #songcsumlog['sid'] = songcsumlog['sid']/songcsum['sid'].max()
 #songcsumlog.plot('plays','sid')
-#plt.show()
-
-# cumulative sum for plays vs songs
-#songplaycsum = songhist.sort('plays')
-#songplaycsum['plays'] = songplaycsum['plays'].cumsum()
-#songplaycsum['plays'] = songplaycsum['plays']/songplaycsum['plays'].max()
-#songplaycsum = songplaycsum.reset_index(drop=True).reset_index()
-#songplaycsum.plot('index', 'plays')
-#plt.legend().set_visible(False)
-#plt.title("Cumulative sum of plays for songs")
-#plt.ylabel("Perecentage of plays")
-#plt.xlabel("Amount of songs sorted with increasing amount of plays")
 #plt.show()

### 2015-05-08 writeup/MSDgroupwriteup.tex
captions and table updates
Updating .tex

@@
 \begin{center}
-\begin{tabular}{lcc}
+\begin{tabular}{lll}
 \hline
@@
 \emph{Note: $\lambda = 10$ was used and $\Omega$ is the set of all indices in the matrix which have an observation.}
-%We keep track of the root mean square error (RMSE) versus the testing set (i.e.
-how close our prediction is as compared to the actual normalized play count in the testing set) and the log joint likelihood of the training set as a function of iteration (see Figures \ref{fig:rmseplot} and \ref{fig:likeplot}).
 
@@
 After testing various values of the parameters, it was soon discovered that matrix factorization results were significantly below the popularity baseline and changing the hyper-parameters around did not help much.
-The best MAP value of 0.0032 was achieved using a high rank $d$ of 80 over 100 iterations with a variance of 0.0001.
+The best MAP value of 0.0038 was achieved using a high rank $d$ of 80 over 100 iterations with a variance of 1.0.
 The log joint likelihood chart can be used to verify that the probabilistic matrix factorization algorithm is running properly in that the values are monotonically decreasing which corresponds to a decreasing training error.
@@
 Due to this result, further exploration of the probabilistic matrix factorization algorithm was limited to 30 iterations to reduce computational time.
+
+\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
+   \centering
+   \includegraphics[width=6in]{../plots/final/defaultPMF.png} 
+   \caption{MAP values and log joint likelihood values for PMF using default playcount values with $\sigma^2 = 1.0$ and $d = 80$}
+   \label{defaultPMF}
+\end{figure}
 
@@
 
+\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
+   \centering
+   \includegraphics[width=6in]{../plots/final/PMF.png} 
+   \caption{MAP values for different scaling methods and hyperparameters after 30 iterations}
+   \label{PMF}
+\end{figure}
+
 Regardless, the best result achieved for PMF in the end was 0.0143, which was only slightly above the popularity baseline.
@@
 \citep{hu2008collaborative}
-
-\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
-   \centering
-   \includegraphics[width=6in]{../plots/final/PMF.png} 
-   \caption{ }
-   \label{fig:t}
-\end{figure}
 

### 2015-05-08 writeup/MSDgroupwriteup.tex
done

@@
 
-The sample data set had $110000$ users, $163206$ songs, and the number of play counts for each user and song pair. The statistics are summarized below:
+The sample data set had $110000$ users, $163206$ songs, and the number of play counts for each user and song pair. The statistics are summarized in Table \ref{tab:stats}:
 
@@
 \end{tabular}
+\caption{Basic summary statistics for the full data set.}
+\label{tab:stats}
 \end{center}
@@
 
-\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
-   \centering
-   \includegraphics[width=4in]{../plots/final/hist-song.png} 
-   \caption{ }
-   \label{fig:t}
-\end{figure}
-
-\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
-   \centering
-   \includegraphics[width=4in]{../plots/final/hist_user_song_count.png} 
-   \caption{ }
-   \label{fig:t}
-\end{figure}
+%\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
+%   \centering
+%   \includegraphics[width=4in]{../plots/final/hist-song.png} 
+%   \caption{ }
+%   \label{fig:t}
+%\end{figure}
+%
+%
+%\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
+%   \centering
+%   \includegraphics[width=4in]{../plots/final/hist_user_song_count.png} 
+%   \caption{ }
+%   \label{fig:t}
+%\end{figure}
 
@@
 \subsubsection{User-Based Collaborative Filtering}
-For user-based recommendation, we first calculated the similarity score between every pair of users, $u$ and $v$, using the equation below:
+For user-based recommendation, we first calculated the similarity score between every pair of users, $u$ and $v$, using equation \ref{eq:item}:
 
@@
 sim(u,v) = \frac{\text{\# common items}(u, v)}{{\text{\# items}(u)}^{1/2} \times {\text{\# items}(v)}^{1/2}}
+\label{eq:item}
 \end{equation}
@@
 \subsubsection{Item-Based Collaborative Filtering}
-For item-based recommendation, which turned out to be our best algorithm in terms of a MAP value, we first calculated the similarity score between every pair of songs, $i$ and $j$, using the equation below and saved the most similar songs to each song in the process:
+For item-based recommendation, which turned out to be our best algorithm in terms of a MAP value, we first calculated the similarity score between every pair of songs, $i$ and $j$, using equation \ref{eq:user} and saved the most similar songs to each song in the process:
 
@@
 sim(i,j) = \frac{\text{\# common users}(i, j)}{{\text{\# items}(i)}^{1/2} \times {\text{\# items}(j)}^{1/2}}
+\label{eq:user}
 \end{equation}

### 2015-05-08 writeup/MSDgroupwriteup.tex
Update MSDgroupwriteup.tex

@@
 \end{tabular}
-\caption{Basic summary statistics for the full data set.}
-\label{tab:stats}
+\caption{Basic summary statistics for the full data set.}\label{tab:stats}
 \end{center}
 \end{table}
+
+It is also easy to see in the below chart that the majority of the songs are listened to by very few users so we can decrease the number of songs in our subset but not lose all of the play count.
+
+\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
+   \centering
+   \includegraphics[width=4in]{../plots/final/cumsum-songs.png} 
+   \caption{Percentage of users accounted for versus songs with increasing number of plays.}
+   \label{fig:t}
+\end{figure}
 
@@
 
-\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
-   \centering
-   \includegraphics[width=4in]{../plots/final/cumsum-songs.png} 
-   \caption{Percentage of users accounted for versus songs with increasing number of plays.}
-   \label{fig:t}
-\end{figure}
-
-\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
-   \centering
-   \includegraphics[width=4in]{../plots/final/cumsum-users.png} 
-   \caption{Percentage of songs accounted for versus users with increasing number of plays.}
-   \label{fig:t}
-\end{figure}
+%\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
+%   \centering
+%   \includegraphics[width=4in]{../plots/final/cumsum-users.png} 
+%   \caption{Percentage of songs accounted for versus users with increasing number of plays.}
+%   \label{fig:t}
+%\end{figure}
 

### 2015-05-08 writeup/MSDgroupwriteup.tex
Update MSDgroupwriteup.tex
Updating .tex

@@
 After testing various values of the parameters, it was soon discovered that matrix factorization results were significantly below the popularity baseline and changing the hyper-parameters around did not help much.
-The best MAP value of 0.0038 was achieved using a high rank $d$ of 80 over 100 iterations with a variance of 1.0.
-The log joint likelihood chart can be used to verify that the probabilistic matrix factorization algorithm is running properly in that the values are monotonically decreasing which corresponds to a decreasing training error.
-The plot shows that there is more room for improvement of the training error, but the MAP plot that was taken from the same run shows that there is only a marginal difference in terms of improving the MAP score after the initial 5 iterations.
-Due to this result, further exploration of the probabilistic matrix factorization algorithm was limited to 30 iterations to reduce computational time.
+The best MAP value of 0.0039 was achieved using a high rank $d$ of 80 over 100 iterations with a variance of 1.0.
+The log joint likelihood chart can be used to verify that the probabilistic matrix factorization algorithm is running properly in that the values are monotonically increasing, which corresponds to a decreasing training error.
+The first graph in Figure /ref{defaultPMF} shows that the log joint likelihood is indeed monotonically increasing and slowed down significantly after the first twenty iterations.
+The second graph in the same figure shows the MAP values from the same run and it shows that the MAP values decreasing after the maximum at five iterations.
+It could be argued that emphasis should be placed on finding the right number of iterations to produce the best MAP values, but since the overall MAP results were poor and the differences were minimal, it was decided that this was not to be a priority for further experimentation.
+Due to this conclusion, further exploration of the probabilistic matrix factorization algorithm was limited to 30 iterations in order to reduce computational time.
 
@@
 One explanation for the very low MAP scores using PMF was that the number of plays that were used are different from what PMF is typically used for which is bounded ratings.
-Using implicit user feedback (number of plays) instead of explicit user feedback (ratings) has been found in literature to create various issues.
+Using implicit user feedback (number of plays) instead of explicit user feedback (ratings) has been found in literature \citep{hu2008collaborative} to create various issues.
 One of the main issues is that implicit feedback is inherently noisy and does not measure negative feedback.
@@
 
-From the experiments, it was found that the rank $d$ did not contribute significantly to the MAP values especially in settings where the MAP values were close to the baseline.
-Higher variance values were found to actually significantly increase the MAP values.
+The results from the experiments were plotted in Figure \ref{PMF}.
+It was found that the rank $d$ did not contribute significantly to the MAP values especially in settings where the MAP values were close to the popularity baseline.
+Higher variance values were found to actually significantly increase the MAP values and it can be argued that the reason why the default PMF works best at variance of 10 instead of 1.0 for the others is that the empirical variance of play count values is at least an order of magnitude higher than the others.
+This also suggests that the variance used in PMF should scale with the empirical variance, but the exact relationship remains unclear.
+
 In terms of scaling schemas, making the data binary produced the best result, with the second best being normalizing the data against the total number of plays for each user.
 Perhaps it can be said that using number of plays introduces more noise than simply considering if the user has played the song or not.
+TF-IDF performed slightly worse than the two, so down-weighting the songs shared more often by users did not produce better recommendations.
 Interestingly enough, normalizing the data against total number of plays for each song performed the worst and this most likely relates to the fact that number of plays is more meaningful relative to each user rather than to each song.
@@
 Regardless, the best result achieved for PMF in the end was 0.0143, which was only slightly above the popularity baseline.
-Based on literature review, we have found that this is better or in line with what others have achieved with using PMF on play counts.
+Based on literature review \citep{li20122million, McFee:2012:MSD:2187980.2188222}, we have found that this is better or in line with what others have achieved with using PMF on play counts.
 Nevertheless, this indicates that PMF is not suitable for the MSD dataset and this can be attributed to the implicit feedback issue or simply to the fact that the dataset is too sparse.
 However, it is still possible to apply matrix factorization techniques to this dataset through using metadata and this has been demonstrated to be viable by \citet{liangcodebook} which achieved MAP values exceeding 0.1.
-
-\citep{hu2008collaborative}
 
@@
 \end{tabular}
+\caption{Final MAP results across various algorithms.}\label{tab:stats}
 \end{center}

### 2015-05-08 writeup/MSDgroupwriteup.tex
rendering PDF

@@
    \includegraphics[width=6in]{../plots/final/defaultPMF.png} 
-   \caption{MAP values and log joint likelihood values for PMF using default playcount values with $\sigma^2 = 1.0$ and $d = 80$}
+   \caption{MAP values and log joint likelihood values for PMF using default play-count values with $\sigma^2 = 1.0$ and $d = 80$}
    \label{defaultPMF}
@@
 Regardless, the best result achieved for PMF in the end was 0.0143, which was only slightly above the popularity baseline.
-Based on literature review \citep{li20122million, McFee:2012:MSD:2187980.2188222}, we have found that this is better or in line with what others have achieved with using PMF on play counts.
+Based on literature review \citep{li2012million, McFee:2012:MSD:2187980.2188222}, we have found that this is better or in line with what others have achieved with using PMF on play counts.
 Nevertheless, this indicates that PMF is not suitable for the MSD dataset and this can be attributed to the implicit feedback issue or simply to the fact that the dataset is too sparse.

### 2015-05-08 algorithms/final/main.py

@@ -1,6 +1,7 @@
 import numpy as np
 import pandas as pd
 import matplotlib as plt
+from matplotlib.ticker import FuncFormatter
 import matplotlib
 import random
 import pandas.io.sql as psql
@@ -90,12 +91,36 @@ omega = [tuple(x) for x in trainsub[trainsub.plays>0][['user_index','song_index'
 omega_test = [tuple(x) for x in testsub[trainsub.plays>0][['user_index','song_index']].values]
 
 
-
 ### Data Exploratory and Plotting Section
+def to_percent(y, position):
+    # Ignore the passed in position. This has the effect of scaling the default
+    # tick locations.
+    s = str(100 * y)
 
+    # The percent symbol needs escaping in latex
+    if matplotlib.rcParams['text.usetex'] == True:
+        return s + r'$\%$'
+    else:
+        return s + '%'
 
+#cumulative sum for plays vs songs
+songplaycsum = songhist.sort('plays')
+songplaycsum['plays'] = songplaycsum['plays'].cumsum()
+songplaycsum['plays'] = songplaycsum['plays']/songplaycsum['plays'].max()
+songplaycsum = songplaycsum.reset_index(drop=True).reset_index()
+songplaycsum.plot('index', 'plays')
+plt.legend().set_visible(False)
+plt.title("Cumulative Sum of Number of Users for each Song")
+plt.ylabel("Perecentage of Users")
+plt.xlabel("Number of Songs")
+#plt.vlines(60000,0,1,linestyles='dotted')
+formatter = FuncFormatter(to_percent)
+plt.gca().yaxis.set_major_formatter(formatter)
+plt.legend().set_visible(False)
+plt.show()
 
-
+print eval.groupby('user_id').count().describe()['plays']
+print eval.groupby('sid').count().describe()['plays']
 
 ### Popularity Baseline Section
 popbase = popularity_baseline(trainsub,omegau_train,omegau_test)

### 2015-05-08 algorithms/final/NMF.py
added algs

@@ -1,0 +1,43 @@
+def NMF(Mtrain, omegau, omegau_test, iter=20, rank = 25):
+    X = Mtrain.T
+
+    rank = 25
+
+    N1 = np.shape(Mtrain)[0]
+    N2 = np.shape(Mtrain)[1]
+
+    W = np.random.uniform(size=[N2,rank])
+    H = np.random.uniform(size=[rank,N1])
+
+    divobj = []
+    for t in xrange(0,iter):
+        temp1 = normalize(np.transpose(W), norm='l1', axis=1)
+        purple = X / (np.dot(W,H) + 10**(-16))
+        H = H * np.dot(temp1,purple)
+        temp2 = normalize(np.transpose(H), norm='l1', axis=0) 
+        purple = X / (np.dot(W,H) + 10**(-16))
+        W = W * np.dot(purple,temp2)
+        temp = np.sum((X * np.log(1/(np.dot(W,H) + 10**(-16)))) + (np.dot(W,H)))
+        divobj.append(temp)
+        
+    plt.plot(divobj)
+    plt.xlabel('Iteration')
+    plt.ylabel('Objective function')
+    plt.title('NMF with Divergence Penalty')
+    
+    rec_user = {}
+    for user in  xrange(0,N1):
+        song_ranks = np.dot(W, H[:,user])
+        rec_user[user] = np.argsort(song_ranks)[::-1]
+
+    apk_sum = 0
+    rec = {}
+    counter = 0
+    for i in omegau_test:
+        if i in omegau:
+            rec[i] = [x for x in rec_user[i] if x not in omegau[i]][0:500]
+            apk_sum += apk(omegau_test[i], rec[i])
+            counter += 1
+            
+    mapval = apk_sum/counter
+    return mapval

### 2015-05-08 algorithms/final/k_means.py

@@ -1,0 +1,80 @@
+def rawkmeans(Mtrain, omegau, omegau_test, song_dict, p2, ncl=10):
+    kmeans_setup = KMeans(n_clusters=ncl)
+    kmeans_model = kmeans_setup.fit_transform(Mtrain)
+
+    ulist = []
+    slist = {}
+    flatsonglist = {}
+    notinOMEGAU = set()
+
+    for i in xrange(0,ncl):
+        cl = i
+        sorted_array = np.argsort(kmeans_model.T[i])[0:10]
+        sort = kmeans_model.T[i, sorted_array]
+        for k in xrange(0,10):
+            uindex = sorted_array[k]
+            if cl in slist:
+                if uindex in omegau:
+                    slist[cl].append(omegau[uindex])
+                else:
+                    #print "This index is not in OMEGA_U Uindex=", uindex
+                    notinOMEGAU.add(uindex)
+            else:
+                if uindex in omegau:
+                    slist[cl] = [(omegau[uindex])]
+                else:
+                    #print "This index is not in OMEGA_U Uindex=", uindex 
+                    notinOMEGAU.add(uindex)
+
+    for cl in xrange(0,10):
+        if cl in slist:
+            flatsonglist[cl] = [item for sublist in slist[cl] for item in sublist]
+
+    # assign one user to one cluster
+    cl_user_dic = defaultdict(list)
+    user_cl_dic = {}    
+
+    #for i in xrange(0, len(kmeans_model)):
+    #    cl_assignment = np.argmin(kmeans_model[i]) 
+    #    cl_user_dic[cl_assignment].append(i)
+
+    #for key, value in cl_user_dic.items():
+    #    print key, len(value)
+
+    for i in xrange(0, len(kmeans_model)):
+        cl_assignment = np.argmin(kmeans_model[i]) 
+        user_cl_dic[i] = cl_assignment
+
+    # for each song how many users have played it
+    cls_songs = {}
+
+    for cl, sindexLIST in flatsonglist.items():
+        nplaycnt = []
+        for sindex in sindexLIST:
+            songh = [key for key, value in song_dict.items() if value == sindex][0]
+            tupole = (sindex, len(p2[songh]))
+        nplaycnt.append(tupole) 
+        sortsongs_t = sorted(nplaycnt, key=lambda x: x[1], reverse=True)
+        sortsonsgs = [i[0] for i in sortsongs_t]
+        cls_songs[cl] = sortsonsgs
+
+    prediction = {}
[…]

### 2015-05-08 writeup/MSDgroupwriteup.tex
references

@@
 
-\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
+\begin{figure}[h] %  figure placement: here, top, bottom, or page
    \centering
@@
 The log joint likelihood chart can be used to verify that the probabilistic matrix factorization algorithm is running properly in that the values are monotonically increasing, which corresponds to a decreasing training error.
-The first graph in Figure /ref{defaultPMF} shows that the log joint likelihood is indeed monotonically increasing and slowed down significantly after the first twenty iterations.
+The first graph in Figure \ref{defaultPMF} shows that the log joint likelihood is indeed monotonically increasing and slowed down significantly after the first twenty iterations.
 The second graph in the same figure shows the MAP values from the same run and it shows that the MAP values decreasing after the maximum at five iterations.
@@
 
-The results from the experiments were plotted in Figure \ref{PMF}.
+The results from the experiments were plotted in Figure \ref{fig:PMF}.
 It was found that the rank $d$ did not contribute significantly to the MAP values especially in settings where the MAP values were close to the popularity baseline.
@@
 
-\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
+\begin{figure}[hbt] %  figure placement: here, top, bottom, or page
    \centering
@@
    \caption{MAP values for different scaling methods and hyperparameters after 30 iterations}
-   \label{PMF}
+   \label{fig:PMF}
 \end{figure}
@@
 Our song-based collaborative filtering technique produced the best results despite being a relatively simple algorithm as compared to matrix factorization.
-Our final results for each technique employed are summarized in the table below.
+Our final results for each technique employed are summarized in Table \ref{tab:map}.
 
@@
 \end{tabular}
-\caption{Final MAP results across various algorithms.}\label{tab:stats}
+\caption{Final MAP results across various algorithms.}\label{tab:map}
 \end{center}
@@
 Apart from the algorithms discussed above, we also explored building a recommendation system based on K-Means Clustering (by both user and song) and Non-Negative Matrix Factorization (NMF).
-In the end, both of these methods yielded worse results as compared to Probabalistic Matrix Factorization so we decided not to explore them further.
+In the end, both of these methods yielded worse results as compared to Probabilistic Matrix Factorization so we decided not to explore them further.
 Next steps include incorporating tags and metadata present in the dataset (e.g.
@@
 
+\newpage
 \bibliography{msdref}

### 2015-05-08 algorithms/final/artist_based.py
Update artist_based.py

@@ -1,6 +1,7 @@
 import numpy as np
 import pandas as pd
 import random
+from collections import OrderedDict
 
 from apk import apk
 

### 2015-05-08 algorithms/final/NMF.py
erin's algs

@@ -1,3 +1,8 @@
+from apk import apk
+import numpy as np
+import matplotlib.pyplot as plt
+from sklearn.preprocessing import normalize
+
 def NMF(Mtrain, omegau, omegau_test, iter=20, rank = 25):
     X = Mtrain.T
 

### 2015-05-08 algorithms/final/k_means.py

@@ -1,4 +1,9 @@
-def rawkmeans(Mtrain, omegau, omegau_test, song_dict, p2, ncl=10):
+from apk import apk
+import numpy as np
+from collections import defaultdict
+from sklearn.cluster import KMeans
+
+def rawkmeans(Mtrain, omegau, omegau_test, omegav, ncl=10):
     kmeans_setup = KMeans(n_clusters=ncl)
     kmeans_model = kmeans_setup.fit_transform(Mtrain)
 
@@ -51,8 +56,7 @@ def rawkmeans(Mtrain, omegau, omegau_test, song_dict, p2, ncl=10):
     for cl, sindexLIST in flatsonglist.items():
         nplaycnt = []
         for sindex in sindexLIST:
-            songh = [key for key, value in song_dict.items() if value == sindex][0]
-            tupole = (sindex, len(p2[songh]))
+            tupole = (sindex, len(omegav[sindex]))
         nplaycnt.append(tupole) 
         sortsongs_t = sorted(nplaycnt, key=lambda x: x[1], reverse=True)
         sortsonsgs = [i[0] for i in sortsongs_t]

### 2015-05-08 algorithms/final/main.py

@@ -1,6 +1,6 @@
 import numpy as np
 import pandas as pd
-import matplotlib as plt
+import matplotlib.pyplot as plt
 from matplotlib.ticker import FuncFormatter
 import matplotlib
 import random
@@ -17,6 +17,9 @@ from artist_based import artist_based
 
 from UserAndSongBasedRec import user_based_rec
 from UserAndSongBasedRec import item_based_rec
+
+from NMF import NMF
+from k_means import rawkmeans
 
 matplotlib.style.use('ggplot')
 
@@ -253,6 +256,9 @@ print user_based_map
 
 
 ### K-Means section
-
+kmeans_map = rawkmeans(M_train, omegau_train, omegau_test, omegav_train, ncl=10):
+print kmeans_map
 
 ### NMF section
+nmf_map = NMF(M_train, omegau_train, omegau_test, iter=20, rank=25):
+print nmf_map

### 2015-05-08 writeup/MSDgroupwriteup.tex

@@
 
-It is also easy to see in the below chart that the majority of the songs are listened to by very few users so we can decrease the number of songs in our subset but not lose all of the play count.
+It is also easy to see in Figure \ref{tab:csongs} that the majority of the songs are listened to by very few users so we can decrease the number of songs in our subset but not lose all of the play count.
 
@@
    \includegraphics[width=4in]{../plots/final/cumsum-songs.png} 
-   \caption{Percentage of users accounted for versus songs with increasing number of plays.}
-   \label{fig:t}
+   \caption{Percentage of users accounted for versus songs with increasing number of plays.}\label{fig:csongs}
 \end{figure}
@@
 
-In terms of scaling schemas, making the data binary produced the best result, with the second best being normalizing the data against the total number of plays for each user.
-Perhaps it can be said that using number of plays introduces more noise than simply considering if the user has played the song or not.
-TF-IDF performed slightly worse than the two, so down-weighting the songs shared more often by users did not produce better recommendations.
-Interestingly enough, normalizing the data against total number of plays for each song performed the worst and this most likely relates to the fact that number of plays is more meaningful relative to each user rather than to each song.
-
-\begin{figure}[hbt] %  figure placement: here, top, bottom, or page
+\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
    \centering
@@
 \end{figure}
+
+In terms of scaling schemas, making the data binary produced the best result, with the second best being normalizing the data against the total number of plays for each user.
+Perhaps it can be said that using number of plays introduces more noise than simply considering if the user has played the song or not.
+TF-IDF performed slightly worse than the two, so down-weighting the songs shared more often by users did not produce better recommendations.
+Interestingly enough, normalizing the data against total number of plays for each song performed the worst and this most likely relates to the fact that number of plays is more meaningful relative to each user rather than to each song.
 

### 2015-05-08 algorithms/final/main.py
Update main.py

@@ -116,7 +116,6 @@ plt.legend().set_visible(False)
 plt.title("Cumulative Sum of Number of Users for each Song")
 plt.ylabel("Perecentage of Users")
 plt.xlabel("Number of Songs")
-#plt.vlines(60000,0,1,linestyles='dotted')
 formatter = FuncFormatter(to_percent)
 plt.gca().yaxis.set_major_formatter(formatter)
 plt.legend().set_visible(False)
@@ -194,8 +193,7 @@ M_train_tfidf = M_train_rownorm*idf
 M_train_binary = M_train.copy()
 M_train_binary[M_train_binary != 0] = 1
 
-# Bunch of plotting stuff
-# First and Second PMF plot
+# First PMF plot
 MAP, L =  MF(M_train,M_test,1.0, omega, omega_test, omegau_train, omegau_test,80,100)
 
 plt.subplot(1,2,1)
@@ -210,7 +208,7 @@ plt.ylabel('Log joint likelihood')
 plt.title('Log joint likelihood across iterations for default PMF')
 plt.show()
 
-# Third PMF plot
+# Second PMF plot
 plotdata = []
 
 iter = 30

### 2015-05-08 writeup/MSDgroupwriteup.tex
small changes to figure placement
FINAL!

@@
 
-It is also easy to see in Figure \ref{tab:csongs} that the majority of the songs are listened to by very few users so we can decrease the number of songs in our subset but not lose all of the play count.
+It is also easy to see in Figure \ref{fig:csongs} that the majority of the songs are listened to by very few users so we can decrease the number of songs in our subset but not lose all of the play count.
 
-\begin{figure}[h] %  figure placement: here, top, bottom, or page
+\begin{figure}[H] %  figure placement: here, top, bottom, or page
    \centering
@@
 
-\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
+\begin{figure}[H] %  figure placement: here, top, bottom, or page
    \centering
@@
 
+\newpage
 \begin{figure}[htbp] %  figure placement: here, top, bottom, or page

### 2015-05-08 algorithms/final/main.py
Update main.py

@@ -254,9 +254,9 @@ print user_based_map
 
 
 ### K-Means section
-kmeans_map = rawkmeans(M_train, omegau_train, omegau_test, omegav_train, ncl=10):
+kmeans_map = rawkmeans(M_train, omegau_train, omegau_test, omegav_train, ncl=10)
 print kmeans_map
 
 ### NMF section
-nmf_map = NMF(M_train, omegau_train, omegau_test, iter=20, rank=25):
+nmf_map = NMF(M_train, omegau_train, omegau_test, iter=20, rank=25)
 print nmf_map

### 2015-05-08 algorithms/final/main.py
Updating main.py

@@ -214,11 +214,11 @@ plotdata = []
 iter = 30
 for var in [100, 10, 1, 0.01, 0.001]:
     for d in [10, 20, 40, 80]:
-        plotdata.append([0,'Default',var,d,MF2(M_train,M_test,var, omegau_train, omegau_test, d, iter)])
-        plotdata.append([1,'Row-norm',var,d,MF2(M_train,M_test,var, omegau_train, omegau_test, d, iter)])
-        plotdata.append([2,'Col-norm',var,d,MF2(M_train,M_test,var, omegau_train, omegau_test, d, iter)])
-        plotdata.append([3,'Binary',var,d,MF2(M_train,M_test,var, omegau_train, omegau_test, d, iter)])
-        plotdata.append([4,'TF-IDF',var,d,MF2(M_train,M_test,var, omegau_train, omegau_test, d, iter)])
+        plotdata.append([0,'Default',var,d,MF2(M_train, M_test, var, omegau_train, omegau_test, d, iter)])
+        plotdata.append([1,'Row-norm',var,d,MF2(M_train_rownorm, M_test, var, omegau_train, omegau_test, d, iter)])
+        plotdata.append([2,'Col-norm',var,d,MF2(M_train_colnorm, M_test, var, omegau_train, omegau_test, d, iter)])
+        plotdata.append([3,'Binary',var,d,MF2(M_train_binary, M_test, var, omegau_train, omegau_test, d, iter)])
+        plotdata.append([4,'TF-IDF',var,d,MF2(M_train_tfidf, M_test, var, omegau_train, omegau_test, d, iter)])
             
 plotdata = pd.DataFrame(plotdata,columns=['method','methodname','variance','d','MAP'])
 
@@ -254,9 +254,9 @@ print user_based_map
 
 
 ### K-Means section
-kmeans_map = rawkmeans(M_train, omegau_train, omegau_test, omegav_train, ncl=10)
+kmeans_map = rawkmeans(M_train, omegau_train, omegau_test, omegav_train, ncl=10):
 print kmeans_map
 
 ### NMF section
-nmf_map = NMF(M_train, omegau_train, omegau_test, iter=20, rank=25)
+nmf_map = NMF(M_train, omegau_train, omegau_test, iter=20, rank=25):
 print nmf_map

### 2015-05-08 writeup/MSDgroupwriteup.tex
oxford commas are the best
Update MSDgroupwriteup.tex
some small grammar edits

@@
 
- \author{Jordan Rosenblum, Justin Law \& Erin Grand}
+ \author{Jordan Rosenblum, Justin Law, \& Erin Grand}
  \affil{Data Science Institute, Columbia University, New York, NY 10027}
@@
 
-It is also easy to see in Figure \ref{fig:csongs} that the majority of the songs are listened to by very few users so we can decrease the number of songs in our subset but not lose all of the play count.
+It is also easy to see in Figure \ref{fig:csongs} that the majority of the songs are listened to by very few users so we can decrease the number of songs in our subset without losing all of the play counts.
 
@@
 This resulted in a dataset which was 10\% the size of the original.
-The sparsity of the data was .0016, meaning that we only had play counts for 0.16\% of the user, song pairs (i.e.
+The sparsity of the data was .0016, meaning that we only had play counts for 0.16\% of the user-song pairs (i.e.
 not all users listen to all songs).
@@
 %\end{figure}
-%
 %
@@
 \begin{equation}
-\mathcal{L} = \sum_{(i,j) \in \Omega} \frac{1}{2\sigma^2} {|| M_{ij} - u_i^\TRANSPOSE  v_j||}^2 - \sum_{i=1}^{N_1} \frac{\lambda}{2} ||u_i^2 || - \sum_{i=1}^{N_2} \frac{\lambda}{2} ||v_j^2 || + \text{constant}
+\mathcal{L} = - \sum_{(i,j) \in \Omega} \frac{1}{2\sigma^2} {|| M_{ij} - u_i^\TRANSPOSE  v_j||}^2 - \sum_{i=1}^{N_1} \frac{\lambda}{2} ||u_i^2 || - \sum_{j=1}^{N_2} \frac{\lambda}{2} ||v_j^2 || + \text{constant}
 \label{eq3} 
@@
 
-There are several hyper-parameters that can be tweaked in order to achieve a better MAP result using matrix factorization.
+There are several hyperparameters that can be tweaked in order to achieve a better MAP result using matrix factorization.
 The main ones that we varied were the variance, the rank $d$ of the factorization, and the number of iterations.
@@
 
-After testing various values of the parameters, it was soon discovered that matrix factorization results were significantly below the popularity baseline and changing the hyper-parameters around did not help much.
+After testing various values of the parameters, it was soon discovered that matrix factorization results were significantly below the popularity baseline and changing the hyperparameters around did not help much.
 The best MAP value of 0.0039 was achieved using a high rank $d$ of 80 over 100 iterations with a variance of 1.0.
@@
 
-One explanation for the very low MAP scores using PMF was that the number of plays that were used are different from what PMF is typically used for which is bounded ratings.
-Using implicit user feedback (number of plays) instead of explicit user feedback (ratings) has been found in literature \citep{hu2008collaborative} to create various issues.
+One explanation for the very low MAP scores using PMF was that the number of plays that were used are different from the typical bounded ratings PMF is typically used on.
+Using implicit user feedback (number of plays) instead of explicit user feedback (ratings) has been shown in literature, such as \citet{hu2008collaborative}, to create various issues.
 One of the main issues is that implicit feedback is inherently noisy and does not measure negative feedback.
-Users may have played a song to simply find out that they dislike the song.
+Listeners may have played a song and simply found out that they disliked the song.
 The other main issue is that the exact number of plays could have a significantly different meaning across different users.
-For example, an avid user who has played a song five times could have a different meaning as compared to a user that has at most only played songs five times.
+For example, an avid listener who has played a song five times could have a different meaning as compared to a listener that has played songs at most five times.
 
@@
 These included normalizing user play counts by the total number of plays for the user, normalizing song play counts by the total number of plays for the song, making the data binary (changing all non-zero play counts to 1), and applying tf-idf.
-While testing out these schemas, we continued to vary the hyper-parameters (i.e.
+While testing out these schemas, we continued to vary the hyperparameters (i.e.
 $d$ and variance) in order to determine if they changed the MAP results significantly.
 
-The results from the experiments were plotted in Figure \ref{fig:PMF}.
-It was found that the rank $d$ did not contribute significantly to the MAP values especially in settings where the MAP values were close to the popularity baseline.
-Higher variance values were found to actually significantly increase the MAP values and it can be argued that the reason why the default PMF works best at variance of 10 instead of 1.0 for the others is that the empirical variance of play count values is at least an order of magnitude higher than the others.
+The results from the experiments are shown in Figure \ref{fig:PMF}.
+It was found that the rank, $d$, did not contribute significantly to the MAP values especially in settings where the MAP values were close to the popularity baseline.
+Higher variance values were found to significantly increase the MAP values and it can be argued that the reason why the default PMF works best at variance of 10 instead of 1.0 for the others is that the empirical variance of play count values is at least an order of magnitude higher than the others.
 This also suggests that the variance used in PMF should scale with the empirical variance, but the exact relationship remains unclear.
 
-\newpage
+%\newpage
 \begin{figure}[htbp] %  figure placement: here, top, bottom, or page
@@
 Perhaps it can be said that using number of plays introduces more noise than simply considering if the user has played the song or not.
-TF-IDF performed slightly worse than the two, so down-weighting the songs shared more often by users did not produce better recommendations.
+Tf-idf performed slightly worse than the two, so down-weighting the songs shared more often by users did not produce better recommendations.
 Interestingly enough, normalizing the data against total number of plays for each song performed the worst and this most likely relates to the fact that number of plays is more meaningful relative to each user rather than to each song.
@@
 However, it is still possible to apply matrix factorization techniques to this dataset through using metadata and this has been demonstrated to be viable by \citet{liangcodebook} which achieved MAP values exceeding 0.1.
-
-%
-%\begin{figure}[ht]
-%\begin{minipage}[b]{0.45\linewidth}
-%\centering
-%\includegraphics[width=\textwidth]{h845BuzFO9CNgAAAABJRU5ErkJggg==.png}
-%\caption{default}
-%\label{fig:figure1}
-%\end{minipage}
-%\hspace{0.5cm}
-%\begin{minipage}[b]{0.45\linewidth}
-%\centering
-%\includegraphics[width=\textwidth]{C9LAAAAABJRU5ErkJggg==.png}
-%\caption{default}
-%\label{fig:figure2}
-%\end{minipage}
-%\end{figure}
 
@@
 
-%\subsection{Artist Based Recommendation}
-
 \section{Results}
@@
 
-\section{Conclusion}
+\section{Future Work}
 Apart from the algorithms discussed above, we also explored building a recommendation system based on K-Means Clustering (by both user and song) and Non-Negative Matrix Factorization (NMF).

### 2015-05-08 lastfm_db.py

@@ -1,82 +1,0 @@
-import sqlite3
-
-conn = sqlite3.connect('lastfm_tags.db')
-
-
-
-# Get list of all tags for a particular track id
-tid = 'TRNIEVD128F147645F'
-print 'We get all tags (with value) for track: %s' % tid
-sql = "SELECT tags.tag, tid_tag.val FROM tid_tag, tids, tags WHERE tags.ROWID=tid_tag.tag AND tid_tag.tid=tids.ROWID and tids.tid='%s'" % tid
-res = conn.execute(sql)
-data_tags = res.fetchall()
-print data_tags
-
-
-conn.close()
-
-
-
-# Create dictionaries for tracks to artists and titles
-f = open('unique_tracks.txt', 'r')
-data = []
-for line in f:
-    data.append(line)
-f.close()
-
-track_to_artist = {}
-track_to_title = {}
-song_to_track = {}
-
-for i in range(len(data)):
-    split = data[i].split('<SEP>')
-    track_to_artist[split[0]] = split[2]
-    track_to_title[split[0]] = split[3]
-    song_to_track[split[1]] = split[0]
-
-
-
-# Map songs id to track id
-f = open('kaggle_visible_evaluation_triplets.txt', 'r')
-data = []
-for line in f:
-    data.append(line)
-f.close()
-    
-
-users = set()
-songs = set()
-
-for i in range(len(data)):
-    split = data[i].split('\t')
-    userhash = split[0]
-    songhash = split[1]
-    users.add(userhash)
-    songs.add(songhash)
-        
-users = list(users) #user hashes
-songs = list(songs) #artist hashes
-
-# make sure that all the songs in the triplet set are in the dictonary
-#i=0
-#for song in songs:
-#    if song not in song_to_track:
-#        i = i + 1
-
-# dictonary from trackID to songID for only the songs in triplets
-
-triplet_song_to_track = {}
-triplet_track_to_artist = {}
-triplet_track_to_title = {}
-
-for song in songs:
-    track = song_to_track[song]
-    triplet_song_to_track[song] = track
-    triplet_track_to_artis = track_to_artist[track]
-    triplet_track_to_title = track_to_title[track]
-
-
-
-
-
-

### 2015-05-08 writeup/MSDgroupwriteup.tex
Rewriting abstract and editing last table

@@
 \begin{abstract}
-We attempt to build a recommendation system for a subset of the Million Song Dataset.
+A recommendation system was attempted to recommend songs to users using data from the Million Song Dataset available on Kaggle.
+The main focus of this paper was using content-based collaborative filtering algorithms that only utilize user feedback in order to predict what songs users may like.
+For benchmarking the algorithms, the same Mean Average Precision truncated at 500 recommended songs used for the original Kaggle competition was utilized.
+It was discovered that probablistic matrix factorization did not improve results much from using a baseline of simply recommending popular songs with MAP values around 0.014, while artist-based popularity along with user-based and item-based collaborative filtering methods yielded much better results exceeding MAP of 0.04.
+
+%We attempt to build a recommendation system for a subset of the Million Song Dataset.
 We explore various algorithms including matrix factorization, user-based collaborative filtering, and item-based collaborative filtering.
@@
 User Based  CF  &  0.0377 \\ 
-Matrix Factorization  &   0.0143  \\ 
+Binary Matrix Factorization  &   0.0143  \\ 
 Top 500 Songs by Count  Baseline &  0.0138  \\ 

### 2015-05-08 writeup/MSDgroupwriteup.tex
small spelling correction, no in depth edits

@@
 For benchmarking the algorithms, the same Mean Average Precision truncated at 500 recommended songs used for the original Kaggle competition was utilized.
-It was discovered that probablistic matrix factorization did not improve results much from using a baseline of simply recommending popular songs with MAP values around 0.014, while artist-based popularity along with user-based and item-based collaborative filtering methods yielded much better results exceeding MAP of 0.04.
+It was discovered that probabilistic matrix factorization did not improve results much from using a baseline of simply recommending popular songs with MAP values around 0.014, while artist-based popularity along with user-based and item-based collaborative filtering methods yielded much better results exceeding MAP of 0.04.
 
@@
 For our project, we explored various song recommendation systems algorithms such as popularity recommendations, artist based recommendations, matrix factorization, and user and item based similarity measures.
-% memory based collaborative filtering
 

### 2015-05-08 writeup/MSDgroupwriteup.tex
Update MSDgroupwriteup.tex

@@
 \begin{abstract}
-A recommendation system was attempted to recommend songs to users using data from the Million Song Dataset available on Kaggle.
-The main focus of this paper was using content-based collaborative filtering algorithms that only utilize user feedback in order to predict what songs users may like.
-For benchmarking the algorithms, the same Mean Average Precision truncated at 500 recommended songs used for the original Kaggle competition was utilized.
-It was discovered that probabilistic matrix factorization did not improve results much from using a baseline of simply recommending popular songs with MAP values around 0.014, while artist-based popularity along with user-based and item-based collaborative filtering methods yielded much better results exceeding MAP of 0.04.
+We attempted to build a recommendation system to recommend songs to users using data from the Million Song Dataset available on Kaggle.
+The main focus of this paper was to use collaborative filtering algorithms that only utilize user feedback in order to predict what songs users may like.
+For benchmarking the algorithms, the same Mean Average Precision truncated at 500 recommended songs used for the original Kaggle competition was chosen.
+It was discovered that probablistic matrix factorization did not improve results much from using a baseline of simply recommending popular songs with MAP values around 0.014, while artist-based popularity along with user-based and item-based collaborative filtering methods yielded much better results exceeding a MAP value of 0.04.
 
@@
 \section{Introduction}
-In the online world, recommendation systems are becoming increasingly important in various industries including retail, entertainment, and even dating. 
-Because of their popularity and usefulness, there are numerous incentives to implement a good recommendation system. 
-In the music industry, companies use recommendation systems to provide a better service to their listeners. 
-With the music industry shifting away from a more traditional distribution model of physical CDs or records, 
-having an efficient and well performing recommendation system is very important.
+With an increasing amount of accessible data, recommendation systems are becoming increasingly viable and important in various industries including retail, entertainment, and even dating.
+Recommendation systems allow for services to extend their reach to customers by allowing their users to discover things of interest.
+Hence, there is a strong interest in implementing a good recommendation system so that their user base can grow.
+It can be said that building such a system is even more important for the music industry because the entire library of artists and songs are far larger compared to something like the movie industry so recommending the right music to a specific user is an even harder task.
+On the other hand, with the shift from a traditional distribution model utilizing physical CDs or records to digital models where users stream or acquire music from internet services, it is now easier than ever to track user preferences and this presents a new opportunity to better predict and recommend songs to users.
 
-In building a recommendation system, we had to take into account various trade-offs in the type of models we chose to fit.
-For example, some methods take more information about the user and the song into account (i.e.
-content-based) while others use a collaborative filtering approach without taking into account user or song characteristics such as age, genre, location, etc.
-In this paper, we focus on collaborative filtering methods.
-There is no set similarity measure between two songs, as evident by the fact that songs by the same artist are often very different. As such, there are many different types of algorithms to explore.
+In building a recommendation system, various trade-offs had to be accounted for in the type of models that need to be fit.
+For example, some methods focus more on the metadata of the users and songs (i.e.
+content-based, such as the user or song characteristics - age, genre, location, etc.
+) while other methods use a collaborative filtering approach where a user feedback system such as a star rating is used without considering other metadata.
+In this paper, the latter approach is the focus.
 
-For our project, we explored various song recommendation systems algorithms such as popularity recommendations, artist based recommendations, matrix factorization, and user and item based similarity measures.
+While songs are typically grouped into genres, it is harder to quantify exactly how similar a song is to another and what characteristics of songs a user actually prefers.
+Two songs by the same artist can produce very different feedback from the same user and thus, determining the exact similarity between two songs is no easy task.
+The benefit of collaborative filtering methods is that we assume that the tastes of a user can be estimated to be similar to another user and thus we can recommend the songs that the second user has enjoyed to the first user.
 
@@
 \subsection{Testing}
-We tested our algorithms against each other using a mean average precision 
-score for each method.
-The mean average precision (MAP) calculation came directly from the Kaggle MSD challenge \citep{McFee:2012:MSD:2187980.2188222}.
-The score was used to rank the competition entries.
+We benchmarked our algorithms using a mean average precision score truncated at 500 (MAP@500) for each algorithm.
+This choice of benchmark was mainly due to MAP@500 also being used as the main method to evaluate algorithms submitted for the Kaggle MSD challenge \citep{McFee:2012:MSD:2187980.2188222}.
 
-Specifically, we used MAP@500 scores which take into account the first 500 recommendations given to each user.
-Generally, the score looks at the precision between a list of recommendations for a user and that user's test set.
+Specifically, MAP@500 scores take into account the first 500 recommendations given to each user.
+Generally, the score looks at the precision for a list of 500 song recommendations for a user and calculates the amount of correct recommendations (true positives) against the total amount of recommended songs.
 This was done by calculating the average precision at each element in the list of recommended items (i.e.
@@
 for each $k \in \{1, ..., 500\}$, calculate the precision and average them together).
-We then averaged the score over all users.
-Intuitively, the score looks at the percentage of recommendations that were in-line with the test set of songs actually listened to but also takes order into account.
-So it is preferable to recommend the songs in the user's test set earlier in the list of 500.
-The MAP score is similar to the AUC of the ROC curve discussed in class except it takes ordering of the recommendations into account, which is an important factor to consider in a recommendation system.
+Then, the average score is simply calculated by taking the average score over all users.
 
-We used two different baseline benchmarks \citep{McFee:2012:MSD:2187980.2188222} to compare our MAP scores against.
-The first was to simply recommend the top 500 most popular songs to every user.
-This naive method resulted in a MAP score of 0.0138 (top 500 songs based on number of users who listened to the song) or 0.0126 (top 500 songs based on number of plays the song had).
-The second method was to first recommend a user the most popular songs of artists the user already had already played (but different songs from that artist).
-If this did not result in enough recommendations, we would then recommend the overall most popular songs as well at the end (similar to baseline method 1).
-Although this method is conservative in that it does not go beyond the tastes the user already has in the training set, it resulted in a significant improvement in the MAP score to 0.0448.
+Intuitively, the score looks at the percentage of recommendations that were in-line with what the user actually liked using the testing set that were not used by the algorithms during training.
+This score also takes into account the order of the songs recommended so it is desirable to have the song with the highest confidence to be liked at the beginning of the recommendation list.
+For example, if a user had a test set list of five songs, recommending these five songs at the very beginning would give a MAP score of 100 while recommending them at the very end of the 500 songs in reverse order would only produce a MAP score of 0.0060.
+The MAP score can be said to be similar to the AUC of the ROC curve discussed in class without taking into account the order of recommendations.
+
+Two different baseline benchmarks based on \citet{McFee:2012:MSD:2187980.2188222} were used.
+The first was to simply to recommend the top 500 most popular songs in the training dataset to every user with no user personalization.
+This naive method produced a MAP score of 0.0138 if the top 500 songs were based on the numbers of users who listened to the song or 0.0126 if the top 500 songs were based on the number of total plays the song had.
+The second baseline used had the bare amount of user personalization where the most popular songs from the artists that the user had also listened to were ordered first.
+Although this method was also relatively simple and naive, it resulted in a very significant improvement of the MAP score to produce a value of 0.0448.
 

### 2015-05-08 writeup/MSDgroupwriteup.tex
Update MSDgroupwriteup.tex

@@
 \subsection{Data and Statistics}
-The main source of data for this project was the Million Song Data Set (MSD)  \citep{Bertin-Mahieux2011}.
+The main data source for this project is the Million Song Data Set (MSD)  \citep{Bertin-Mahieux2011}.
 We used the sample data set from the Kaggle competition which contains tuples of user id, song id, and play count.
-We also used data which links songs to artists through track ids.
+We also used data which links songs to artists through track ids to calculate the artist-based popularity baseline MAP score.
 The links to these datasets can be found in the Appendix.
@@
 \hline
-Measurement & Plays Across Songs & Plays Across Users \\
+Measurement & # of Listeners Per Song & # of Songs Played Per User \\
 \hline
-Count &  163206 & 110000 \\
 Mean  &     8.890194 & 13.190300  \\
@@
 
-It is also easy to see in Figure \ref{fig:csongs} that the majority of the songs are listened to by very few users so we can decrease the number of songs in our subset without losing all of the play counts.
+The scale of the data was too large to explore on our computers for the algorithms that we used, so we subset the data down to a size easily handled by an average laptop.
+
+It can be seen in Figure \ref{fig:csongs} that the majority of the songs are listened to by very few users so we can decrease the number of songs in our subset without losing all of the play counts.
 
@@
 
-The scale of the data was too large to explore on our computers, so we subset the data down to a size easily handled by an average laptop.
-
 We choose to subset the data to users who have listened to more than 27 songs and songs which were listened to by more than 22 users.
@@
 not all users listen to all songs).
-Despite subsetting the data, though, we are still dealing with a matrix of $8130$ users by $11861$ songs, so some of our algorithms will take a fair amount of time to run.
+Despite subsetting the data, though, we are still dealing with $8130$ users and $11861$ songs, so some of our algorithms will take a fair amount of time to run.
 
@@
 In order to perform the analysis, we had to fill in training and testing matrices which contained play counts for user and song pairs.
-We did a random selection of data elements to split between the training (80\%) and testing (20\%) sets.
+We did a random sampling of 80\%/20\% to produce training and testing sets.
 Next, we implemented various algorithms on the dataset in order to recommend songs to users, including matrix factorization, user-based collaborative filtering, and item-based collaborative filtering.
@@
 The log joint likelihood chart can be used to verify that the probabilistic matrix factorization algorithm is running properly in that the values are monotonically increasing, which corresponds to a decreasing training error.
+
 The first graph in Figure \ref{defaultPMF} shows that the log joint likelihood is indeed monotonically increasing and slowed down significantly after the first twenty iterations.
@@
 It was found that the rank, $d$, did not contribute significantly to the MAP values especially in settings where the MAP values were close to the popularity baseline.
-Higher variance values were found to significantly increase the MAP values and it can be argued that the reason why the default PMF works best at variance of 10 instead of 1.0 for the others is that the empirical variance of play count values is at least an order of magnitude higher than the others.
+Higher variance values were found to significantly increase the MAP values.
+It can be argued that the reason why the default PMF works best at variance of 10 instead of 1.0 for the others is that the empirical variance of play count values is at least an order of magnitude higher than the others.
 This also suggests that the variance used in PMF should scale with the empirical variance, but the exact relationship remains unclear.
@@
 The sum is a proxy for how likely the user is to like that particular song.
-We went through that process for all songs and recommended the user the top 500 songs which had the highest scores.
+This was repeated for all songs and recommended the user the top 500 songs which had the highest scores.
 This method led to a MAP value of 0.0377.
@@
 Then, for each user, we found all the songs listened to.
-Next, we got the most similar songs to each of those songs listened to by the user.
+Next, we calculated the most similar songs to each of those songs listened to by the user.
 If the same similar songs came up multiple times, we added the weights together.
@@
 \section{Results}
-After comparing the various algorithms, the simpler models gave us the best results.
+After comparing the various algorithms, we found that item-based collaborative filtering gave us the best result.
 Our most complicated model, matrix factorization, gave us results barely above the popularity baseline.
 On the other hand, simply recommending users songs by artists they have already listened to (the artist-based popularity baseline) resulted in a significantly better recommendation system.
-Our song-based collaborative filtering technique produced the best results despite being a relatively simple algorithm as compared to matrix factorization.
+Our song-based collaborative filtering technique produced the best results despite it being more naive as compared to matrix factorization.
 Our final results for each technique employed are summarized in Table \ref{tab:map}.

### 2015-05-08 writeup/MSDgroupwriteup.tex
updated PDF, some small changes

@@
 For benchmarking the algorithms, the same Mean Average Precision truncated at 500 recommended songs used for the original Kaggle competition was chosen.
-It was discovered that probablistic matrix factorization did not improve results much from using a baseline of simply recommending popular songs with MAP values around 0.014, while artist-based popularity along with user-based and item-based collaborative filtering methods yielded much better results exceeding a MAP value of 0.04.
+It was discovered that probabilistic matrix factorization did not improve results much from using a baseline of simply recommending popular songs with MAP values around 0.014, while artist-based popularity along with user-based and item-based collaborative filtering methods yielded much better results exceeding a MAP value of 0.04.
 
@@
 We also used data which links songs to artists through track ids to calculate the artist-based popularity baseline MAP score.
-The links to these datasets can be found in the Appendix.
+The links to all these datasets can be found in the Appendix.
 
@@
 \hline
-Measurement & # of Listeners Per Song & # of Songs Played Per User \\
+Measurement & \# of Listeners Per Song & \# of Songs Played Per User \\
 \hline
@@
 The scale of the data was too large to explore on our computers for the algorithms that we used, so we subset the data down to a size easily handled by an average laptop. 
-
-It can be seen in Figure \ref{fig:csongs} that the majority of the songs are listened to by very few users so we can decrease the number of songs in our subset without losing all of the play counts.
+It can be seen in Figure \ref{fig:csongs} that the majority of the songs are listened to by very few users we were able to decrease the number of songs in our subset without losing all of the play counts.
 

### 2015-05-08 writeup/MSDgroupwriteup.tex
Update MSDgroupwriteup.tex

@@
 The main focus of this paper was to use collaborative filtering algorithms that only utilize user feedback in order to predict what songs users may like.
-For benchmarking the algorithms, the same Mean Average Precision truncated at 500 recommended songs used for the original Kaggle competition was chosen.
-It was discovered that probabilistic matrix factorization did not improve results much from using a baseline of simply recommending popular songs with MAP values around 0.014, while artist-based popularity along with user-based and item-based collaborative filtering methods yielded much better results exceeding a MAP value of 0.04.
+For benchmarking the algorithms, we used the same Mean Average Precision score truncated at 500 recommended songs that was used in the original Kaggle competition.
+It was discovered that probabilistic matrix factorization with a MAP value of 0.014 did not improve results much from using a baseline of simply recommending popular songs, while artist-based popularity along with user-based and item-based collaborative filtering methods yielded much better results with the best method giving a MAP value of 0.048.
 
@@
 \section{Introduction}
-With an increasing amount of accessible data, recommendation systems are becoming increasingly viable and important in various industries including retail, entertainment, and even dating.
+With an increasing amount of accessible data, recommendation systems are becoming more viable and important in various industries including retail, entertainment, and even dating.
 Recommendation systems allow for services to extend their reach to customers by allowing their users to discover things of interest.

### 2015-05-08 writeup/MSDgroupwriteup.tex
Update MSDgroupwriteup.tex

@@
 Hence, there is a strong interest in implementing a good recommendation system so that their user base can grow.
-It can be said that building such a system is even more important for the music industry because the entire library of artists and songs are far larger compared to something like the movie industry so recommending the right music to a specific user is an even harder task.
+It can be said that building such a system is even more important for the music industry because the entire library of artists and songs are far larger as compared to something like the movie industry so recommending the right music to a specific user is an even harder task.
 On the other hand, with the shift from a traditional distribution model utilizing physical CDs or records to digital models where users stream or acquire music from internet services, it is now easier than ever to track user preferences and this presents a new opportunity to better predict and recommend songs to users.
@@
 For example, some methods focus more on the metadata of the users and songs (i.e.
-content-based, such as the user or song characteristics - age, genre, location, etc.
-) while other methods use a collaborative filtering approach where a user feedback system such as a star rating is used without considering other metadata.
+content-based) such as user or song characteristics (e.g.
+age, genre, location, etc.) while other methods use a collaborative filtering approach where a user feedback system such as a star rating is used without considering other metadata.
 In this paper, the latter approach is the focus.
@@
 The scale of the data was too large to explore on our computers for the algorithms that we used, so we subset the data down to a size easily handled by an average laptop. 
-It can be seen in Figure \ref{fig:csongs} that the majority of the songs are listened to by very few users we were able to decrease the number of songs in our subset without losing all of the play counts.
+It can be seen in Figure \ref{fig:csongs} that the majority of the songs were listened to by very few users so we were able to decrease the number of songs in our subset without losing all of the play counts.
 
@@
 
-We choose to subset the data to users who have listened to more than 27 songs and songs which were listened to by more than 22 users.
+We chose to subset the data to users who have listened to more than 27 songs and songs which were listened to by more than 22 users.
 This resulted in a dataset which was 10\% the size of the original.
@@
 This score also takes into account the order of the songs recommended so it is desirable to have the song with the highest confidence to be liked at the beginning of the recommendation list.
-For example, if a user had a test set list of five songs, recommending these five songs at the very beginning would give a MAP score of 100 while recommending them at the very end of the 500 songs in reverse order would only produce a MAP score of 0.0060.
+For example, if a user had a test set list of five songs, recommending these five songs at the very beginning would give a MAP score of 1 while recommending them at the very end of the 500 songs in reverse order would only produce a MAP score of 0.0060.
 The MAP score can be said to be similar to the AUC of the ROC curve discussed in class without taking into account the order of recommendations.
@@
 Two different baseline benchmarks based on \citet{McFee:2012:MSD:2187980.2188222} were used.
-The first was to simply to recommend the top 500 most popular songs in the training dataset to every user with no user personalization.
+The first was to simply recommend the top 500 most popular songs in the training dataset to every user with no user personalization.
 This naive method produced a MAP score of 0.0138 if the top 500 songs were based on the numbers of users who listened to the song or 0.0126 if the top 500 songs were based on the number of total plays the song had.
@@
 Training and testing matrices were constructed, of which both have $N_1$ users (rows denoted by $u_i$) and $N_2$ songs (columns denoted by $v_j$).
-As users have only listened to a fraction of all the songs available, the matrices are sparse, containing zeros in all entries except for those in which a user (a user representing a single row) has listened to.
+As users have only listened to a fraction of all the songs available, the matrices are sparse, containing zeros in all entries except for those in which a user (representing a single row) has listened to.
 The goal of matrix factorization is to factor the training matrix into the product of two matrices $U$ and $V$.
@@
 There are several hyperparameters that can be tweaked in order to achieve a better MAP result using matrix factorization.
-The main ones that we varied were the variance, the rank $d$ of the factorization, and the number of iterations.
+The main ones that we explored were the variance, the rank $d$ of the factorization, and the number of iterations.
 As the project was being done using Python and the main computational package (\emph{scikit-learn}) did not have probabilistic matrix factorization functions, we decided to apply the algorithm that we wrote ourselves for a different class \citep{koren2009matrix}.

### 2015-05-08 writeup/MSDgroupwriteup.tex
Update MSDgroupwriteup.tex

@@
 
-One explanation for the very low MAP scores using PMF was that the number of plays that were used are different from the typical bounded ratings PMF is typically used on.
+One explanation for the very low MAP scores using PMF was that the number of plays that were used were different from the typical bounded ratings PMF is typically used on.
 Using implicit user feedback (number of plays) instead of explicit user feedback (ratings) has been shown in literature, such as \citet{hu2008collaborative}, to create various issues.

### 2015-05-08 writeup/MSDgroupwriteup.tex
Update MSDgroupwriteup.tex
Updating .tex
Removed "in reverse" when describing the MAP@500 because it's not true.
Added a sentence about sparsity in PMF. Still figuring out how to talk
about running time and space requirements for PMF

@@
 This score also takes into account the order of the songs recommended so it is desirable to have the song with the highest confidence to be liked at the beginning of the recommendation list.
-For example, if a user had a test set list of five songs, recommending these five songs at the very beginning would give a MAP score of 1 while recommending them at the very end of the 500 songs in reverse order would only produce a MAP score of 0.0060.
+For example, if a user had a test set list of five songs, recommending these five songs at the very beginning would give a MAP score of 1 while recommending them at the very end of the 500 songs would only produce a MAP score of 0.0060.
 The MAP score can be said to be similar to the AUC of the ROC curve discussed in class without taking into account the order of recommendations.
@@
 Nevertheless, this indicates that PMF is not suitable for the MSD dataset and this can be attributed to the implicit feedback issue or simply to the fact that the dataset is too sparse.
+A dataset that is too sparse prevents PMF from learning enough information about how users rank songs to produce more accurate recommendations.
 However, it is still possible to apply matrix factorization techniques to this dataset through using metadata and this has been demonstrated to be viable by \citet{liangcodebook} which achieved MAP values exceeding 0.1.
@@
 On the other hand, simply recommending users songs by artists they have already listened to (the artist-based popularity baseline) resulted in a significantly better recommendation system.
+This is probably because users tend to listen to multiple songs by the same artist.
 Our song-based collaborative filtering technique produced the best results despite it being more naive as compared to matrix factorization.

### 2015-05-08 writeup/MSDgroupwriteup.tex
Updating .tex

@@
 \subsection{Probabilistic Matrix Factorization}
-Matrix factorization is a type of collaborative filtering algorithm that can be used to build a recommendation system for users based on user feedback of objects (in this case songs).
+Probabilistic matrix factorization (PMF) is a type of collaborative filtering algorithm that can be used to build a recommendation system for users based on user feedback of objects (in this case songs).
 This allows us to recommend songs to users based on their listening history without the need for content based approaches.
@@
 As the project was being done using Python and the main computational package (\emph{scikit-learn}) did not have probabilistic matrix factorization functions, we decided to apply the algorithm that we wrote ourselves for a different class \citep{koren2009matrix}.
+It should be noted that while PMF algorithms can be parallelized to significantly improve running time and scaled to handle large amounts of data, being limited to a single machine to run these computations meant that sub setting the data was necessary in order to actually execute the algorithm successfully while limiting computational time to durations that were feasible for exploration.
 

### 2015-05-09 writeup/MSDgroupwriteup.tex
Changing sub setting to subsetting

@@
 As the project was being done using Python and the main computational package (\emph{scikit-learn}) did not have probabilistic matrix factorization functions, we decided to apply the algorithm that we wrote ourselves for a different class \citep{koren2009matrix}.
-It should be noted that while PMF algorithms can be parallelized to significantly improve running time and scaled to handle large amounts of data, being limited to a single machine to run these computations meant that sub setting the data was necessary in order to actually execute the algorithm successfully while limiting computational time to durations that were feasible for exploration.
+It should be noted that while PMF algorithms can be parallelized to significantly improve running time and scaled to handle large amounts of data, being limited to a single machine to run these computations meant that subsetting the data was necessary in order to actually execute the algorithm successfully while limiting computational time to durations that were feasible for exploration.
 

### 2015-05-09 algorithms/final/matrix_factorization.py
Update the MF and main .py

@@ -3,7 +3,7 @@ import pandas as pd
 import random
 from apk import apk
 
-def MF(Mtrain, Mtest, var, omega, omega_test, omegau, omegau_test, d=20, iter=20):
+def MF(Mtrain, var, omega, omega_test, omegau, omegau_test, d, iter):
     
     N1 = np.shape(Mtrain)[0]
     N2 = np.shape(Mtrain)[1]
@@ -73,7 +73,7 @@ def MF(Mtrain, Mtest, var, omega, omega_test, omegau, omegau_test, d=20, iter=20
 
     return MAP, loglik
 
-def MF2(Mtrain, Mtest, var, omegau, omegau_test, d, iter):
+def MF2(Mtrain, var, omegau, omegau_test, d, iter):
     
     N1 = np.shape(Mtrain)[0]
     N2 = np.shape(Mtrain)[1]

### 2015-05-09 algorithms/final/main.py
Update matrix_factorization.py and main.py

@@ -194,7 +194,7 @@ M_train_binary = M_train.copy()
 M_train_binary[M_train_binary != 0] = 1
 
 # First PMF plot
-MAP, L =  MF(M_train,M_test,1.0, omega, omega_test, omegau_train, omegau_test,80,100)
+MAP, L =  MF(M_train, 1.0, omega, omega_test, omegau_train, omegau_test, 80, 100)
 
 plt.subplot(1,2,1)
 plt.plot([0] + range(4,100,5), MAP)
@@ -212,13 +212,13 @@ plt.show()
 plotdata = []
 
 iter = 30
-for var in [100, 10, 1, 0.01, 0.001]:
+for var in [100, 10, 1, 0.1, 0.01, 0.001]:
     for d in [10, 20, 40, 80]:
-        plotdata.append([0,'Default',var,d,MF2(M_train, M_test, var, omegau_train, omegau_test, d, iter)])
-        plotdata.append([1,'Row-norm',var,d,MF2(M_train_rownorm, M_test, var, omegau_train, omegau_test, d, iter)])
-        plotdata.append([2,'Col-norm',var,d,MF2(M_train_colnorm, M_test, var, omegau_train, omegau_test, d, iter)])
-        plotdata.append([3,'Binary',var,d,MF2(M_train_binary, M_test, var, omegau_train, omegau_test, d, iter)])
-        plotdata.append([4,'TF-IDF',var,d,MF2(M_train_tfidf, M_test, var, omegau_train, omegau_test, d, iter)])
+        plotdata.append([0,'Default',var,d,MF2(M_train, var, omegau_train, omegau_test, d, iter)])
+        plotdata.append([1,'Row-norm',var,d,MF2(M_train_rownorm, var, omegau_train, omegau_test, d, iter)])
+        plotdata.append([2,'Col-norm',var,d,MF2(M_train_colnorm, var, omegau_train, omegau_test, d, iter)])
+        plotdata.append([3,'Binary',var,d,MF2(M_train_binary, var, omegau_train, omegau_test, d, iter)])
+        plotdata.append([4,'TF-IDF',var,d,MF2(M_train_tfidf, var, omegau_train, omegau_test, d, iter)])
             
 plotdata = pd.DataFrame(plotdata,columns=['method','methodname','variance','d','MAP'])
 

### 2015-05-09 algorithms/final/matrix_factorization.py

@@ -73,6 +73,7 @@ def MF(Mtrain, var, omega, omega_test, omegau, omegau_test, d, iter):
 
     return MAP, loglik
 
+# MF without calculating log likelihood to reduce compute time
 def MF2(Mtrain, var, omegau, omegau_test, d, iter):
     
     N1 = np.shape(Mtrain)[0]
